---
title: "AE 13: Foundations of Inference"
author: "Your Name"
date: "10.07.2021"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

```{r load-packages}
library(tidyverse)
library(knitr)
library(infer)
```

## Bulletin

-   Due tomorrow:

    -   lab 05

-   Upcoming:

    - [project](https://sta199-f21-001.netlify.app/project/project.html) proposal (work on in lab tomorrow)
    - `hw03` released today, due next Thursday
    
- Misc:

    - Exam 01 grades are out (check gradescope for feedback)
    - 1 week from yesterday for regrade requests
    - make sure you make three commits per application exercise

## Midterm grade calculation

Your midterm grade is computed using the assignment grades reported on sakai but according to the weight scale described on the [syllabus](https://sta199-f21-001.netlify.app/syllabus/syllabus.pdf)

For purposes of the midterm grade, no lowest assignment scores were dropped.

```
mean ae score (as %) * 2.5
+ mean quiz score (as %) * 5
+ mean lab score (as %) * 15
+ exam 01 score (as %) * 17.5
+ hw score (as %) * 25

all divided by 65.
```

## Learning goals

- Understand how statistical inference is used to draw conclusions from sample data 
- Define and understand sampling variability
- Introduce bootstrapping 


## Vocabulary activity

What's the difference between a parameter and a statistic?


Is a parameter "random" or "fixed"? Typically "Unknown" or "known"? Why?


## Bootstrapping activity: Rent in Manhattan 

On a given day in 2018, twenty one-bedroom apartments were randomly selected on Craigslist Manhattan from apartments listed as "by owner". The data are in the `manhattan` data frame.  We will use this sample to conduct inference on the typical rent of 1 bedroom apartments in Manhattan. 

```{r load-data}
manhattan <- read_csv("data/manhattan.csv")
```

### Exercise 1

Visualize the distribution of `rent`. Is the mean or the median a better measure of typical rent of one-bedroom apartments in Manhattan?


### Exercise 2

What is a point estimate of the typical rent? 


### Exercise 3

Let's bootstrap! 

- To bootstrap we will sample with replacement by drawing a value from the bowl. 
- How many draws do we need for our bootstrap sample? 

Fill in  the values from the bootstrap sample conducted in class. Once the values are filled in, uncomment the code. 
```{r boot-sample}
# class_bootstrap <- c()

```


## Exercise 4

- About what value do you expect the bootstrap statistic to take? 
- Calculate the statistic from the bootstrap sample. 

```{r calc-boot-stat}
# add code
```

Does this statistic align with your expectations? 

<hr> 

Here we've take one bootstrap sample, but in practice we will need about 10,000 - 15,000! In the next lecture we will discuss how we can calculate bootstrap samples using the `infer` package in R. 

**Sneak peek!**
```{r}
boot_dist = manhattan %>%
  # specify the variable of interest
  specify(response = rent) %>% 
  # generate 15000 bootstrap samples
  generate(reps = 15000, type = "bootstrap") %>% 
  # calculate the statistic of each bootstrap sample
  calculate(stat = "mean")

boot_dist %>%
  ggplot(aes(x = stat)) +
  geom_histogram() + 
  labs(x = "mean rent")
  
```

