---
title: "AE 21: Regression pt. 2 and multiple regression"
author: "YOUR NAME GOES HERE"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, warning = FALSE, message  = FALSE}
library(tidyverse)
library(tidymodels)
library(scatterplot3d)
```

# Bulletin

- Due today
  - [todo]

- Upcoming
  - [todo]


# Learning goals

- Gain proficiency with multiple linear regression
- Understand what a dummy variable is

To begin, we'll work with a new set of pokemon data.

```{r load-pokemon}
pokemon <- read_csv("data/pokemon150.csv")
```

## Exercise 1) 

Previously we looked at building a model to predict a pokemon's hit points using height.

Last time we noticed *both* height *and* weight are correlated with a pokemon's hit points.

Other variables may be correlated with `hp`, e.g. a pokemon's legendary status. 

Do legendary pokemon have higher `hp` than non-legendary pokemon? Compare mean `hp` between groups to support your answer.

```{r hp-vs-legendary}
pokemon %>%
  group_by(is_legendary) %>%
  summarize(mean_hp = mean(hp), n = n())
```


Write down a model to predict a pokemon's hitpoints based on their height, weight, legendary status (use $x$, $y$, $\beta$ notation). Define each variable.

$y$: hit points, (`hp`)
$x_1$: height, (`height_m`)
$x_2$: weight, (`weight_kg`)
$x_3$: legendary, (`is_legendary`)

$y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_3$

## Exercise 2) 

Use `tidymodel` syntax to build a linear model and estimate each $\beta$.

```{r hp-lm}
# code here
linear_reg() %>%
  set_engine("lm") %>%
  fit(hp ~ height_m + weight_kg + is_legendary, data = pokemon) %>%
  tidy()
```
Interpret the meaning of your estimates and write a brief description below.

The intercept tells us the approximate hp of a pokemon with no height, no weight and no legendary status This is meaningless in context.

**All else held constant**

- the `height_m` coefficient tells us the that pokemon 1 m taller have on average 5.1 more hp.

- the `weight_kg` coefficient tells us that a pokemon gains on average .08 hp per additional kg they weigh.

- the `is_legendary` coefficient tells us that legendary pokemon has on average 5.3 more hp than a non-legendary pokemon.

## Exercise 3)

Do the coefficients match your expectations for the previous exercise? Why or why not? Write code to explore any oddities further.

At first glance, it is strange that the coefficient of `is_legendary` is only 5.3 considering the difference in means between groups is 42.

To explain this

```{r explain-legendary-coefficient}
# code here
pokemon %>%
  group_by(is_legendary) %>%
  summarize(mean_weight = mean(weight_kg),
            mean_height = mean(height_m),
            mean_hp = mean(hp))
```

we see that legendary pokemon also weigh more and are taller on average than non-legnedaries, helping explain the weaker contribution conferred by legendary status.

## Exercise 4)

Some think that certain pokemon types have higher `hp` than others.

```{r hp-vs-type}
pokemon %>%
  group_by(type_1) %>%
  summarize(mean_hp = mean(hp), n = n())
```
Construct a linear model in `R` to determine the effect of pokemon type on `hp`. 

```{r hp-type-lm}
# code here
linear_reg() %>%
  set_engine("lm") %>%
  fit(hp ~ type_1, data = pokemon) %>%
  tidy()
```


Interpret the meaning of your estimates and write a brief description below. Are any types missing?

- the intercept tells us the approximate baseline hp of a pokemon with bug type (set to the baseline by default since it's the first variable).

- the `type_1` coefficients tell us the average contribution of that pokemon's type on hp.

Because the category `type_1` has many levels, it is encoded as several dummy variables. The idea is that each type is now yes/no for each pokemon.

Given the output of your code, write down the linear model (use $x$, $y$, $\beta$ notation). Define each variable.

$y = \beta_0 + x_1 \beta_1 + x_2 \beta_2 + x_3 \beta_3 + \ldots x_17 \beta_17$

$x_1$ through $x_17$ are boolean variables telling us whether or not a pokemon is a dark type,..., all the way until water type.
