---
title: "AE 20: Intro to Linear Models"
author: "YOUR NAME GOES HERE"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, warning = FALSE, message  = FALSE}
library(tidyverse)
library(tidymodels)
library(scatterplot3d)
```

# Bulletin

- Upcoming
  - prep quiz due Thursday
  - lab 08 due Friday
  - Exam 02 next Friday through Tuesday. Exam 02 will cover topics up until Tues Nov. 9th.


# Learning goals

-   Understand the language and notation of linear modeling. 
-   Use `tidymodels` to make inference under a linear regression model

To begin, let's load the data. Again, we'll work with the pokemon data from last week.

```{r load-pokemon}
pokemon <- read_csv("data/pokemon.csv")
```

# The language of linear modeling

What is a linear model?

A linear model is a simple way to mathematically model the relationship between two or more observed phenomena. 

## Exercise 1) 

Is there a relationship between a pokemon's size and their hit points (hp)? Specifically, does a pokemon's height/weight tell us something about how many hit points they have?

Create two scatterplots, one for hp vs height and another illustrating hp vs weight.
```{r visualize-relationships}
# code here
```

Describe the relationship you see here.


Label the following:

Response variable:

Explanatory variable(s):

## Exercise 2)

For now, let's focus on just two of these variables, namely hp and height.

Write down a model (using $x$, $y$, $\beta$ notation) that describes a linear relationship between hp and height. Define each variable.

[write here]

## Exercise 3)

[Click here](https://seeing-theory.brown.edu/regression-analysis/index.html#section1) to interact with an ordinary least squares (OLS) linear regression model.

Select I and move the data points around.

Describe what you see.

## Exercise 4)

Now let's fit a ordinary least squares `lm`.

Use functions from the preparation video/slides to fit a linear model to `hp` and `height` as described in exercise 2. Save the model in a variable entitled `hp_height_fit`.

```{r fit-hp-height}
# code here

```

Write out the equation of the fitted model.

Uncomment and fill in blanks below to visualize the linear model on top of the scatterplot from Exercise 1.
```{r visualize-the-model}
#hp_height_fit_aug <- augment(hp_height_fit$fit) # need $fit for plotting

# hp_height_fit_aug %>%
#   ggplot() +
#   geom_point(aes(___, ___)) +
#   geom_line(aes(x = ___, y = .fitted), size = 0.75, color = "darkred") +
#   theme_minimal()
```


## Exercise 5)

Use the equation of the fitted model to predict the hp of an unknown pokemon with a height of 1.1 meters.
```{r predict-from-equation}
# code here
```
If you were asked to predict the hp of a pokemon that is 2.5 meters tall, is this model appropriate? Why or why not?

## Exercise 6)

What about a linear model with multiple explanatory variables? 

```{r plot-all-three, warning = FALSE}
scatterplot3d(pokemon[,c("weight_kg", "height_m", "hp")], pch = 19, color="steelblue") 
```

Describe what you see.

```{r look-at-hyperplane, message = FALSE, warning = FALSE, echo = FALSE}
plot3d = scatterplot3d(pokemon[,c("weight_kg", "height_m", "hp")], pch = 19, color="steelblue") 
model  = lm(pokemon$hp ~ pokemon$weight_kg + pokemon$height_m)
plot3d$plane3d(model)
```
