---
title: "The Four Nations"
author: "STA199"
date: "9.16.2021"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

### Bulletin
-   finish ae07 then ae08
-   on your radar:
    -   hw 02 released today
    -   quiz 03 released today
    -   group lab tomorrow
    -   exam next week
-   due today
    -   hw 01 due tonight at 11:59pm
-   due Tuesday
    -   prep
    -   lab 04
    -   quiz 03

### Practice

```{r load-packages, message = FALSE}
library(tidyverse)
```

Load in the data

```{r}
class = read_csv("data/classData.csv")
```

Take a look at the data

```{r look-at-data}
glimpse(class)
```

1. Which nation is the most represented? Least represented? Make a visualization to illustrate.
```{r plot-1}
```


2. Are any lab sections predicted to do better than others on an airbender test? Is there a correlation between lab section and predicted score?

```{r}
```

3. Why?

```{r}

```

4. How could you see this on the plot above?

```{r}

```

