---
title: "Data Visualization II"
output: html_document
author: "STA199"
date: "9-2-2021"
editor_options: 
  chunk_output_type: console
---

## Bulletin
  - Due Friday (tomorrow) at 11:59pm: 
    - lab 01 (gradescope, github) **reminder: label questions on gradescope pdf**
    - ae 03 (github)
    
  - Due Sunday at 11:59pm:
    - ae 04 (github)
    
  - Due Tuesday by class time:
    - prepare (videos)
    - lab 02 (gradescope, github)
    - quiz (sakai)

# Main Ideas

- There are different types of variables. 
- Visualizations and summaries of variables must be consistent with the 
variable type.

# Lecture Notes and Exercises

Load the `tidyverse` and `datasauRus` packages

```{r message = FALSE}
library(tidyverse)
library(datasauRus)
```

There are two types of variables **numeric** and **categorical**.

### Types of variables

Numerical variables can be classified as either **continuous** or **discrete**. 
Continuous numeric variables have an infinite number of values between any 
two values. Discrete numeric variables have a countable number of values.

- height
- number of siblings

Categorical variables can be classified as either **nominal** or **ordinal**. 
Ordinal variables have a natural ordering.

- hair color
- education

### Numeric Variables

To describe the distribution of a numeric we will use the properties below.

- shape
  - skewness: right-skewed, left-skewed, symmetric
  - modality: unimodal, bimodal, multimodal, uniform
- center: mean (`mean`), median (`median`)
- spread: range (`range`), standard deviation (`sd`), interquartile range (`IQR`)
- outliers: observations outside the pattern of the data

We will continue our investigation of home prices in Minneapolis, Minnesota.

```{r load-data, message = FALSE}
mn_homes <- read_csv("data/mn_homes.csv")
```

Add a `glimpse` to the code chunk below and identify the following variables as
numeric continuous, numeric discrete, categorical ordinal, or categorical
nominal. 

- area
- beds
- community

```{r glimpse-data}
glimpse(mn_homes$community)
```

The `summary` command is also useful in looking at numerical variables. Use this command to look at the numeric variables from the previous chunk.

```{r summarize}
summary(mn_homes$beds)
```

We can use a **histogram** to summarize a numeric variable.

```{r histogram}
ggplot(data = mn_homes, 
       mapping = aes(x = salesprice)) + 
   geom_histogram(bins = 25)
```

A **density plot** is another option. We just connect the boxes in a histogram
with a smooth curve.

```{r density-plot}
ggplot(data = mn_homes, 
       mapping = aes(x = salesprice)) + 
   geom_density()
```

Side-by-side **boxplots** are helpful to visualize the distribution of a numeric
variable across the levels of a categorical variable.

```{r boxplots}
ggplot(data = mn_homes, 
       mapping = aes(x = community, y = salesprice)) + 
       geom_boxplot() + coord_flip() + 
       labs(main= "Sales Price by Community", x= "Community", y="Sales Price")
```

**Question:** What is `coord_flip()` doing in the code chunk above? Try removing it to see.

### Categorical Variables

**Bar plots** allow us to visualize categorical variables.

```{r bar-plot}
ggplot(data = mn_homes) + 
  geom_bar(mapping = aes(x = community)) + coord_flip() + 
  labs(main= "Homes by Community", x= "Community", y="Number of Homes")
```

**Segmented bar plots** can be used to visualize two categorical variables.

```{r segmented-bar-plot}
library(viridis)
ggplot(data = mn_homes, mapping = aes(x = community, fill = fireplace)) + 
  geom_bar() +
  coord_flip() + 
  scale_fill_viridis(discrete=TRUE, option = "D", name="Fireplace?") +
  labs(main= "Fireplaces by Community", x= "Community", y="Number of Homes")
```

```{r segmented-bar-plot-fill}
ggplot(data = mn_homes, mapping = aes(x = community, fill = fireplace)) + 
  geom_bar(position = "fill") + coord_flip() + 
  scale_fill_viridis(discrete=TRUE, option = "D", name="Fireplace?") +
  labs(main= "Percentage of Homes with a Fireplace by Community", x=
  "Community", y="Percentage of Homes")
```

**Question:** Which of the above two visualizations do you prefer? Why? Is this answer always the same?

There is something wrong with each of the plots below. Run the code for each plot, read the error, then identify and fix the problem.

```{r fix-errors, eval = FALSE}
ggplot(data = mn_homes) + 
  geom_point(mapping = aes(x = lotsize, y = salesprice,
                           shape = 21, size = .85))
ggplot(data = mn_homes) + 
  geom_point(x = lotsize, y = area), shape = 21, size = .85)
ggplot(data = mn_homes) +
  geom_point(mapping = aes(x = lotsize, y = area), color=community), size = 0.85)
ggplot(data = mn_homes) +
  geom_point(mapping = aes(x = 1otsize, y = area))
```

General principles for effective data visualization

- keep it simple
- use color effectively
- tell a story

Why is data visualization important? We will illustrate using the 
`datasaurus_dozen` data from the `datasauRus` package.

```{r glimpse-dinos}
glimpse(datasaurus_dozen)
```

The code below calculates the correlation, mean of y, mean of x, standard deviation of y, and standard deviation of x for each of the 13 datasets.

**Question:** What do you notice? 

```{r summarize-dinos}
datasaurus_dozen %>% 
   group_by(dataset) %>%
   summarize(r = cor(x, y), 
             mean_y = mean(y),
             mean_x = mean(x),
             sd_x = sd(x),
             sd_y = sd(y))
```

Let's visualize the relationships

```{r visualize-dinos}
ggplot(data = datasaurus_dozen, 
       mapping = aes(x = x, y = y)) + 
   geom_point(size = .5) + 
   facet_wrap( ~ dataset)
```

**Question:** Why is visualization important?

## Practice

(1) Modify the code outline to create a faceted histogram examining the 
distribution of year built within each community.

When you are finished, remove `eval = FALSE` and knit the file to see the changes.

Note: depending on the variable you use, you might want to change the bin width as I do below.

```{r}
ggplot(data = mn_homes, mapping = aes(x = area)) +
  geom_histogram(binwidth = 100) +
  facet_wrap(~ community) +
  labs(x = "Area", 
      title = "Which Communities Have the Largest Homes?", 
      subtitle = "Faceted by Community")
```

### Additional Resources

- https://ggplot2.tidyverse.org/
- https://raw.githubusercontent.com/rstudio/cheatsheets/master/data-visualization-2.1.pdf
- http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html
- https://medium.com/bbc-visual-and-data-journalism/how-the-bbc-visual-and-data-journalism-team-works-with-graphics-in-r-ed0b35693535
- https://ggplot2-book.org/
= https://ggplot2.tidyverse.org/reference/geom_histogram.html
- https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf
- https://github.com/GraphicsPrinciples/CheatSheet/blob/master/NVSCheatSheet.pdf