---
title: "AE 02: Part 2, Bechdel Test + R Markdown"
subtitle: "due Friday, May 13 at 9:29am"
author: "Bora Jin"
editor_options:
  chunk_output_type: console
output: 
  html_document: 
    toc: true
    toc_float: true
---

**Your task is to fill in the blanks denoted by `___`.**

## Introduction

In this mini analysis we work with the data used in the FiveThirtyEight story titled ["The Dollar-And-Cents Case Against Hollywood's Exclusion of Women"](https://fivethirtyeight.com/features/the-dollar-and-cents-case-against-hollywoods-exclusion-of-women/). The data contains information about if movies pass the Bechdel test which checks the following three criteria: (1) it has to have at least two women in it, who (2) who talk to each other, about (3) something besides a man.

## Data and Packages 

We start with loading the packages we'll use.

```{r ex-package, message=FALSE}
library(fivethirtyeight) # for data
library(tidyverse) 
```

The dataset contains information on `r nrow(bechdel)` movies released between `r min(bechdel$year)` and `r max(bechdel$year)`. However, we'll focus our analysis on movies released between 2000 and 2013. We first make a new dataset with year between 2000 and 2013 by the pipe operator `%>%` and `filter()`. The function `filter()` is to filter observations meeting the given condition, and you can think of the pipe operator `%>%` as a "then". Hence, the following code translates to "you take the bechdel data, then you filter the data with year between 2000 and 2013". 

```{r data-0013}
bechdel00_13 <- bechdel %>% 
  filter(between(year, 2000, 2013))
```

Let's preview our data with the `glimpse()` function:
```{r glmpse-data}
glimpse(bechdel00_13)
```
There are `___` such movies.

The financial variables we'll focus on are the following:

-   `budget_2013`: Budget in 2013 inflation adjusted dollars
-   `domgross_2013`: Domestic gross (US) in 2013 inflation adjusted dollars
-   `intgross_2013`: Total International (i.e., worldwide) gross in 2013 inflation adjusted dollars

And we'll also use the `binary` and `clean_test` variables for **grouping**.

## Analysis

### Part 1

Let's take a look at how median budget and gross vary by whether the movie passed the Bechdel test, which is stored in the `binary` variable.

```{r bechdel-budget, message = FALSE}
bechdel00_13 %>%
  group_by(binary) %>%
  summarise(med_budget = median(budget_2013),
            med_domgross = median(domgross_2013, na.rm = TRUE),
            med_intgross = median(intgross_2013, na.rm = TRUE))
```

### Part 2

Next, let's take a look at how median budget and gross vary by a more detailed indicator of the Bechdel test result.
This information is stored in the `clean_test` variable, which takes on the following values:

-   `ok` = passes test
-   `dubious`
-   `men` = women only talk about men
-   `notalk` = women don't talk to each other
-   `nowomen` = fewer than two women

```{r cleantest-budget, message = FALSE}
bechdel00_13 %>%
  # group_by(___) %>%
  summarise(med_budget = median(budget_2013),
            med_domgross = median(domgross_2013, na.rm = TRUE),
            med_intgross = median(intgross_2013, na.rm = TRUE))
```

### Part 3

In order to evaluate how return on investment varies among movies that pass and fail the Bechdel test, we'll first create a new variable called `roi` as the ratio of the gross to budget.

```{r roi}
bechdel00_13 <- bechdel00_13 %>%
  mutate(roi = (intgross_2013 + domgross_2013) / budget_2013)
```

Let's see which movies have the highest return on investment.

```{r arrange-roi}
bechdel00_13 %>%
  arrange(desc(roi)) %>% 
  select(title, roi, year)
```

### Part 4

Below is a visualization of the return on investment by test result, however it's difficult to see the distributions due to a few extreme observations.

```{r boxplot, warning = FALSE}
ggplot(data = bechdel00_13, 
       mapping = aes(x = clean_test, y = roi, color = binary)) +
  geom_boxplot() +
  labs(title = "Return on investment vs. Bechdel test result",
       x = "Detailed Bechdel result",
       y = "___",
       color = "Binary Bechdel result")
```

What are those movies with *very* high returns on investment?

```{r high-roi}
bechdel00_13 %>%
  filter(roi > 150) %>%
  select(title, budget_2013, domgross_2013, year)
```

Zooming in on the movies with `roi < ___` provides a better view of how the medians across the categories compare:

```{r betterviz-roi, warning = FALSE}
ggplot(data = bechdel00_13, mapping = aes(x = clean_test, y = roi, fill = binary)) +
  geom_boxplot() +
  labs(title = "___",
       subtitle = "___", # Something about zooming in to a certain level
       x = "___",
       y = "___",
       fill = "___") + 
  coord_cartesian(ylim = c(0, 15)) # zooming in without dropping data 
```

## Submitting Application Exercises

- Once you have completed the activity, push your final changes to your GitHub repo. 
- Make sure you committed at least *three* times. 
- Check that your repo is updated on GitHub, and that’s all you need to do to submit application exercises for participation.

## References

This assignment was adapted from [Bechdel](https://github.com/rstudio-education/datascience-box/blob/main/course-materials/application-exercises/ae-02-bechdel-rmarkdown/bechdel.Rmd) exercise.

