---
title: "Flint Water Crisis"
author: "Your Name"
date: "01.12.2022"
output: 
  html_document 
---

## R basics: vocabulary

A **variable** is a container for storing data.

`<-` and `=` are known as "assignment operators" and they save a value (or collection of values) as a **variable**.

```{r ex-variable}
y = 2
x = 1:10
```
x is a variable that contains _

When you type the name of an object into the console, or at the end of a code chunk, it **prints** to the screen. "Print to screen" just means "show on screen"

```{r ex-print-to-screen}
x
```
A **function** is literally just a bunch of instructions. Most often, we will work with functions that take input (also called **arguments**) and return an output.
```{r ex-function}
mean(x)
```

`mean()` is a function that takes a numeric input and returns the mean of that input

All functions in `R` have `()` and inside those parentheses is where inputs (aka "arguments") go.

**Data frames** are lists of variables. Typically the columns are called "variables", while the rows are thought of as "observations"

```{r ex-data, eval=F}
data(mtcars)
mtcars
```
`mtcars` is a data frame with _ variables and _ observations.

Finally, a **package** (also called a **library**) is just a collection of functions and/or data.

```{r ex-package, eval=F}
# library(dadjokes)
```


## Packages

```{r load-packages, warning = FALSE, message=FALSE}
library(tidyverse)
```

## Introduction

The data set consists of 271 homes sampled with three water lead contaminant values at designated time points. The lead content is in parts per billion (ppb). Additionally, some location data is given about each home.

## Data

To get started, read in the `flint.csv` file using the function `read_csv`. First, use the Upload button under Files.

```{r flint, warning = FALSE, message = FALSE}
flint <- read_csv("data/flint.csv")
```

In this file, there are five variables:

- **id**: sample ID number
- **zip**: ZIP code in Flint of the sample's location
- **ward**: ward in Flint of the sample's location
- **draw**: which time point the water was sampled from
- **lead**: lead content in parts per billion

Let's preview the data with the `glimpse()` function:

```{r glimpse-data, echo = TRUE}
glimpse(flint)
```

## Analysis

### Part 1

Let’s see how many samples were taken from each zip code.

```{r count-by-zip, echo = TRUE}
flint %>%               # data
  group_by(zip) %>%     # perform a grouping by zip code
  count()               # count occurrences
```

Which ZIP code had the most samples drawn?

### Part 2

Next, let's look at the mean and median lead contaminant values for each zip code and draw combination. We have eight zip codes and samples taken at three times. How many combinations do we have?

```{r mean-ppb-by-zip-draw, echo = TRUE}
flint %>% 
  group_by(zip, draw) %>% 
  summarise(mean_pb = mean(lead))
```

```{r median-ppb-by-zip-draw, echo = TRUE}
flint %>% 
  group_by(zip, draw) %>% 
  summarise(median_pb = median(lead))
```

How many rows are in each of two above data frames?

### Part 3

Modify the code below to compute the mean and median lead contaminant values for zip code 48503 at the first draw. What should you put in for `draw == "_____"`? Don’t forget to uncomment the second line of code.

```{r, echo = TRUE}
flint %>% 
  # filter(zip == 48503, draw == "_____") %>% 
  summarise(mean_pb = mean(lead),
            median_pb = median(lead))
```

### Part 4

Let’s make some plots, where we will focus on zip codes 48503, 48504, 48505, 48506, and 48507. We will restrict our attention to samples with lead values less than 1,000 ppb.

```{r, echo = TRUE}
flint_focus <- flint %>% 
  filter(zip %in% 48503:48507, lead < 1000)
```

Below are side-by-side box plots for the three flushing times in each of the five zip codes we considered. Add `x` and `y` labels; add a title by inserting `title = "title_name"` inside the `labs()` function.

```{r}
ggplot(data = flint_focus, aes(x = factor(zip), y = lead)) +
  geom_boxplot(aes(fill = factor(draw))) +
  labs(x = "--------", y = "--------", fill = "Flushing time") +
  scale_fill_discrete(breaks = c("first", "second", "third"),
                      labels = c("0 (sec)", "45 (sec)", "120 (sec)")) +
  coord_flip() +
  theme_bw()
```

Add labels for `x`, `y`, a `title`, and `subtitle` to the code below to update the corresponding plot.

```{r}
ggplot(data = flint_focus, aes(x = factor(zip), y = lead)) +
  geom_boxplot(aes(fill = factor(draw))) + 
  labs(x = "--------", y = "--------", fill = "Flushing time",
       subtitle = "--------") +
  scale_fill_discrete(breaks = c("first", "second", "third"),
                      labels = c("0 (sec)", "45 (sec)", "120 (sec)")) +
  coord_flip(ylim = c(0, 50)) +
  theme_bw()
```

What is the difference between the two plots?

## References

1. Langkjaer-Bain, R. (2017). The murky tale of Flint’s deceptive water data. Significance, 14: 16-21.
