---
title: "AE 06: Data Wrangling II"
output: html_document
date: "01.26.2022"
editor_options: 
  chunk_output_type: inline
---

# Bulletin
- Homework 01 due Friday
- Finish `ae05`, due Friday
- Lab 02 was due Monday
- Academic Honesty

# Today

By the end of the today you will combine data from different sources using a well-chosen join function.

# Lecture Notes and Exercises

```{r message = FALSE, warning = FALSE}
library(tidyverse)
```

Instead of working with a single dataset, usually you will have to work with many different related datasets. To answer research questions using related datasets, we need to develop tools to join datasets together.

There are many possible types of joins. All have the format 
`something_join(x, y)`.

- `inner_join()`: join all rows from `x` where there are matching values in `y`. 
Return all combinations in case of multiple matches
- `left_join()`: include all rows from `x`
- `right_join()`: include all rows from `y`
- `full_join()`: include all rows in `x` or `y`
- `semi_join()`: return all rows from `x` with match in `y`
- `anti_join()`: return all rows from `x` without a match in `y`

```{r test-joins}
x <- tibble(value = c(100, 200, 300),
            xcol = c("x1", "x2", "x3"))
y <- tibble(value = c(100, 200, 400),
            ycol = c("y1", "y2", "y4"))
x
y
```

We will demonstrate each of the joins on these small, toy datasets. 

```{r inner-join}
inner_join(x, y)
```

```{r left-join}
left_join(x, y)
```

```{r right-join}
right_join(x, y)
```

```{r full-join}
full_join(x, y)
```

```{r semi-join}
semi_join(x, y)
```

```{r anti-join}
anti_join(x, y)
```

How do the join functions above know to join `x` and `y` by `value`? Examine the names to find out.

```{r names}
names(x)
names(y)
```

We will again work with data from the `nycflights13` package. 

```{r load-flights-data, message=FALSE, warning=FALSE}
library(nycflights13)
```

Examine the documentation for the datasets `airports`, `flights`, and `planes`.

**Question:** How are these datasets related? Suppose you wanted to make a map of the route of every flight. What variables would you need from which datasets?

Join flights to airports. Note these two datasets have no variables in common so we will have to specify the variable to join by using `by = `. Check out the documentation for more information.

```{r flights-join}
flights %>% 
  left_join(airports, by = c("dest" = "faa"))
```

## Practice

(1) Create a new dataset `dest_delays` with the median arrival delay for each destination. Note this question does not require you to use joins.

```{r make-dest-delays}
```

(2) Create a new dataset by joining `dest_delays` and `airports`. Only include observations that have both delay and airport information. Note `dest_delays` and `airports` have no variables in common so you will need to specify the variables to join using `by` as in the example above.

```{r new-dataset}
```

**Question:** Are all of the observations in `dest_delays` included in the new dataset you created by joining `dest_delays` and `airports`? Use an appropriate join function to investigate this issue and determine what is going on here.


```{r anti-join2}
```

(3) Is there a relationship between the age of a plane and its delays? The plane tail number is given in the `tailnum` variable in the `flights` dataset. The year the plane was manufactured is given in the `year` variable in the `planes` dataset. 

- Step #1: Start by finding the average arrival delay for each plane and store the resulting dataset in `plane_delays`.

```{r step-1}
```

- Step #2: Join `plane_delays` to the `planes` data using an appropriate join and then use mutate to create an age variable. Note this data is from 2013.

```{r step-2}
```

- Step #3: Finally, create an effective visualization of the data.

```{r step-3}
```

### Additional Resources

- https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf