---
title: "AE 10: Introduction to Probability"
subtitle: "due Thursday, May 26 at 9:29am"
author: "Bora Jin"
editor_options:
  chunk_output_type: console
output: 
  html_document: 
    toc: true
    toc_float: true
---

## Getting Started 

- Complete this [survey](https://forms.gle/FQiDDMQ2Nmfzsdf2A)!

- Wait until we finish watching a video

  - Clone the repository entitled "ae10-GitHubUsername" at [course GitHub organization page](https://github.com/sta199-summer22) on your [RStudio](https://vm-manage.oit.duke.edu/containers).
  - Open the `.Rmd` file and replace "Your Name" with your name. 

## Introduction 

Today we load `knitr` to use `kable()` function which neatly displays tables:

```{r load-packages, message = FALSE}
library(tidyverse)
library(knitr)
```

We will also use a new function `pivot_wider()` in `dplyr` package ($\in$ `tidyverse`):

- `r emo::ji("movie_camera")` [Video](https://www.youtube.com/watch?v=x3KM5uxaFdI) and  [Slides](https://rstudio-education.github.io/datascience-box/course-materials/slides/u2-d09-tidying/u2-d09-tidying.html#1)

For this Application Exercise, we will look at our newly collected data.
Remove `eval = FALSE` to read data. 

```{r load-data, eval = FALSE}
sta199 <- read_csv("data/sta199-ae10.csv")
```

The dataset includes

- `year`: Year in school
- `animal`: Whether you prefer cats or dogs
- `movie`: Favorite movie genre
- `stat_major`: statistical science or not

## Events, Sample Space 

### Part 1

Give two examples of an event from the dataset.

### Part 2

Let's take a look at favorite movie genre. Note that we have categorized genres 
so that each person can only have one favorite genre. 

**Q- ** What is the sample space for favorite movie genre? You can use code to 
identify the sample space. 

```{r movie-sample-space}
# code here
```

### Part 3

**Q- ** How large is the sample space of any individual's response? 

```{r response-sample-size}
# code here
```

The sample space for the four survey questions contains $4 \times 2 \times 4 \times 2 = 64$ different outcomes.

## Probability 

### Part 4

Let's make a table that includes `year`, the number of students in each, 
and the associated probabilities. 

**Q- ** What is the probability a randomly selected STA199 student is a freshman?

```{r year}
# code here
```

### Part 5

**Q- ** What is the probability a randomly selected STA199 student favors cats?
Answer it with a table that includes `animal`, the number of students who prefer each, 
and the associated probabilities. 

```{r animal}
# code here
```

### Part 6

**Q- ** What is the probability a randomly selected STA199 student is not a 
senior **and** prefers dogs? 

Let $A$ be the event that someone is not a senior and prefers dogs.

```{r year-animal} 
# code here
```

### Part 7

**Q- ** What is the probability a randomly selected STA199 student likes either 
action movies or comedy **and** is a statsci major? 

Let $B$ be the event someone is an action or comedy movie lover and a statsci major.

```{r movie-major} 
# code here
```

### Part 8

Now we examine the relationship between favorite animal and favorite movie. 
Let's make a table of the number of students for every combination of favorite 
animals and movie genres. 

```{r animal-movie}
# code here
```

Using `pivot_wider()`, we'll reformat the data into a **contingency table**, 
a table frequently used to study the association between two categorical variables. 
In this contingency table, each row will represent an animal, each column will represent a movie, 
and each cell is the number of students have a particular combination of animal and movie.  

```{r pivot-wider}
# code here
```

**Q - ** How many students in STA199 like sci-fi movies? 

**Q - ** How many students in STA199 like dogs and dramas?

### Practice

For each of the following exercises:

- Calculate the probability using a relevant contingency table.  
- Then write code to check your answer using the `sta199` data frame and `dplyr` functions. 

(1) Relationship between `year` and `major`

**Q - **What is the probability a randomly selected STA199 student is a junior **or** not a statsci major?

```{r ex1-1}
# code here
```

```{r ex1-2}
# code here
```

**Q - ** What is the probability a randomly selected STA199 student is a statsci major?

```{r ex1-3}
# code here
```

(2) Relationship between `animal` and `major`

**Q - **What is the probability a randomly selected STA199 student likes dogs **and** a statsci major?

```{r ex2-1}
# code here
```

```{r ex2-2}
# code here
```

(3) Relationship between `year` and `movie`

**Q - ** What is the probability a randomly selected STA199 student is a senior **or** does not pick sci-fi as the favorite movie genre? 

```{r ex3-1}
# code here
```

```{r ex3-2}
# code here
```

## Submitting Application Exercises

- Once you have completed the activity, push your final changes to your GitHub repo. 
- Make sure you committed at least *three* times. 
- Check that your repo is updated on GitHub, and that’s all you need to do to submit application exercises for participation.

