---
title: 'Conditional Probability'
output:
  html_document:
    df_print: paged
date: "02.11.2022"
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

# Bulletin
- Due today
  - lab 04, double check group members are linked on gradescope
  - QUIZ: mid-course survey on sakai (find under quizzes)
- [Project proposal](https://sta199-sp22-003.netlify.app/project/project.html)

# Today

By the end of today you will

- be able to define and compute **marginal**, **joint** and **conditional** probabilities
- identify when events are independent
- apply Bayes' theorem to examine COVID test specificity

```{r load-packages}
library(tidyverse)
library(knitr)
```

## Definitions

Let A and B be events.

-   **Marginal probability:** The probability an event occurs regardless of values of the other event
    - *P(A)*
    - Example: What's the probability a student in STA199 favors dogs?

-   **Joint probability**: The probability two or more events simultaneously occur
    - Example: What's the probability a student is a junior **and** favors dogs?
    - *P(A and B)* 
  
-   **Conditional probability**: The probability an event occurs given the other has occurred
    - *P(A|B)* or *P(B|A)*
    - Eample: What is the probability a student is a junior **given** they favor dogs?

-   **Independent events**: Knowing one event has occurred does not lead to any change in the probability we assign to another event.
    - *P(A|B) = P(A)* or *P(B|A) = P(B)*
    - Example: P(Junior | dogs) = P(junior)

## Bayes' Theorem

The global coronavirus pandemic illustrates the need for accurate testing of COVID-19, as its extreme infectivity poses a significant public health threat. Due to the time-sensitive nature of the situation, the FDA enacted emergency authorization of a number of serological tests for COVID-19 in 2020. Full details of these tests may be found on its website [here](https://www.fda.gov/medical-devices/coronavirus-disease-2019-covid-19-emergency-use-authorizations-medical-devices/eua-authorized-serology-test-performance).

We will define the following events:

-   **Pos**: The event the Alinity test returns positive.
-   **Neg**: The event the Alinity test returns negative.
-   **Covid**: The event a person has COVID
-   **No Covid**: The event a person does not have COVID

The Abbott Alinity test has an estimated sensitivity of 100%, *P(Pos \| Covid) = 1*, and specificity of 99%, *P(Neg \| No Covid) = 0.99*.

Suppose the prevalence of COVID-19 in the general population is about 2%, *P(Covid) = 0.02*.

**Exercise 1**: Use the Hypothetical 10,000 to calculate the probability a person has COVID given they get a positive test result, i.e. *P(Covid | Pos)*.

|           | Covid | No Covid | Total |
|-----------|-------|----------|-------|
| **Pos**   |       |          |       |
| **Neg**   |       |          |       |
| **Total** |       |          | 10000 |


**Exercise 2** Use Bayes' Theorem to calculate *P(Covid|Pos)*. 

## Simpson's paradox

This example comes from *Confounding and Simpson's paradox*^[Julious, S A, and M A Mullee. “Confounding and Simpson's paradox.” BMJ (Clinical research ed.) vol. 309,6967 (1994): 1480-1. doi:10.1136/bmj.309.6967.1480] by Julious and Mullee.

The data examines 901 individuals with diabetes and includes the following variables

- `insulin_dep`: whether or not the patient has *insulin dependent* or *non-insulin dependent* diabetes
- `less_than_40`: whether or not the individual is less than 40 years old
- `survival`: whether or not the individual survived the length of the study

```{r load-data}
diabetes = read_csv("data/diabetes.csv")
```
One might be interested in the mortality associated with each type of diabetes.

```{r aggregate}
# code here
```
Is the aggregate reported above misleading and if so, why?

```{r separate}
# code here
```

