---
title: "AE 15: Hypothesis Testing 1"
subtitle: "due Friday, June 3 at 9:29am"
author: "Bora Jin"
editor_options:
  chunk_output_type: console
output: 
  html_document: 
    toc: true
    toc_float: true
---

## Getting Started 

- Clone the repository entitled "ae15-GitHubUsername" at [course GitHub organization page](https://github.com/sta199-summer22) on your [RStudio](https://cmgr.oit.duke.edu/containers).

- Open the `.Rmd` file and replace "Your Name" with your name. 

```{r load-packages0, message = FALSE}
library(tidyverse)
```

## Coin Flips 

Suppose we flipped a coin 20 times and observed 17 tails and 3 heads.

```{r coin-flips}
coin_flips <- c(0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 1, 0, 0, 0)
```

But this looks suspicious - we would think: "what's the probability of 17T and 3H if the coin is fair?"

### Part 1: Manual testing with a known distribution

**Step 1:** Define the hypotheses.

Without realizing it, we are setting up a null hypothesis:

- $H_0$: The coin is fair. 

Mathematically we state this null hypothesis in terms of $p$ (the probability of the coin landing heads):

**Q - **What should $p$ equal if the coin is fair?

- $H_0$: $p = ?$ 

This is a hypothesis about "truth". The "true" parameter $p$ that belongs to this particular coin when flipped infinitely many times. We don't know the true $p$, we make an assumption about it.

**Step 2:** Summarize data. 

All we have is a **sample** of coin flips. 

**Q - ** What is our **statistic** for the proportion of heads? 

$\hat{p} = ?$

In words, $\hat{p}$ is called "sample-p" or "p-hat". 
See a brief table of mathematical notation for common quantities of interest below:

| Quantity           | Parameter  | Statistic   |
|--------------------|------------|-------------|
| Mean               | $\mu$      | $\bar{x}$   |
| Variance           | $\sigma^2$ | $s^2$       |
| Standard deviation | $\sigma$   | $s$         |
| Median             | $M$        | $\tilde{x}$ |
| Proportion         | $p$        | $\hat{p}$   |

**Step 1 (revisited):** Define the hypotheses.

Now we ask ourselves: "Is $\hat{p}$ so different from our assumption that we have to reject our assumption as true?"

To answer this, the alternative hypothesis matters. Our options are:

- 1) $H_1: p < 0.5$, the coin is biased to land tails
- 2) $H_1: p > 0.5$, the coin is biased to land heads
- 3) $H_1: p \neq 0.5$, the coin is biased

Choice of hypothesis affects calculation of p-value in the next step and thus affects our conclusions. For this reason, one should choose an alternative hypothesis before looking at the data.

**Step 3:** Assess the evidence. 

In this step, we compute a p-value.
By definition, the p-value is the probability of seeing what we observed (in this case, $\hat{p} = 3/20$) or something more extreme if $H_0$ is true. The direction of alternative hypotheses decide what "something more extreme" is. 

The following probabilities are the p-value for each alternative hypothesis given above: 

- 1) The probability of observing a $\hat{p} \leq 3/20$ if the null is true?
- 2) The probability of observing a $\hat{p} \geq 3/20$ if the null is true?
- 3) The probability of observing a $\hat{p} \leq 3/20$ **or** $\hat{p} \geq 17/20$ if the null is true?

We will find the three p-values using simulations. First, we need to know what 20 coin flips would look like "under the null". In other words, "if the null was actually true".

**Q - ** Create 200000 random numbers from Binomial(20, 0.5). This generates a distribution for $\hat{p}$ (the sample proportion of heads out of 20 coin flips) given the coin is fair. This is called the null distribution. 

```{r null-flips, eval = FALSE}
set.seed(123)
null_heads <- rbinom(____, size = __, prob = __)
null <- data.frame(null_heads)
```

**Q - ** Create a new column `p` that is the proportion of heads in each row and create a histogram with `bins = 20`. 

```{r null-flips-viz}
```

**Q - ** Find the three p-values using `null`. 

```{r p-value}
## 1)

## 2)

## 3)
```

**Step 4:** Make a conclusion. 

**Q - ** Make a testing decision at the significance level 0.05 ($\alpha = 0.05$) for each alternative:

- 1) $H_1: p < 0.5$, the coin is biased to land tails
- 2) $H_1: p > 0.5$, the coin is biased to land heads
- 3) $H_1: p \neq 0.5$, the coin is biased

## Rent in Manhattan 

We will use the following data one more time to conduct hypothesis testing on the typical rent of one-bedroom apartments in Manhattan: 

```{r load-data, message = FALSE}
manhattan <- read_csv("data/manhattan.csv")
```

### Part 2: Testing using functions in `infer`

We were lucky in Part 1 because we knew the theoretical distribution of the number of heads from $n$ coin flips with a probability $p$. The distribution was Binomial($n$, $p$), and `rbinom()` function generated random numbers from Binomial distributions. 

Often, however, you do not know the exact distribution of the statistics of interest. In that case, we can use **bootstrapping** to capture variability in sample statistics from the original sample. 

In Part 2, we will conduct hypothesis testing in a `tidy` way. We load the `tidymodels` package that contains the `infer` package.  

```{r load-packages, message = FALSE}
library(tidymodels)
```

**Step 1:** Define the hypotheses.

**Q - ** Suppose you are interested in whether the mean rent of one-bedroom apartments in Manhattan is less than $3000. State the null and alternative hypotheses. 

**Step 2:** Summarize data. 

**Q - ** What is an appropriate statistic for $\mu$? Find its value. 

```{r point-est, eval = FALSE}
xbar <- manhattan %>% 
  summarize(xbar = ____) %>% 
  pull()
```

The function `pull()` extracts a single column from a data frame. In the code chunk above, without `pull()` the resulting output is a `tibble`, while with `pull()` the output becomes a scalar.  

**Step 3:** Assess the evidence. 

**Q - **Let's first approximate the null distribution by completing the code chunk below. Use 1,000 reps for the in-class activity. You will use about 15,000 reps for assignments outside of class.

```{r null-dist, eval = FALSE}
set.seed(602)

# save resulting bootstrap distribution
null_dist <- manhattan %>%
  # specify the variable of interest
  specify(response = ____) %>%
  # declare the null hypothesis 
  hypothesize(null = ____, mu = ____) %>% 
  # generate reps bootstrap samples
  generate(reps = 1000, type = ____) %>%
  # calculate the statistic of each bootstrap sample
  calculate(stat = ____)
```

Compared to bootstrapping for estimation, we need one additional function for testing: 

-   `hypothesize`: declare a null hypothesis about a **point estimate**, e.g. proportion, mean, median, variance (today) or about independence (tomorrow) by setting `null = point` or `null = independence`. 
  - If making a hypothesis about a point estimate, set the null value, e.g., `mu = 10`.
  - This shifts the bootstrap distribution to be centered at the null value. 
  - This is necessary for hypothesis testing unlike constructing confidence interval because we assume "the null hypothesis is actually true". 
  
**Q - ** Visualize the null distribution. Fill in the blanks in the code chunk below. 

```{r null-viz, eval = FALSE}
visualize(null_dist) +
  labs(x = "______", y = "______", 
       title = "Simulated null distribution of sample means",
       subtitle = "[about the assumption]") 
```

After simulating under the null, next we want to find where our observed statistic lands in the null distribution.

**Q - ** Use the code below to plot visualize the p-value on the graph above.

```{r null-viz2, eval = FALSE}
visualize(null_dist) +
  labs(x = "Sample mean", y = "Count", 
       title = "Simulated null distribution of sample means",
       subtitle = "if the true mean rent for one-bedroom apartments was $3000 in Manhattan") +
  shade_pvalue(obs_stat = ____, direction = ____)
```

-   `shade_pvalue()`: plots a p-value region on top of `visualize()`. We plug the observed statistic in `obs_stat`. `direction = "left"`, `direction = "right"`, `direction = "two-sided"` corresponding to the directionality of your alternative hypothesis.

**Q - ** Manually compute the p-value and compare to the `easy-p-value` code chunk below.

Manual p-value:

```{r manual-p-value}

```

Easy p-value:

```{r easy-p-value}

```

**Step 4:** Make a conclusion. 

**Q - **At the significance level 0.05, make your conclusion with the following 3 components: 

- How the p-value compares to the significance level.
- The decision you make with respect to the hypotheses (reject $H_0$ / fail to reject $H_0$).
- The conclusion in the context of the analysis question.


### Practice

(1) Suppose instead you wanted to test the claim that the mean price of rent is not equal to $3000. Which of the following would change? Select all that apply.

a. Null hypothesis
b. Alternative hypothesis
c. Null distribution
d. p-value

(2) Let's test the claim in (1). Conduct the hypothesis test, then state your conclusions in the context of the data at significance levels 0.05 and 0.01.

```{r test-not-equal}

```

## Submitting Application Exercises

- Once you have completed the activity, push your final changes to your GitHub repo. 
- Make sure you committed at least *three* times. 
- Check that your repo is updated on GitHub, and that’s all you need to do to submit application exercises for participation.

