---
title: "AE 16: Hypothesis Testing 2"
subtitle: "due Monday, June 6 at 9:29am"
author: "Bora Jin"
editor_options:
  chunk_output_type: console
output: 
  html_document: 
    toc: true
    toc_float: true
---

## Getting Started 

- Clone the repository entitled "ae16-GitHubUsername" at [course GitHub organization page](https://github.com/sta199-summer22) on your [RStudio](https://vm-manage.oit.duke.edu/containers).

- Open the `.Rmd` file and replace "Your Name" with your name. 

## Inference Overview 

What do we want to do? 

- Estimation $\Rightarrow$ Point estimate and confidence interval 
- Decision $\Rightarrow$ Hypothesis test

Always ask 

- How many variables?
- What types of variables?
- What is the research question? 

## Different Options Inside `generate()`

The function `generate` allows three options for its `type` argument. Discussion of `type = bootstrap`, `type = draw`, and `type = permute` is available [here](https://infer.netlify.app/reference/generate.html).

- `type = permute`: shuffle the data without replacement
  - for hypothesis testing (HT) on a difference in the outcome between groups 
  - example: HT for a difference in proportions of yawners in the treatment and the control group 

- `type = draw`: sample from a theoretical distribution 
  - only for HT on a single proportion
  - example: HT for proportion of the number of heads in coin flips
  
- `type = bootstrap`: re-sample the original data with replacement
  - for confidence intervals (CI) or for HT on a single mean / median
  - example: CI and HT for the true mean rent of one-bedroom apartments in Manhattan

## Push Ups and Pull Ups 

First load the relevant packages:

```{r load-packages, message = FALSE}
library(tidyverse)
library(tidymodels)
```

Today's dataset `push_pull` comes from a "mini study" by [mountain tactical institute](https://mtntactical.com/knowledge/mini-study-grease-the-groove-beats-density-for-push-up-pull-up-improvement/). 

```{r load-data, message = FALSE}
push_pull <- read_csv("data/push_pull.csv")
push_pull %>%
  slice(1:3, 24:26)
```

26 individuals completed 1 of 2 exercise regiments for 3.5 weeks to increase their push ups and pull ups. See the codebook below:

- `participant_id`: unique identifier for each participant
- `age`: age of participant
- `push1` / `push2`: push-ups at beginning and end of program, respectively
- `pull1` / `pull2`: pull-ups at beginning and end of program, respectively
- `training`: which training protocol the individual participated in - either "density" or "gtg" (grease-the-groove)

We create new variables for relative change in push-ups and pull-ups before and after the training. Recall, relative change = (new - old)/old.

```{r rel-change}
push_pull <- push_pull %>%
  mutate(rel_change_push = (push2 - push1)/push1, 
         rel_change_pull = (pull2 - pull1)/pull1)
```

### Part 1: Is the average relative change in pull-ups of a gtg trainee significantly greater than a density trainee?

In other words, we wonder if the group variable `training` affects the average relative increase in pull-ups. Let $\mu_{den}$ and $\mu_{gtg}$ be the true average relative change in pull-ups among density trainees and gtg trainees, respectively.

Let's perform a hypothesis testing. 

**Q - Step 1:** State the null hypothesis and the alternative hypothesis both in words and math.

**Q - Step 2:** Find the relevant statistic from the data. 

```{r statistic, eval = FALSE}
mu_hat_diff <- push_pull %>% 
  group_by(____) %>% 
  summarize(mu_hat = ____) %>% 
  pull(mu_hat) %>% 
  diff()
mu_hat_diff
```

**Step 3:** Simulate from the null distribution and compute the p-value. 

**Q - ** Which of the `type` option in the `generate()` function is the most appropriate in this case? Why?

**Q - ** Complete the code chunk below to simulate 10,000 sample statistics under the null hypothesis. Hint: check the help page of `calculate()` for its `stat` argument. 

```{r part1-null, eval = FALSE}
set.seed(603)

null_dist <- push_pull %>% 
  specify() %>%
  hypothesize() %>% 
  generate() %>%
  calculate()
```

**Q - ** Visualize the p-value region under the null hypothesis with informative labels and compute the p-value. 

```{r part1-pvalue, eval = FALSE}
visualize(null_dist) +
  labs(x = "_____", 
       y = "Count") +
  shade_pvalue(obs_stat = _____, direction = _____) 
  
pvalue1 <- null_dist %>%
  get_pvalue(obs_stat = ______, direction = _____) 
pvalue1
```

**Q - Step 4:** State your conclusion with $\alpha = 0.01$.

### Part 2. Most people who train consistently will see at least a 15% increase in push-ups

**Q - ** State the null hypothesis and the alternative hypothesis.
  
**Q - ** Create a binary outcome `over15` which takes TRUE if `rel_change_push` is larger than 0.15 and FALSE otherwise. 

```{r over15}
```

**Q - ** Find the sample proportion $\hat{p}$. 

```{r phat, eval = FALSE}
p_hat
```

**Q - ** Which of the `type` option in the `generate()` function is the most appropriate in this case? Why?

**Q - ** Perform a hypothesis test, compute the p-value, and state your conclusion with $\alpha = 0.05$.

```{r part2-null, eval = FALSE}
set.seed(603)

null2_dist <- push_pull %>% 
  specify() %>%
  hypothesize() %>% 
  generate() %>%
  calculate()

visualize(null2_dist) +
  labs(x = "Sample proportion of people with a 15% increase in push-ups", 
       y = "Count") +
  shade_pvalue(obs_stat = _______, direction = _______) 
  
pvalue2 <- null2_dist %>%
  get_pvalue(obs_stat = _______, direction = _______) 
pvalue2
```

Now we will construct a confidence interval to evaluate the hypotheses above. 

**Q - ** Which of the `type` option in the `generate()` function is the most appropriate in this case? Why?

**Q - ** Simulate from a bootstrap distribution with `reps = 10000` and visualize the distribution. What is it centered at? 

```{r part2-ci-null}
set.seed(603)

```

**Q - ** We want to construct a confidence interval at a confidence level equivalent to the significance level of $\alpha = 0.05$. What do you think the confidence level should be? Hint: The alternative hypothesis is one-sided. 

**Q - ** Construct a confidence interval with the confidence level equivalent to $\alpha = 0.05$. Interpret the confidence interval. Is the conclusion drawn from the confidence interval consistent with the conclusion from the hypothesis test? 

```{r part2-ci}
```

## Submitting Application Exercises

- Once you have completed the activity, push your final changes to your GitHub repo. 
- Make sure you committed at least *three* times. 
- Check that your repo is updated on GitHub, and that’s all you need to do to submit application exercises for participation.

