---
title: 'Hypothesis testing for independence and summary'
output:
  html_document:
    df_print: paged
date: "03.04.2022"
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

# Bulletin

- Due Friday:

    - homework 04
    - quiz 05
    
- Other:
  
    - Lab hopefully short enough to finish in section (still due Monday)
    - Project report
    
# Today

By the end of today you will

- finish bootstrapping confidence intervals practice from last time (AE15)
- understand the three `generate` options: `bootstrap`, `draw` and `permute`
- practice hypothesis testing for independence

```{r load-packages}
library(tidyverse)
library(tidymodels)
```

```{r load-bass}
bass_mercury = read_csv("data/mercury.csv")
```


# The three generate functions

`permute`

- null: an observation's group membership does not matter
- typical example: no difference in the mean/median between groups
- idea:



# Rent in Manhattan 

On a given day in 2018, twenty one-bedroom apartments were randomly selected on Craigslist Manhattan from apartments listed as "by owner". The data are in the `manhattan` data frame.  We will use this sample to conduct inference on the typical rent of 1 bedroom apartments in Manhattan. 

# Part 1: Drawing a bootstrap sample

Let's start by using bootstrapping to estimate the **mean** rent of one-bedroom apartments in Manhattan. 

## Exercise 1

What is the point estimate of the typical rent? 

## Exercise 2

Let's bootstrap! 

- To bootstrap we will sample with replacement by drawing a value from the bowl. 
- How many draws do we need for our bootstrap sample? 

Fill in  the values from the bootstrap sample conducted in class. Once the values are filled in, uncomment the code. 

```{r boot-sample}
# class_bootstrap <- c()

```

## Exercise 3

- About what value do you expect the bootstrap statistic to take? 
- Calculate the statistic from the bootstrap sample. 

```{r calc-boot-stat}
# add code

```

# Part 2: Bootstrap confidence interval

**We will use the `infer` package, included as part of `tidymodels` to calculate a 95% confidence interval for the mean rent of one-bedroom apartments in Manhattan.**

We start by setting a seed to ensure our analysis is reproducible. 

## Generating the bootstrap distribution

We can use R to take many bootstrap samples and generate a bootstrap distribution

Uncomment the lines and fill in the blanks to create the bootstrap distribution of sample means and save the results in the data frame `boot_dist`.    

Use **100 reps** for the in-class activity. (You will use about 15,000 reps for assignments outsdie of class.)

```{r create-bootstrap-dist}
set.seed(2252022)

boot_dist <- manhattan #%>%
  #specify(______) %>%
  #generate(______) %>%
  #calculate(______)
```

- How many rows are in `boot_dist`? 
- What does each row represent? 
- What are the variables in `boot_dist`? What do they mean?

## Visualize the bootstrap distribution 

Visualize the bootstrap distribution using a histogram. Describe the shape, center, and spread of this distribution. 

```{r viz-bootstrap}
# add code

```

## Calculate the confidence interval

Uncomment the lines and fill in the blanks to construct the 95% bootstrap confidence interval for the mean rent of one-bedroom apartments in Manhattan. 

```{r confidence-interval-95}
#___ %>%
#  summarize(lower = quantile(______),
  #          upper = quantile(______))
```

## Interpret the interval

Write the interpretation for the interval calculated above. 

# Part 3: Changing the confidence level 

- Modify the code used to calculate a 95% confidence interval to calculate a **90% confidence interval** for the mean rent of one-bedroom apartments in Manhattan. How does the width of this interval compare to the width of the 95% confidence interval? 

```{r confidence-interval-90}
#calculate a 90% confidence interval
```

- Now let's calculate a 99% confidence interval for the mean rent of one-bedroom apartments in Manhattan. How does the width of this interval compare to the width of the 95% confidence interval? 

```{r confidence-interval-99}
#calculate a 99% confidence interval
```

- Question: Does a confidence interval have to be symmetric?

- What is one advantage to using a 90% confidence interval instead of a 95% confidence interval to estimate a parameter? - What is one advantage to using a 99% confidence interval instead of a 95% confidence interval to estimate a parameter? 

# Part 4: Additional practice

Next, use bootstrapping to estimate the median rent for one-bedroom apartments in Manhattan. 
- Generate the bootstrap distribution of the sample medians. Use 100 reps. Save the results in `boot_dist_median`.

```{r boot-dist-median}
set.seed(100)
## add code
```

- Calculate a 92% confidence interval. 

```{r confidence-interval-92}
## add code
```

- Interpret the 92% confidence interval. 