---
title: 'Review: using generate with tidymodels'
output:
  html_document:
    df_print: paged
date: "03.02.2022"
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

# Bulletin

- Due Friday:

    - homework 04
    - quiz 05
    
- Other:
  
    - Lab hopefully short enough to finish in section
    
# Today

By the end of today you will

- finish bootstrapping confidence intervals practice from last time (AE15)
- practice the three `generate` options: `permute`, `draw`, and `bootstrap`

```{r load-packages}
library(tidyverse)
library(tidymodels)
```


# Data

```{r load-bass}
push_pull = read_csv("data/push_pull.csv")
```

```{r peak-at-data}
push_pull %>%
  slice(1:3, 24:26)
```


The `push_pull` dataset comes from a "mini study" by [mountain tactical institute](https://mtntactical.com/knowledge/mini-study-grease-the-groove-beats-density-for-push-up-pull-up-improvement/). 

26 individuals completed 1 of 2 exercise regiments for 3.5 weeks to increase their pushups and pullups. Codebook below:

- `participant_id`: unique identifier for each participant
- `age`: age of participant
- `push1/2`: push-ups at beginning and end of program respectively
- `pull1/2`: pull-ups at beginning and end of program respectively
- `training`: which training protocol the individual participated in

```{r pct-increase-column, results = 'hide'}
push_pull = push_pull %>%
  mutate(
    pct_push_inc = (push2 / push1 ) - 1,
    pct_pull_inc = (pull2 / pull1) - 1)
```


# The three generate functions

## 1. `permute`

- description: permutes variables
- typical case: is there a difference in the outcome between groups?
- associated null: group membership does not matter i.e. group A and group B have the same outcome
- Example (test for independence):

Two exercise regimes:

- "density" training
- "grease-the-groove" (gtg)

We want to know, is the average pull-up percent increase of a gtg trainee significantly greater than a density trainee?

Fundamentally, does the categorical variable `training` affect the average percentage increase in pull-ups?

**State the null hypothesis**

What we want to do to simulate data under this null:

```{r conceptual-permute, results = 'hide'}
random_training = sample(push_pull$training, replace = FALSE)

push_pull %>%
  select(pct_pull_inc) %>%
  mutate(random_training = random_training)
```



### Exercise 1:

- Complete the hypothesis specification above by stating the alternative.
- Perform a hypothesis test, compute the p-value and state your conclusion with $\alpha = 0.05$.


```{r ex-1}
set.seed(1)
# code here
```

--- 

## 2. `draw`

- description: flip a coin with probability p, or roll a die with probabilities associated with each side. [See here](https://cran.r-project.org/web/packages/infer/vignettes/observed_stat_examples.html#one-categorical-2-level---gof) for reference to the multinomial setting.
- typical case: test the proportion of a binary outcome
- null: proportion $p$ is some fixed number
- Example (hypothesis test for a single proportion):

"Most people who train consistently will see at least a 15% increase in push-ups over an 3.5 week training period."

Breaking it down:

- "Most" i.e. "greater than 50%". This indicates we should examine a proportion.

**What's the null?**

- "will see at least a 15% increase". Each person either increases by 15% over 3.5 weeks or does not. This is our binary outcome.

```{r binary-outcome, results = 'hide'}
push_up = push_pull %>%
  select(pct_push_inc) %>%
  mutate(over15pct =
           pct_push_inc > 0.15)

push_up
```

- Our "NULL" means that the TRUE and FALSE in the `over15pct` column are equally likely.

### Exercise 2

- Complete the hypothesis specification above by stating the alternative.
- Perform a hypothesis test, compute the p-value and state your conclusion with $\alpha = 0.01$.

```{r ex-2}
set.seed(2)
# code here
```

---

## 3. `bootstrap`

- description: re-sample your data with replacement
- typical case (in hypothesis testing): does the mean equal a specific value? Does the median equal a specific value?
- null: what would the data have looked like if nothing but the point estimate changed?
- Example:

"The mean age of push-up/pull-up training partcipants is greater than 30".

**What's the null?**

Bootstrapping does the following...

```{r test-mean, results = 'hide'}

# find observed statistic
obs_mean_age = push_pull %>%
  drop_na(age) %>%
  summarize(meanAge = mean(age)) %>%
  pull()

# subtract observed_mean - desired_mean from age
age_and_null = push_pull %>% 
  select(age) %>%
  drop_na(age) %>%
  mutate(nullAge = age - (obs_mean_age - 30))

# show data frame
age_and_null

# show the means of each column
age_and_null %>%
  summarize(meanAge = mean(age),
            mean_nullAge = mean(nullAge))
```
If we take bootstrap samples from this new `nullAge` column, we are sampling from data with the **same variability** as our original data, but a different mean. This is a nice way to explore the null!

### Exercise 3

- Complete the hypothesis specification above by stating the alternative.
- Perform a hypothesis test, compute the p-value and state your conclusion with $\alpha = 0.05$.


```{r ex-3}
set.seed(3)
# code here
```

