---
title: "AE 17: Central Limit Theorem"
subtitle: "due Tuesday, June 7 at 9:29am"
author: "Bora Jin"
editor_options:
  chunk_output_type: console
output: 
  html_document: 
    toc: true
    toc_float: true
---

## Getting Started 

- Clone the repository entitled "ae17-GitHubUsername" at [course GitHub organization page](https://github.com/sta199-summer22) on your [RStudio](https://vm-manage.oit.duke.edu/containers).

- Open the `.Rmd` file and replace "Your Name" with your name. 

## Bone Density for Women Over 65

```{r load-package, message=F}
library(tidyverse)
theme_set(theme_bw())
```

Suppose the bone density for 65-year-old women is normally 
distributed with mean $809 ~mg/cm^3$ and standard deviation of $140 ~mg/cm^3$.

Let $X$ be the bone density of 65-year-old women. We can write this distribution of $X$ in mathematical notation as 

$$X \sim N(809, 140)$$
### Part 1: Original density 

The original density is given below: 

```{r visualize-normal}
ggplot(NULL, aes(c(809 - 140*3, 809 + 140*3))) +
  geom_area(stat = "function", fun = dnorm, 
            args = list(mean = 809, sd = 140), color = "black", fill = NA) +
  labs(title = "Population distribution", x = "", y = "Density") 
```

**Q - ** Before typing any code, based on what you know about the normal distribution, what do you expect the median bone density to be?

The median is the same as the mean, 809. 

**Q - ** What bone densities correspond to Q1 (25th percentile), Q2 (50th percentile), and Q3 (the 75th percentile) of this distribution? Use the `qnorm()` function to calculate these values. 

```{r}
(Q1 <- qnorm(0.25, mean = 809, sd = 140))
(Q2 <- qnorm(0.5, mean = 809, sd = 140))
(Q3 <- qnorm(0.75, mean = 809, sd = 140))
```

## Exercise 2

The densities of three woods are below:

- Plywood: $540 ~mg/cm^3$
- Pine: $600 ~mg/cm^3$
- Mahogany: $710 ~mg/cm^3$

- What is the probability that a randomly selected 65-year-old woman has bones less dense than Pine?

```{r}

```

- Would you be surprised if a randomly selected 65-year-old  woman had bone density less than Mahogany? What if she had bone density less than Plywood? Use the respective probabilities to support your response. 

## Exercise 3

Suppose you want to analyze the mean bone density for a group of 10 randomly selected 65-year-old women. 

- Are the conditions met use the Central Limit Theorem to define the distribution of $\bar{x}$, the mean density of 10 randomly selected 65-year-old women? 
  - Independence? 
  - Sample size/distribution? 
  
- What is the shape, center, and spread of the distribution of $\bar{x}$, the mean bone density for a group of 10 randomly selected 65-year-old women?

- Write the distribution of $\bar{x}$ using mathematical notation. 

## Exercise 4

- What is the probability that the mean bone density for the group of 10 randomly-selected 65-year-old women is less dense than Pine?

- Would you be surprised if a group of 10 randomly-selected 65-year old women had a mean bone density less than Mahogany? What the group had a mean bone density less than Plywood? Use the respective probabilities to support your response. 

## Exercise 5

Explain how your answers differ in Exercises 2 and 4.


## Submitting Application Exercises

- Once you have completed the activity, push your final changes to your GitHub repo. 
- Make sure you committed at least *three* times. 
- Check that your repo is updated on GitHub, and that’s all you need to do to submit application exercises for participation.

