---
title: "Regression 4: Review"
date: "March 30 2022"
output:
  html_document:
    df_print: paged
---

```{r, warning = FALSE, message  = FALSE}
library(tidyverse)
library(tidymodels)
```

# Bulletin

- Friday
  - peer reviews due
  - exam 02 released

# Today

Today we review and practice linear modeling concepts with the Palmer penguins data.

```{r load-penguins, warning = F, message = F}
data(penguins)
```

Use `?penguins` or [click here](https://github.com/allisonhorst/palmerpenguins) for more info about the dataset.

# Main vs interaction effects

$y$: body mass (g)

$x_1$: bill length

$x_2$: island Dream

$x_3$: island Torgersen

$$
\hat{y} = \hat{\beta_0} + \hat{\beta_1} x_1 + \hat{\beta_2} x_2 + \hat{\beta_3} x_3
$$

Bill length only impacts body mass via the term $\beta_1 x_1$. $x_2$ and $x_3$ can be thought of as turning on or off a constant.

## Main

```{r main, echo = F, message = F, warning = F}
main_fit = linear_reg() %>%
  set_engine("lm") %>%
  fit(body_mass_g ~ bill_length_mm + island, data = penguins)
  main_fit %>%
  tidy()

b0 = 1225.79107
bD = -919.07448
bT = -523.29201
bBill = 77.11933

# show_col(hue_pal()(3))  

penguins %>%
ggplot(aes(x = bill_length_mm, y = body_mass_g, color = island)) +
  geom_point() +
  theme_bw() +
  geom_abline(intercept = b0, slope = 77.11933, color = "#F8766D", lwd = 1) +
  geom_abline(intercept = b0 + bD, slope = 77.11933, color = "#00BA38", lwd = 1) +
  geom_abline(intercept = b0 + bT, slope = 77.11933, color = "#619CFF", lwd = 1) +
  labs(x = "Bill length (mm)",
       y = "Body mass (g)",
       title = "Island main effect")
```


## Interaction

$$
\hat{y} = \hat{\beta_0} + \hat{\beta_1} x_1 + \hat{\beta_2} x_2 + \hat{\beta_3} x_3 + \hat{\beta_4} x_1 x_2 + \hat{\beta_5} x_1 x_3
$$

What's different? The interaction terms ($\beta_4 x_1 x_2$ and $\beta_5 x_1 x_3$). 

```{r interaction, echo = F, warning = F, message = F}

interaction_fit = linear_reg() %>%
  set_engine("lm") %>%
  fit(body_mass_g ~ bill_length_mm * island, data = penguins)

  interaction_fit %>%
  tidy()

penguins %>%
ggplot(aes(x = bill_length_mm, y = body_mass_g, color = island)) +
geom_point() +
theme_bw() +
geom_smooth(method = 'lm', se = F) +
labs(x = "Bill length (mm)",
     y = "Body mass (g)",
     title = "Island interaction effect")
```

Interpreting interactions can be difficult, especially without writing things down. To make it easier, we will compare the implied linear models:

Plug in 0 for `islandDream` ($x_2$) and 0 for `islandTorgersen` ($x_3$) to get the linear model for `islandBiscoe` penguins

Plug in 1 for `islandDream` ($x_2$) and 0 for `islandTorgersen` ($x_3$) to get the linear model for `islandDream` penguins

Plug in 0 for `islandDream` ($x_2$) and 1 for `islandTorgersen` ($x_3$) to get the linear model for `islandTorgersen` penguins

- Biscoe fitted model:

$$
\hat{y} = -1726.0+ 142.3 x_1
$$

- Dream fitted model:

$$
\hat{y} = -1726.0 + 142.3 x_1 + 4478.7 -120.6 x_1
$$

Combine terms:

$$
\hat{y} = 2752.7 + 21.7 x_1 
$$

## Exercise 1 

Write out the fitted model for Torgersen island below.

- Torgersen model:
$$
\hat{y} = [\text{write here}]
$$

# Interpreting

Now we can interpret the interaction model by comparing bill length slopes between islands.

- For a unit increase in bill length of a penguin from the island Dream, how much do we expect the body mass to increase?

## Exercise 2

- You measured the bill length of a penguin from island Biscoe and a penguin from island Torgersen a year ago. You re-measure them today and find the bill length of each one grew by exactly 2 mm. How much mass do you expect each penguin to have gained?

## Exercise 3

Are the intercepts meaningful?

# Practice

## Exercise 4

Is the relationship between Body mass (g) and Bill depth (mm) positive or negative? Create a convincing argument from the data.

## Exercise 5

Let's build a linear model of body mass using bill depth and one other predictor of your choosing (hint: see previous exercise!)

- Write out a linear model with both predictors and fit the model.

- Fit the linear model

- Are any of the predictors insignificant at the $\alpha = 0.05$ level? If so, state the null and alternative hypothesis, p-value and conclusion in context.

- Do you prefer this model to the interaction effects model from above? Why?



