---
title: 'AE 24: Final Project in R Markdown - Report'
author: "Bora Jin"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document: default
---

## Bulletin 

Course evaluation is now open! It will be closed on Tuesday, June 21 at 11:59pm. 
If **more than 80%** of you complete the evaluation, everyone will receive **a bonus point**!  

Check your emails or visit [duke.evaluationkit.com](https://duke.evaluationkit.com/) to complete the evaluation. 

## Getting Started 

- Clone the repository entitled "ae24-GitHubUsername" at [course GitHub organization page](https://github.com/sta199-summer22) on your [RStudio](https://vm-manage.oit.duke.edu/containers).

- Open the `.Rmd` file. 

## Introduction

Refer to [STA199: Final Project - Tips](https://sta199-summer22.netlify.app/project/project-tips_bj) for general guidelines for formatting of your report. 

## Text 

In order to make bullet points, skip a line and use a hyphen `-`. 

- *italic*
- **bold**
- `function` or `variable`
- [links](https://rmarkdown.rstudio.com/lesson-8.html)
  - `tab` once for a sub-list 
  - **Q - **Knit the document to PDF and see how the link appears. 
  - If you know your PDF will be printed, add `links-as-notes: TRUE` to YAML to 
  display URLs as a footnote in the knitted PDF. 
- Latex equations 
  - inline: $y = \beta_0 + \beta_1x_1 + \epsilon$ 
  - chunk: $$\hat{y} = \hat{\beta}_0 + \hat{\beta}_1x_1$$
- The default font size is 10pt. You can increase it to 11pt or 12pt. 
In order to change the font size to 11pt, add `fontsize: 11pt` to YAML.

## Citation 

Your report will include citations about the data source, previous research, and other sources as needed. At a minimum, you should have a citation for the data source. 

Follow the steps below to easily include citations in your report:

1. Prepare a `.bib` file. All of your bibliography entries will be stored in it. 
Let's take a look at `references.bib`. 

**Q -** Add a bibliography entry for the book *R markdown: The definitive guide* to `references.bib`. 

- Search the title in Google scholar. 
- Click "Cite". 
- Click "BibTeX".
- Copy all and paste it in `references.bib`. 

2. Include `bibliography: references.bib` in the YAML. 

3. At the end of the report, include `## References`. This will list all of the references at the end of the document under the section "References". 

4. If you want to include an Appendix after References, include the additional code shown at the end of this document. 

Here are examples of different in-text citations: 

- In @gorman2014structural, the authors focus on the analysis of Adelie penguins.
- Studies have shown whether environmental variability in the form of winter sea ice is associated with differences in male and female pre-breeding foraging niche [@gorman2014structural].

**Q - ** Add a citation for *R markdown: The definitive guide* in the sentence below.

The following code chunk is adapted from ______.

## Options for Code Chunks 

Code chunk options are used to customize how the code and output is displayed in the knitted R Markdown document. There are two ways to set code chunk options: 

- In the header of an individual code chunk 
- As a global setting to apply to all code chunks
  
A few options to change what we show / hide in the knitted document: 

- `message = FALSE` to hide messages.
- `warning = FALSE` to hide warnings (with caution!) 
  - While you conduct analyses, you should know what warning messages you receive and decide if it is okay to ignore them. For the final knitting, you may use `warning = FALSE`. 
- `echo = FALSE` to hide code. 
  - For the project, you will set the option `echo = FALSE` to hide all code in your final report. 
- `include = FALSE` to run code but hide code, output, messages, warnings, etc.
  - Avoid using this option as a global setting.
- `eval = FALSE` to not run code. 
  - Useful when your code chunk is incomplete. You should not have any incomplete code chunks in your final report! 

**Q - ** Clean up the mess below. 

```{r setup, include = TRUE}
knitr::opts_chunk$set(message = TRUE, 
                      warning = TRUE, 
                      echo = TRUE,
                      fig.width = 6, # width of figure
                      fig.asp = .618, # aspect ratio of figure
                      out.width = "70%", # width relative to text
                      fig.align = "center" # figure alignment
                      )
```

```{r load-packages}
library(knitr)
library(tidyverse)
theme_set(theme_bw()) # set global theme 
```

**Q - ** Try `echo = FALSE` and take a look at the updated PDF. 

## Theme for Plotting

For `ggplots`, you can set a global theme (see the code chunk above). Click [here](https://ggplot2.tidyverse.org/reference/ggtheme.html) to explore available themes and choose your favorite. In case your PDF is printed, gray or dark background for plots are not recommended. 

You can also set a color theme of your choosing. You can call any color by any name you would like in `my_palette`. 

**Q - ** Add "dukeblue" in `my_palette`. Hint: Find hex code. 

```{r color-options, include = FALSE}
my_palette <- list(gray = "#999999", 
                   yelorange = "#E69F00", 
                   blue = "#56B4E9", 
                   green = "#009E73", 
                   yellow = "#F0E442", 
                   darkblue = "#0072B2", 
                   orangered = "#D55E00",
                   deeppink = "#CC79A7"
                   )
```

## Customizing Plots 

```{r basic-plot}
mpg %>% 
  ggplot(aes(y = manufacturer, fill = drv)) + 
  geom_bar()
```

We can do better than this. 

**Q - ** Incorporate the following recommended changes in the plot. 

- Axes labels begin with a capital letter.
- The title of the fill legend is "The type of drive train". 
- The levels of the fill legend are "front-wheel" for "f", 
"rear-wheel" for "r", and "four-wheel" for "4" in written order.
- Use fill colors from `my_palette`.
- Manufacturer names begin with a capital letter. 
- Shorter bars locate on top. 
- Add a caption "Counts of the type of drive train by manufacturer".

```{r customize-plot}

```


## Customizing Tables

Calculate mean, median, and standard deviation of `cty`. 

```{r summary-stats}
mpg %>% 
  summarize(mean_cty = mean(cty), median_cty = median(cty), sd_cty = sd(cty))
```

We can also do better than this. 

**Q - ** Incorporate the following recommended changes in the table. 

  - Print it as a nice table. 
  - Results are rounded to 3 decimal places. 
  - Column names are "mean mpg", "median mpg", "sd of mpg".
  - Add a caption "Summary statistics of city miles per gallon (mpg) in city". 
  
```{r format-table}

```

## Acknowledgements 

These notes were adapted from the following: 

- [Markdown Basics](https://rmarkdown.rstudio.com/lesson-8.html)
- [Communicating data analysis](https://sta-212-f19.netlify.app/slides/lec-slides/15-communicating-data-analysis.html#1) by Dr. Lucy McGowan
- [Customizing R Markdown](https://rladiesstl.github.io/intro-meetup4-oct-2021/customize-rmd.html#1) by Dr. Mine Çetinkaya-Rundel
- [Take a sad plot and make it better](https://alison.netlify.app/rlm-sad-plot-better/#1) by Dr. Alison Hill

<!-- <div id="refs"></div> -->

<!-- ## Appendix -->


