---
title: "Homework #03: Probability and Simulation-based Inference"
subtitle: "due Wednesday, June 8 at 11:59pm"
output: 
  tufte::tufte_html:
    css: "sta199-hw.css"
    tufte_variant: "envisioned"
    highlight: pygments
    toc: true
    toc_depth: 2
link-citations: yes
---

## Goals

-   Gain proficiency calculating marginal, joint, and conditional probabilities
-   Use visualizations and probabilities to analyze multivariate associations between categorical variables
-   Construct confidence intervals
-   Conduct hypothesis tests
-   Interpret confidence intervals and results of hypothesis tests in context of the data 

## Getting started

- Go to [course GitHub organization page](https://github.com/sta199-summer22) and find the repository entitled "hw03-GitHubUsername"
- Clone the repo and make a new project in RStudio. 
- Find `hw03.Rmd` to open the template R Markdown file. 

## Instructions

For each exercise:

- Make sure we see all relevant code and output in the knitted PDF. 
If you use inline code, make sure we can still see the code used to derive that answer.

- Write all narrative in complete sentences with **decimal numbers rounded to three decimal places.** 

- Use a small number of `reps` (about 500) as you write and test out your code. 
Once you have finalized all of your code, increase the number of `reps` to 15,000 
to produce your final results.
    
- Write your code and narrative in a reproducible way, so you can easily change the number of reps. 
For example, consider ways you can write your narrative using inline code, so the 
relevant values update when you change the number of reps.

- For each simulation exercise, use the seed specified in the exercise instructions.

## Formatting 

- All plots should follow best visualization practices; plots should include:

  - an informative title, 
  - axes that are labeled, and 
  - careful consideration should be given to aesthetic choices.
  - Please choose colorblind-friendly color maps such as `viridis`, `scico`, and many others.

- Place all plots in the center and properly adjust their size so that they are placed nicely in a written report. 

- Don’t forget to label your R chunk as well. Your label should be short, informative, shouldn’t include spaces, and shouldn’t repeat a previous label.

## Packages 

We’ll use the `tidyverse` package for much of the data wrangling and visualization, 
the `tidymodels` package for inference, and the data lives in the `openintro` package.

```{r load-package, message=FALSE, warning=FALSE}
library(tidyverse)
theme_set(theme_bw())
library(tidymodels)
library(openintro)
```

## United States Births 

Every year, the US releases a large dataset containing information on births 
recorded in the country. This dataset is useful to researchers 
studying the relation between habits and practices of expectant mothers and 
the birth of their children. We will work with a random sample of 1,000 observations
from the dataset released in 2014. 

The subsetted data can be found in the `openintro` package, and it's called `births14`.
Each observation represents a birth in the US. You can find out more about the 
dataset by running `?births14` in the Console.

## Probability 

1. The table below tells us that each birth is classified as full-term if 
the gestational age is larger than or equal to 37 weeks. What is the probability 
a randomly selected baby in the US is premature? 

```{r echo = FALSE}
births14 %>% 
  group_by(premie) %>% 
  summarize(min_week = min(weeks), 
            max_week = max(weeks)) %>% 
  knitr::kable()
```


2. Let $A$ be the event that a baby is female and $B$ be the event that a baby 
weighs more than 9.5 pounds. Determine if the two events are disjoint or not. 
Also determine if they are independent. Explain your reasoning. 


3. What is the probability that a baby is premature given the baby is female? 
What about the probability that a baby is premature given the baby is male? 
Create a horizontal stacked bar plot of `sex` with relative frequencies of `premie`. 
Have the sex of a baby on the y-axis and fill the bars according to whether 
the baby was premature or not. 

4. Using the results in exercises above and Bayes' theorem, compute the probability 
that a baby is female given the baby is premature.
Provided that the event $A$ is a baby is female and the event $B$ is a baby is premature, 
is $A$ independent of $B$? Why or why not?

::: {.commit}
`r emo::ji("yarn")` `r emo::ji("white_check_mark")` `r emo::ji("arrow_up")` Knit, commit, and push your final changes to GitHub with a meaningful commit message.
:::

## Simulation-based Inference 

### Baby weights

According to this [article](https://www.medicalnewstoday.com/articles/325630#by-age), 
the World Health Organization (WHO)-released average birth weight of a full-term 
female baby is 7.125 pounds (lbs). 

We want to evaluate whether the average weight of full-term female babies in the US is 
significantly different than 7.125 lbs. 

5. Conduct a hypothesis test following the steps described below. 

- (a). State the null hypothesis and the alternative hypothesis in math. Clearly define 
  all parameters you introduce. 

- (b). Create a filtered data frame called `births_girl` that contain data only for 
  full-term female babies. Then, calculate the mean of the weights of these babies.

- (c). Simulate data under the null hypothesis, visualize it with the p-value 
   region shaded, and calculate the p-value. You may start with the code chunk provided 
   below. Set seed number at 5. Use `rep = 15000` for your final turn in. Remove 
   `eval = FALSE` once you fill in the blanks. 

- (d). Make a conclusion at $\alpha = 0.05$ and interpret the results in 
   context of the data.
   
- (e). Construct a confidence interval at the equivalent level to the hypothesis test above
   and interpret the interval in context of the data. Use the same seed as Ex 5-c.

::: {.commit}
`r emo::ji("yarn")` `r emo::ji("white_check_mark")` `r emo::ji("arrow_up")` Knit, commit, and push your final changes to GitHub with a meaningful commit message.
:::

### Baby weight vs. smoking

Consider the possible relationship between a mother's smoking habit and the weight of her baby. 
Plotting the data is a useful first step because it helps us quickly visualize trends, 
identify strong associations, and develop research questions.

6.  Make side-by-side boxplots displaying the relationship between `habit` and `weight`.
    What does the plot highlight about the relationship between these two variables?

7.  Before moving forward, save a version of the dataset omitting observations where there are NAs for `habit`.
    You can call this version `births_habitgiven`.

8. Let's conduct the appropriate hypothesis test. 

- (a). State the null and the alternative hypothesis in math with a clear definition 
  of any parameter you introduce. 

- (b). Calculate the observed difference between the average weights of babies 
  born to smoking and non-smoking mothers. 

- (c). Simulate data under the null hypothesis, visualize it with the p-value 
  region shaded, and calculate the p-value. Set seed number at 8. 
  Use `rep = 15000` for your final turn in.
  
- (d). State your conclusion in context of the research question with $\alpha = 0.01$.

- (e). Construct a confidence interval, at the equivalent level to the hypothesis test, 
  for the difference between the average weights of babies born to smoking and non-smoking mothers, 
  and interpret this interval in context of the data. Again, use the seed number 8. 

::: {.commit}
`r emo::ji("yarn")` `r emo::ji("white_check_mark")` `r emo::ji("arrow_up")` Knit, commit, and push your final changes to GitHub with a meaningful commit message.
:::

## Submission

Knit to PDF to create a PDF document. Stage and commit all remaining
changes, and push your work to GitHub. Make sure all files are updated
on your GitHub repo.

Only upload your PDF document to Gradescope. Before you submit the
uploaded document, mark where each answer is to the exercises. If any
answer spans multiple pages, then mark all pages. Associate the
*"Workflow & formatting"* section with the first page.

## Grading (60 pts)

<br>

| Component | Points |
|:----------|:-------|
| Ex 1      | 1      |
| Ex 2      | 4      |
| Ex 3      | 4      |
| Ex 4      | 5      |
| Ex 5      | 18     |
| Ex 6      | 2      |
| Ex 7      | 1.5    |
| Ex 8      | 18.5   |
| Workflow & formatting   | 6      |

**Grading notes**:

- The "Workflow & formatting" grade is to assess the reproducible workflow. This includes updating the name on the assignment to your name, having at least 3 informative commit messages, labeling the code chunks, using tidyverse style throughout, and following best visualization practices. 

