---
title: "Data Visualization 2"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Material 

`r emo::ji("movie_camera")` Watch [Visualizing Numerical Data](https://youtu.be/waBabVTI8ec) 

- [Slides](https://rstudio-education.github.io/datascience-box/course-materials/slides/u2-d03-viz-num/u2-d03-viz-num.html#1)

`r emo::ji("movie_camera")` Watch [Visualizing Categorical Data](https://youtu.be/21h3rEO8k2E) 

- [Slides](https://rstudio-education.github.io/datascience-box/course-materials/slides/u2-d04-viz-cat/u2-d04-viz-cat.html#1)

---

## Today's Goal 

- Explain continuous, discrete, and categorical variables
- Understand how to make visualizations and summarize variables according to their type
- Develop a faceted plot

---

## Quiz 

**Q - ________ variables can be classified as either continuous or discrete.**

--

Numerical

--

**Q - ________ categorical variable has a natural ordering.**

--

Ordinal 

--

**Q - Classify the following variables:**

- monthly expenses
--
: numeric, continuous
--

- number of shoes
--
: numeric, discrete 
--

- course satisfaction rating (“extremely dislike”, “dislike”, “neutral”, “like”, “extremely like”)
--
: categorical, ordinal 
--

- eye color
--
: categorical, nominal
---

## Quiz 

**Q - Describe the shape of the following distributions of a numeric.**

```{r left-skewed, echo = FALSE, fig.align='center', fig.height=3}
set.seed(123)
x <- seq(-20, 0, length = 10000)
y <- dgamma(-x, shape = 2, scale = 1)
ggplot(data = data.frame(x, y)) +
  geom_line(aes(x,y)) + labs(y="density")
```

---

## Quiz 

**Q - Describe the shape of the following distributions of a numeric.**

```{r uniform, echo = FALSE, fig.align='center', fig.height=3}
set.seed(123)
x <- seq(-0.1, 1.1, length = 10000)
y <- dunif(x)
ggplot(data = data.frame(x, y)) +
  geom_line(aes(x,y)) + labs(y="density")
```

---

## Quiz 

**Q - Describe the shape of the following distributions of a numeric.**

```{r bimodal, echo = FALSE, fig.align='center', fig.height=3}
set.seed(123)
x <- seq(-5, 10, length = 10000)
y <- dnorm(x, mean = -2, sd = 1) + dnorm(x, mean = 7, sd = 2)
ggplot(data = data.frame(x, y)) +
  geom_line(aes(x,y)) + labs(y="density")
```

---

## Quiz 

**Q - Fill in the blanks with appropriate** `R` **functions**

- center: mean (`___`), median (`___`)
- spread: range (`range`), standard deviation (`___`), interquartile range (`IQR`)

---

## Quiz 

**Q - Fill in the blanks with appropriate** `R` **functions**

- center: mean (`mean`), median (`meduan`)
- spread: range (`range`), standard deviation (`sd`), interquartile range (`IQR`)

--

**Q - What plot might you draw if you want to detect potential outliers?**

--

Box plot 

---

## Quiz 

**Q - Which of these commands are inappropriate to visualize distribution of a single numerical variable?**

a. `geom_histogram()`

b. `geom_point()`

c. `geom_density()`

d. `geom_boxplot()` 

e. `geom_hex()`

---

## Quiz 

**Q - Which of these commands are inappropriate to visualize distribution of a single numerical variable?**

a. `geom_histogram()`

**b. `geom_point()` - to visualize relationships between two numerical variables** 

c. `geom_density()`

d. `geom_boxplot()` 

**e. `geom_hex()` - relationships between two numerical variables through binning** 

---

## Quiz 

**Q - Which of these commands are inappropriate to visualize relationships between numerical and categorical variables?**

a. `geom_boxplot()`

b. `geom_violin()`

c. `geom_density_ridges()` 

d. `geom_bar()` 

---

## Quiz 

**Q - Which of these commands are inappropriate to visualize relationships between numerical and categorical variables?**

a. `geom_boxplot()`

b. `geom_violin()`

c. `geom_density_ridges()` 

**d. `geom_bar()` - visualize distribution of a categorical variable or relationship between categorical variables**

---

## Quiz 

**Q - Which of these is the most relevant for the difference between two bar plots?**

```{r, echo = FALSE}
library(openintro)
loans <- loans_full_schema %>%
  select(loan_amount, interest_rate, term, grade, 
         state, annual_income, homeownership, debt_to_income)
```

.pull-left[
```{r, echo = FALSE, out.width = '100%'}
plot1 <- ggplot(loans, aes(x = homeownership, 
                  fill = grade)) +
  geom_bar() + 
  labs(
    x = "Proportion",
    y = "Homeownership",
    fill = "Grade"
  ) +
  scale_fill_viridis_d()
plot2 <- ggplot(loans, aes(x = homeownership,
                  fill = grade)) +
  geom_bar(position = "fill") +
  labs(
    x = "Proportion",
    y = "Homeownership",
    fill = "Grade"
  ) + 
  scale_fill_viridis_d()
ggpubr::ggarrange(plot1, plot2, ncol = 2, common.legend = T, legend = "bottom")
```
]

.pull-right[
a. `aes(x = homeownership, fill = grade)`

b. `position = "fill"` 

c. `labs()`
]

---

## Quiz 

**Q - Which of these is the most relevant for the difference between two bar plots?**

.pull-left[
```{r, echo = FALSE, out.width = '100%'}
plot1 <- ggplot(loans, aes(x = homeownership, 
                  fill = grade)) +
  geom_bar() + 
  labs(
    x = "Proportion",
    y = "Homeownership",
    fill = "Grade"
  ) +
  scale_fill_viridis_d()
plot2 <- ggplot(loans, aes(x = homeownership,
                  fill = grade)) +
  geom_bar(position = "fill") +
  labs(
    x = "Proportion",
    y = "Homeownership",
    fill = "Grade"
  ) + 
  scale_fill_viridis_d()
ggpubr::ggarrange(plot1, plot2, ncol = 2, common.legend = T, legend = "bottom")
```
]

.pull-right[
a. `aes(x = homeownership, fill = grade)`

**b. `position = "fill"` - relative frequency within x**

c. `labs()`
]


---

class: middle, center

# Questions?

---

## Let's Practice Together! 

Go to [AE 04](https://sta199-summer22.netlify.app/appex/ae04_BJ.html)

---

## For Monday 

- Watch videos for [Prepare: May 17](https://sta199-summer22.netlify.app/prepare/week02_may17_BJ.html)

- Submit your `AE04`

