---
title: "Spatial Data and Visualization"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Material 

`r emo::ji("movie_camera")` Watch [Spatial Data](https://warpwire.duke.edu/w/Yw4GAA/)

- [Slides](https://sta199-summer22.netlify.app/slides/spatial.html#1)

---

## Today's Goal 

- Understand spatial data frame structure
- Be able to create a visualization from a spatial data frame

---

## Spatial Data are Interesting and Important! 

- Straightforward visualization through maps
- Help detecting spatial trends 
- Many real examples have spatial patterns 
    - air pollutants 
    - water pollutants 
    - disease spread 
    - home sales prices
    - $\cdots$
- Analyses of such data should account for their spatial relationships 

---

## Quiz

**Q - Which package are we going to use in this class to work with spatial data?**

--

`library(sf)`

- represents "simple features" as records in a `data.frame` or `tibble` with a **geometry** column
  - simple features can contain both spatial and non-spatial data
- compatible with `dplyr` functions we've learnt
- functions begin with `st_`
--

- how to draw an `sf` object in ggplot?

--
`geom_sf()`

---

## Quiz

**Q - What do you need to check before visualizing and wrangling multiple `sf` objects?**

--

Coordinate reference system (CRS) - they should match

---

class: middle, center

# Questions?

---

## Hot Keys

<br> 

| **Task / function**  | **Windows & Linux** |      **macOS**      |
|:--------------------:|:-------------------:|:-------------------:|
| Insert R chunk       |      Ctrl+Alt+I     |   Command+Option+I  |
| Knit document        |     Ctrl+Shift+K    |   Command+Shift+K   |
| Run current line     |      Ctrl+Enter     | Command+Enter       |
| Run current chunk    |   Ctrl+Shift+Enter  | Command+Shift+Enter |
| Run all chunks above |      Ctrl+Alt+P     |   Command+Option+P  |
| Comment current line |     Ctrl+Shift+C    |   Command+Shift+C   |
| `<-`                 |        Alt + -      |       Option + -    |
| `%>%`                |     Ctrl+Shift+M    |   Command+Shift+M   |

---

## Let's Practice Together! 

Go to [AE 07: Spatial Data and Visualization](https://sta199-summer22.netlify.app/appex/ae07_BJ.html)

---

## Bulletin

- Lab 02 due tonight at 11:59pm

- Submit your `ae07`

- Watch videos for [Prepare: May 23](https://sta199-summer22.netlify.app/prepare/week03_may23_BJ.html)

- HW 01 due Tuesday, May 24 at 11:59pm

- General questions about GitHub $\rightarrow$ Slack

- Personal matter $\rightarrow$ email me with "STA199" in the subject

- Late submission for HW and Lab assignments up to three days with 10% deduction for each 24-hour period

