---
title: "Conditional Probability"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Material 

`r emo::ji("movie_camera")` Watch [Conditional Probability](https://warpwire.duke.edu/w/9RUGAA/)

- [Slides](https://sta198f2021.github.io/website/slides/week-02/w2-l04-condprob.html)

`r emo::ji("movie_camera")` Watch [Simpson's Paradox](https://www.youtube.com/watch?v=sdas62v0iJU)

- [Slides](https://rstudio-education.github.io/datascience-box/course-materials/slides/u2-d16-simpsons-paradox/u2-d16-simpsons-paradox.html#1)

---

## Today's Goal 

- Define marginal, joint, and conditional probabilities, and calculate each “manually” and in a reproducible way
- Identify when events are independent
- Apply Bayes' theorem to examine COVID-19 test specificity

---

## Quiz 

**Q - Find the correct term, a mathematical notation with events** $A$ **and** $B$, **and an example for each of the following descriptions.**

"The probability an event occurs regardless of values of the other event"

--

- Marginal probability 

--

- $P(A)$ or $P(B)$

--

- The probability that a randomly selected student in STA199 favors dogs over cats.

---

## Quiz 

**Q - Find the correct term, a mathematical notation with events** $A$ **and** $B$, **and an example for each of the following descriptions.**

"The probability two or more events simultaneously occur"

--

- Joint probability 

--

- $P(A \cap B)$ 

--

- The probability that a randomly selected student in STA199 favors dogs **and** their favorite movie genre is drama.

---

## Quiz 

**Q - Find the correct term, a mathematical notation with events** $A$ **and** $B$, **and an example for each of the following descriptions.**

"The probability an event occurs given the other has occurred"

--

- Conditional probability 

--

- $P(A|B) = \frac{P(A \cap B)}{P(B)}$ or $P(B|A)$

--

- The probability that a randomly selected student in STA199 favors dogs **given** their favorite movie genre is drama.

--

- $A \text{ given } B$ restricts our attention to $B$ from the whole sample space $\Omega$.

---

## Quiz 

**Q - What is the multiplicative rule of probability?** 

--

$$P(A \cap B) = P(A|B)P(B)$$

Comes directly from definition of conditional probability.

--

**Q - What is the law of total probability?** 

--

$$P(A) = P(A|B)P(B) + P(A|B^c)P(B^c) = P(A \cap B) + P(A \cap B^c)$$

Let's verify it with a Venn Diagram!

---

## Quiz 

**Q - What are the independent events?**

--

- Knowing one event has occurred does not lead to any change in the probability we assign to another event.

--

- $P(A|B) = P(A)$ or $P(B|A) = P(B)$

- $P(A\cap B) = P(A)P(B)$ if $A$ and $B$ are independent.

--

- P(dogs|drama) = P(dogs)

- Knowing one's movie taste tells us nothing about his/her preference of dogs over cats. 

---

## Quiz 

**Q - Are two disjoint events with positive marginal probability independent?**

--

No

$P(A)P(B) > 0$ while $P(A \cap B)=0$ and thus $P(A \cap B) \neq P(A)P(B)$.

---

**Q - Are dying and abstaining from coffee independent events?**

Let $A$ be the event that a man died and $B$ be the event that a man was a non-coffee drinker. 

.midi[
```{r, echo = F}
coffee <- c(rep("Does not drink coffee", 6477), 
            rep("Drinks coffee occasionally", 29809), 
            rep("Drinks coffee regularly", 28535))
died <- c(rep("Died", 1039), rep("Did not die", 5438), 
          rep("Died", 4440), rep("Did not die", 25369),
          rep("Died", 3601), rep("Did not die", 24934))
knitr::kable(table(coffee, died))
```
]

--

- $P(A) = 9080/64821$
- $P(B) = 6477/64821$
- $P(A \cap B) = 1039/64821$
--

- $P(A)P(B) = 0.014 \approx P(A \cap B) = 0.016$

--
They are independent!

---

## Quiz 

**Q - What is the probability that a man was a non-coffee drinker given he died?**

- $P(A) = 9080/64821$
- $P(B) = 6477/64821$
- $P(A \cap B) = 1039/64821$

--

$P(B|A) = P(B \cap A) / P(A) = 1039 / 9080 = 0.114$ 

---

## Quiz 

**Q - What is the probability that a man died given he was a non-coffee drinker?**

- $P(A) = 9080/64821$
- $P(B) = 6477/64821$
- $P(A \cap B) = 1039/64821$

--

$P(A|B) = P(A \cap B) / P(B) = 1039 / 6477 = 0.160$ 

---

## Quiz 

**Q - ** $P(A|B)$ **and** $P(B|A)$ **are not the same. Are they related in some way?**

--

Yes - According to Bayes' theorem, 

$$P(A|B) = \frac{P(B|A)P(A)}{P(B)}.$$
If we know marginal probabilities and one conditional probability, we can always compute the other conditional probability! 

This can be viewed as updating prior belief $P(A)$ to posterior belief $P(A|B)$ after seeing new information $B$. 

---

## Quiz 

**Q - Let's recompute** $P(A|B)$ **using Bayes' theorem.**

- $P(A) = 9080/64821$
- $P(B) = 6477/64821$
- $P(B|A) = 1039/9080$

--

\begin{align}
P(A|B) &= \frac{P(B|A)P(A)}{P(B)} \\
&= \frac{1039/9080 \times 9080/64821}{6477/64821} \\
&= \frac{1039}{6477} = 0.160
\end{align}

---

## Quiz 

**Q - What is Simpson's paradox?**

--

The effect where the inclusion of a third variable in the analysis can change the apparent relationship between the other two variables

--

**Q - Explain Simpson's paradox in UC Berkeley admission data in mathematical expressions.**

--

Overall, P(Admit|Female) $<$ P(Admit|Male) 

However, P(Admit|Dept, Female) $\approx$ P(Admit|Dept, Male) or even P(Admit|Dept, Female) $>$ P(Admit|Dept, Male) 

---

## Quiz 

**Q - The effect of the function** `group_by()` **applies to all the subsequent operations. What function do we need to undo grouping?**

--

- `ungroup()`

- `summarize()` also peels off one layer of grouping by default; if there were multiple grouping variables, then the dataset would still be grouped even after summarize.

---

class: middle, center

# Questions?

---

## Let's Practice Together! 

Go to [AE 11: Conditional Probability](https://sta199-summer22.netlify.app/appex/ae11_BJ.html)

---

## Bulletin

- Exam 01 on Friday! 



