---
title: "Final Project Proposal"
author: "Bora Jin"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
courage <- read_csv("../labs/data/courage.csv")
```

## Project Proposal Steps

1. Find a dataset that satisfies the [final project guidelines](https://sta199-summer22.netlify.app/project/project_BJ.html)

2. Write about:

  - the source of data
  - when and how it was originally collected (by the curator, not necessarily how you found the data)
  - a brief description of the observations and variables you intend to explore

3. Choose 1-2 research questions

4. `glimpse` the data

---

## Ex: Introduction / data

Dataset #1: NC Courage Homefield Advantage 

Our first dataset comes from the [National Women's Soccer League (NSWL) Github](https://github.com/adror1/nwslR) and was sourced from [nwslsoccer.com](https://www.nwslsoccer.com/).

The dataset contains 78 observations (soccer games) played by the NC courage 
spanning three seasons: 2017, 2018, 2019. There are 10 variables in this dataset. 
Some of the variables we care about are `home_team`, `away_team`, and `result` 
(of the game).

---
## Ex: Research question(s):

1. Does NC Courage have a home-field advantage? We hypothesize that NC Courage is more likely to win on their home field than another team's field.

  - To answer this question we will use information about the `home_team` and the 
`result` of the game.

2. Does winning propagate winning? When NC Courage win a game, does it increase the probability of winning the very next game?

  - To answer this question we will use information about the `result` of the game 
  and the `game_number`.

---

## Ex: Glimpse
```{r glimpse-courage}
glimpse(courage)
```


