---
title: "Strings in R"
subtitle: "<br> <br> Introduction to Data Science"
author: "[introds.org](https://www.introds.org/)"
output:
  xaringan::moon_reader:
    mathjax: "https://cdn.bootcss.com/mathjax/2.7.1/MathJax.js?config=TeX-MML-AM_HTMLorMML"
    css: ["xaringan-themer.css", "slides.css"]
    # logo: img/sta199-sticker-icon.png
    lib_dir: libs/font-awesome
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: "%current%" 
      ratio: "16:9"
---

```{r setup, include=FALSE}
# R options
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE,     # for regression output
  warm = 1
  )
# Set dpi and height for images
knitr::opts_chunk$set(fig.height = 3, fig.width = 5, dpi = 300, 
                      warning = FALSE, 
                      message = FALSE, 
                      fig.align = "center") 
# ggplot2 color palette with gray
color_palette <- list(gray = "#999999", 
                      salmon = "#E69F00", 
                      lightblue = "#56B4E9", 
                      green = "#009E73", 
                      yellow = "#F0E442", 
                      darkblue = "#0072B2", 
                      red = "#D55E00", 
                      purple = "#CC79A7")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
# For magick
dev.off <- function(){
  invisible(grDevices::dev.off())
}
# For ggplot2
ggplot2::theme_set(ggplot2::theme_bw())
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(ape)
# library(seqinr)
# library(usmap)
# library(ggthemes)
```


```{r echo = FALSE}
states <- c("alabama", "alaska", "arizona", "arkansas", "california", 
            "colorado", "connecticut", "delaware", "florida", "georgia", 
            "hawaii", "idaho", "illinois", "indiana", "iowa", "kansas", 
            "kentucky", "louisiana", "maine", "maryland", "massachusetts", 
            "michigan", "minnesota", "mississippi", "missouri", "montana", 
            "nebraska", "nevada", "new hampshire", "new jersey", 
            "new mexico", "new york", "north carolina", "north dakota", "ohio", 
            "oklahoma", "oregon", "pennsylvania", "rhode island",
            "south carolina", "south dakota", "tennessee", "texas", "utah", 
            "vermont", "virginia", "washington", "west virginia", "wisconsin",
            "wyoming")
```


## What is a "string"?

A string is a collection of characters placed between quotes.

A character is a single input from your keyboard (e.g. a single letter or a single punctuation mark).

```{r message = FALSE, warning = FALSE, eval = FALSE}
string1 <- "Hi!"
string2 <- 'Hello, I am C-3PO, it is a pleasure to meet you.'
```


You can combine strings in a vector.

```{r message = FALSE, warning = FALSE}
string3 <- c("It's against", "my programming", "to use inconsistent notation.")
string3
```

---
class: middle, center

`stringr`
---

```{r message = FALSE, warning = FALSE}
library(stringr)
```

... but it's also included in the `tidyverse`!

--

`stringr` provides many tools to work with strings, including functions that

- count the characters in a string: `str_count()`

- concatenate string vectors `str_c()`

- detect patterns `str_detect()`

- trim whitespace `str_trim()`

--

Begin with `str_`

All take a vector of strings as their first argument


---

## Include a quotation in a string?

Why doesn't the code below work?

.midi[
```{r eval = TRUE, error = TRUE}
string3 <- "I say "Hello" to the class"
```
]

--

To include a double quote in a string, *escape it* using a backslash `\`.

--

.midi[
```{r}
string4 <- "I say \"Hello\" to the class"
```
]

--

What if you want to include an actual backslash?

--

.midi[
```{r}
string5 <- "\\"
```
]

This may seem tedious but it will come up later!

---

## `writeLines`

`writeLines` shows the contents of the string not
including escapes.

.pull-left[
```{r}
string4
writeLines(string4)
```
]
.pull-right[
```{r}
string5
writeLines(string5)
```
]


---
## rockyou.txt

- RockYou developed software for social media platforms such as MySpace and Facebook

- Stored user passwords in plain text files

- Hacked in 2009 and over 32 million passwords leaked

```{r, echo = FALSE}
rockyou = readLines("/home/athos/GitFolders/hashcat/rockyouSmall.txt")
```
Let's look at the first 20
```{r}
rockyou20 <- rockyou[1:20] 
rockyou20
```

---

## `str_length`

Given a string, return the number of characters.

.midi[
```{r}
password = "qwerty"
str_length(password)
```
]

Given a vector of strings, return the number of characters in each string.

.midi[
```{r}
str_length(rockyou20)
```

```{r}
rockyou20
```
]

--

.pull-left[
- Alabama: 7
- Alaska: 6
- Arizona: 7
- Arkansas: 8
]
.pull-right[
- California: 10
- Colorado: 8
- Connecticut: 11
- ...
]
 

---

## `str_c`

Combine two or more strings.

```{r}
str_c("My", "password", "is", "qwerty")
```


--

Use `sep` to specify how the strings are separated.

```{r}
str_c("My", "password", "is", "qwerty", sep = " ")
```


---

## `str_to_lower` and `str_to_upper`

Convert the case of a string from lower to upper or upper to lower.

.midi[
```{r}
str_to_upper(rockyou20)
```
]


---

## `str_sub`

Extract parts of a string from `start` to `end`, inclusive.

.midi[
```{r}
str_sub(rockyou20, 1, 4)
```
]

--

.midi[
```{r}
str_sub(rockyou20, -4, -1)
```
]

---

## `str_sub` and `str_to_upper`


Can combine `str_sub` and `str_to_upper` to capitalize each password.

.midi[
```{r}
str_sub(rockyou20, 1, 1) <- str_to_upper(str_sub(rockyou20, 1, 1))
rockyou20
```
]

---

## `str_sort`

Sort a string. Here we sort in decreasing alphabetical order.

.midi[
```{r}
str_sort(rockyou20, decreasing = TRUE)
```
]

---

## Regular Expressions

A .vocab[regular expression] is a sequence of characters that allows you to 
describe string patterns. We use them to search for patterns. 

- extract a phone number from text data
- determine if an email address is valid
- determine if a password has the required number of letters, characters, and symbols
- count the number of times "statistics" occurs in a corpus of text
- ...

---

## Basic Match

To demonstrate the power of regular expressions, let's see if any of the 32 million leaked passwords contain the exact phrase "dog"

```{r, echo = FALSE}
# R2D2_lines = c(1058522, 2101277, 2101278, 2467375, 10709025, 10709026)
# rockyou =  readLines("/home/athos/GitFolders/hashcat/R2D2_lines.txt")
```


```{r}
str_subset(rockyou, "dog")[1:30]
```



---

## Basic Match

What about "d-g"? Match any character using `.`


```{r}
str_subset(rockyou, "d.g")[1:30]
```


---

## Anchors

Match the start of a string using `^`

```{r echo = FALSE}
rockyou20 = rockyou20[1:9]
```


.inverse[
```{r}
str_view_all(rockyou20, "^P")
```
]

```{r echo = FALSE}
rockyou20 = rockyou[1:20]
```


```{r}
rockyou20
```

---

## Anchors

Match the end of a string using `$`

.inverse[
```{r}
str_view_all(rockyou20, "u$", match = TRUE)
```
]
---

## `str_detect`

Determine if a character vector matches a pattern.

```{r}
rockyou20
str_detect(rockyou20, "a")
```

---

## `str_count`

How many matches are there in a string?


```{r}
rockyou20
str_count(rockyou20, "s")
```

---

## `str_replace_all`

Replace all matches with new strings.

```{r}
str_replace_all(rockyou20, "s", "-")
```



---

## Many Matches

The regular expressions below match more than one character.

- Match any digit using `\d` or `[[:digit:]]`
- Match any whitespace using `\s` or `[[:space:]]`
- Match f, g, or h using `[fgh]` 
- Match anything but f, g, or h using `[^fgh]`
- Match lower-case letters using `[a-z]` or `[[:lower:]]`
- Match upper-case letters using `[A-Z]` or `[[:upper:]]`
- Match alphabetic characters using `[A-z]` or `[[:alpha:]]`

Remember these are regular expressions! To match digits you'll need to *escape*
the string, so use `"\\d"`, not `"\d"`

---

## Additional resources

- `stringr` website: https://stringr.tidyverse.org/
- `stringr` and `regex` [cheat sheet](https://raw.githubusercontent.com/rstudio/cheatsheets/master/strings.pdf)
- [Chapter 14: Strings](https://r4ds.had.co.nz/strings.html#matching-patterns-with-regular-expressions) in R for Data Science

