---
title: '`R` errors, troubleshooting, and getting help'
output:
  html_document:
    df_print: paged
date: "9.23.2021"
editor_options:
  chunk_output_type: console
---

### Bulletin

- Due today: Homework 02
- Upcoming:
  - No lab this week
  - Exam 01 released tomorrow morning (GitHub), due Tuesday by 11:59p
  - Probability next Tuesday, Quiz 04 due Thursday

### Main Ideas

- Encountering errors and misbehaving code is common at all levels of ability.
- Employing a few simple strategies and steps to follow will help you diagnose
and fix code problems quickly and independently.
- Troubleshooting code takes practice.

### Lecture Notes and Exercises

We will use the packages below.

```{r load-packages, message = FALSE}
library(tidyverse)
```

`R` or `R Markdown` will provide an error message if a problem with a
computation occurs.

Encountering errors is extremely common and figuring out how to fix them 
is frustrating and time-consuming. But errors are helpful! They provide 
you with information that will help you fix your code.

In today's lecture we will introduce a step-by-step process you can follow 
whenever you encounter trouble with code. Additionally, we will introduce a few
strategies to help you fix problematic code.

## Troubleshooting steps

**Step #1:** Don't panic! 

Errors are extremely common and fixing them quickly takes practice. Try not to 
panic or get frustrated. 

**Step #2:** Find the relevant code.

Navigate to the code chunk where the error or problem occurred. If the error 
occurred while running a code chunk this is already done. If the error occurred
while knitting, `R` usually provides a line number. If possible, navigate to the
code chunk in question so you can see both the code and error at the same time.

**Step #3:** Read the error.

Pause and read the error carefully and in full. What does it say in plain 
English? **Usually**, there is enough information provided to both diagnose and 
fix the problem.

Below are some common errors that have enough information to fix the problem.

- `...could not find function...`

`R` can't find a function that you used. Did you include a code chunk loading
the necessary packages? If you did is `eval = FALSE` included as a code chunk
option?

- `...object 'some-name-of-object' not found...`

`R` can't find `some-name-of-object`. Is `some-name-of-object` spelled and
formatted consistently, including correct capitalization? Was 
`some-name-of-object` actually created in a code chunk where `eval` is not set
to `FALSE`? Is the code chunk located **above** the code chunk where the
error occurred? Is it stored via `<-`? The "Environment" pane can be helpful 
here.

- `...unexpected 'some-symbol' in...`

Here `some-symbol` can be a comma, parenthesis, bracket, etc. This means the 
code you are running is not correct syntactically. Do you have an extra comma, 
parenthesis, bracket, or other symbol? Did you make a typo?

- `Error: attempt to use zero-length variable name`

This generally happens when you highlight the backticks in the code chunk 
in addition to the code. It can also occur if there is a stray `%>%` or `+` at 
the end of a code chunk.

- `...requires the following missing aesthetics...`

`ggplot()` is missing a necessary aesthetic. Is the `ggplot()` formatted 
correctly? Is the aesthetic provided consistent with what is required? 

- `object of type 'closure' is not subsettable`

The above error is included because it is pretty common. It (usually) means
that you tried to subset a function.

- `non-numeric argument to binary operator`

Occurs when you mix different data types in a calculation.

Often, closely reading the error will allow you to fix the problem. 

**Step #4:** Run the code line by line.

For a code chunk with a number of lines, it is sometimes tricky to tell where
exactly the error is occurring. In this situation, it is helpful to run the 
code line-by-line to figure out where the error is occurring.

The code chunk below has a few small errors. Let's demonstrate how running the
code line-by-line helps us troubleshoot.

```{r line-by-line, eval = FALSE}
mpg %>%
  filter(class == "subcompact") %>%
  group_by(drv)
  summarize(median_cty_mpg = median(cty),
            sd_cty_mpg = sd(cty),
            avg_cty_mpg = average(city)))
```

You try!
How does highway mileage vary by car manufacturer? Debug the code below.
```{r, eval = FALSE}
car_companies = c("toyota", "dodge", "ford", "honda", "subaru" "volkswagen")
mpg %>%
  filter(is.element(manufacturer, car_companies))
  ggplot(aes(x = manufacturer, y = hwy)) +
  geo_point()
```


**Step #5: Examine the Documentation**

If an error is from a particular function or argument, pulling up the 
documentation is a good step.

Documentation in `R` is extremely helpful. If you want to understand what a 
function does, its arguments, or examples of usage, examine the documentation.
In many situations, it should be higher than the fifth troubleshooting step.

Documentation can be examined using `?` or `help()`. 

- Description: a general description of what the function does
- Usage: the arguments of the function and their defaults
- Arguments: a description of each argument and what it does
- Details: details about the function
- Examples: examples of function usage

**Step 6: Search online**

Generally you are not the first person to encounter a particular error in `R`. 
A well-thought out search can lead to others who have encountered the
same or a similar problem and potentially a solution.

**Include** general search terms related to the error. Include quotation
marks to search for an exact phrase and include aspects of the error that are
unique. If the error is with a function in a particular package (`rvest`, 
`dplyr`, etc) include that with your search. Include `R` as a search term.

**Avoid** search terms that are specific to your current project. This includes
datasets, variables, and aspects of your personal system (file paths, etc).

Google searches offer advanced search operators. A helpful official Google link is
[here](https://support.google.com/websearch/answer/2466433?hl=en) and an 
unofficial blog post is [here](https://www.lifewire.com/advanced-google-search-3482174).
Check out partial search, domain search, and words by proximity.

[StackOverflow](https://stackoverflow.com/) is an extremely helpful site.
Check out the [`R` related topics](https://stackoverflow.com/questions/tagged/r)
with the tag `R`.

**Examine a Vignette**

For a broad overview of the capabilities of a package it is helpful to examine
a vignette. Vignettes are "discursive documents meant to illustrate and explain
facilities in [a] package" (R-Project).

Use `browseVignettes()` to see vignettes from all installed packages and
`browseVignettes(package = some-package)` to see vignettes from a particular
package. Use `vignette("vignette-name")` to see a particular vignette.

Let's try examining a vignette from a previous lecture.

**Step #7:** Send a message on slack.

First (if possible) push your most recent work to GitHub.

On slack, note

- What assignment and problem you are working on.
- What you are trying to accomplish.
- What issue you encountered.
- A (brief) overview of the steps you have taken to solve the issue.

Include both the code and error in a code chunk using the "Markdown editor". 
You can create a code chunk in the same way we do in an `R` Markdown document.

If possible, include a reproducible example of the error.

### Reproducible examples

A reproducible example is a very small, toy example used to recreate
the issue for yourself and others. Often, devising a reproducible example helps you diagnose
the issue. 

Check out [How to make a great R Reproducible Example](https://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example)
and [How to create a Minimal, Reproducible Example](https://stackoverflow.com/help/minimal-reproducible-example).

A reproducible example should be:

- minimal: use as little code as possible to reproduce the issue
- complete: should include all aspects necessary to reproduce the problem
- reproducible: should generate the issue in question

Check out the examples [here](https://stackoverflow.com/questions/24576515/relative-frequencies-proportions-with-dplyr/24576703#24576703) and [here](https://stackoverflow.com/questions/35090883/remove-all-of-x-axis-labels-in-ggplot).

The function `tribble()` is helpful for creating small, toy datasets row-by-row.

### Exam prep

Compare the highway fuel efficiency of Dodge vs other manufacturers.
```{r, eval=FALSE}
mpg 
```

Create a new column `fuel_efficiency` that separates vehicles into 3 categories: worst gas mileage (hwy <16 mpg) from average (hwy 16-24 mpg) and efficient vehicles (hwy >24 mpg).

```{r, eval=FALSE}

```


### Additional Resources

- [R Project Help](https://www.r-project.org/help.html)
- [Advanced Debugging](https://adv-r.hadley.nz/debugging.html)
- [Reprex package](https://www.tidyverse.org/help/#reprex-pkg)
- [How to write a reproducible example](http://adv-r.had.co.nz/Reproducibility.html)