---
title: 'AE 17: Hypothesis tests for independence'
author: "Your name"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_height: 3
    fig_width: 5
---

```{r load-packages, warning = F, message = F}
library(tidyverse)
library(tidymodels)
```

```{r load-data, message = F}
yawn <- read_csv("data/yawn.csv")
```

# Bulletin

-   Due today:

    -   lab 06

-   Due Thursday:

    -   prep quiz 06

-   Upcoming

    -   proposals back to you by end of Wednesday

# Learning goals

Use and understand simulation-based methods to ...

-   test a claim about a population proportion
-   test a claim about independence between two groups

## Test for independence

-   First let's, watch the experiment from *Mythbusters*.

-   Let $t$ be the treatment group who saw a person yawn, $c$ be the control group who did not see anyone yawn, and $p$ be the proportion of people who yawned.

## Exercise 1[^1]

[^1]: Simulation activity from [Data science in a box](https://datasciencebox.org/)

We want to use simulation-based inference to assess whether or not yawning and seeing someone yawn are independent.

-   State the null and alternative hypotheses in words:

-   Select the appropriate null and alternative hypotheses written in mathematical notation:

a. $H_0: p_t = p_c \text{ vs. }H_a: p_t < p_c$
b. $H_0: p_t = p_c \text{ vs. }H_a: p_t > p_c$
c. $H_0: p_t = p_c \text{ vs. }H_a: p_t \neq p_c$
d. $H_0: \hat{p}_t = \hat{p}_c \text{ vs. }H_a: \hat{p}_t < \hat{p}_c$
e. $H_0: \hat{p}_t = \hat{p}_c \text{ vs. }H_a: \hat{p}_t > \hat{p}_c$
f. $H_0: \hat{p}_t = \hat{p}_c \text{ vs. }H_a: \hat{p}_t \neq \hat{p}_c$

## Exercise 1.5

Type I and Type II error

| Truth         | Reject $H_0$  | Fail to reject $H_0$|
|---------------|---------------|----------------|
| $H_0$ is true  | Type 1 error (false +)  | Good decision  |
| $H_0$ is false | Good decision | Type 2 error (false -)   |

What does type I error look like in this case?

What does type II error look like?

Is type I or type II error more worrisome here?

How might we mitigate the more dangerous error?

[Click here](https://openintro-ims.netlify.app/decerr.html?q=type%20I%20error#controlling-the-type-1-error-rate) for further reading.

## Exercise 2

Before using R to construct the null distribution, let's generate the null distribution using playing cards! See AE-17 for the simulation instructions. You can also find them in the `README` of this application exercise.

## Exercise 3

Uncomment the code to see read in the data from the class and visualize the null distribution.

```{r sim-results}
#sim_data <- read_csv("https://sta199-f21-001.netlify.app/appex/data/yawn-sim.csv")
```

```{r sim-dist, eval = F}
#ggplot(data = sim_data, mapping = aes(x = diff_in_prop)) +
#  geom_histogram(binwidth = 0.05) + 
#  labs(title = "Your Results: Difference in Proportion of Yawners")
```

-   What is the approximate center of the distribution? Is this what you expected? Why or why not?

-   The observed difference in proportions from the Mythbusters episode is 0.0441. Based on your simulated distribution, do yawning and seeing someone yawn appear to be dependent?

## Exercise 4

Let's use the data from the *Mythbusters* episode and simulation-based inference in R to test this claim. Based on their experiment, do yawning and seeing someone yawn appear to be dependent?

Evaluate this question using a simulation based approach. We will use the same null and alternative hypotheses as before. The data from *Mythbusters* is available in the `yawn` data frame.

-   Fill in the code below to generate the null distribution. Uncomment the code once it is complete.

```{r mythbusters-null}
set.seed(101921)
#null_dist <- yawn %>%
#  specify(response = ____, explanatory = _____, success = "yawn") %>%
#  hypothesize(null = "______") %>%
#  generate(100, type = "permute") %>%
#  calculate(stat = "______", 
#            order = c("trmt", "ctrl"))
```

-   Visualize the null distribution and shade in the area used to calculate the p-value.

```{r mythbusters-null-viz}
# add code 
```

-   Calculate p-value. Then use the p-value to make your conclusion using a significance level of 0.1.

```{r mythbusters-calc-p-value}
# add code
```

## Exercise 5

Do you believe the conclusions from this experiment? Why or why not?
