---
title: 'AE 15: Bootstrapping'
author: "Your Name"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_height: 3
    fig_width: 5
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

```{r load-packages}
library(tidyverse)
library(tidymodels)
```

```{r load-data}
manhattan <- read_csv("data/manhattan.csv")
law = read_csv("data/lsat_gpa.csv")
```
# Bulletin
- Due Thursday
  - homework 03
  - quiz 05
- Due Friday
  - project proposal
  
- Upcoming
  - stats experience
  
# Learning goals

- Understand how to draw a bootstrap sample and calculate a bootstrap statistic 
- Use `infer` to obtain a bootstrap distribution 
- Calculate a confidence interval from the bootstrap distribution
- Interpret a confidence interval in context of the data 
 
# Rent in Manhattan 

On a given day in 2018, twenty one-bedroom apartments were randomly selected on Craigslist Manhattan from apartments listed as "by owner". The data are in the `manhattan` data frame.  We will use this sample to conduct inference on the typical rent of 1 bedroom apartments in Manhattan. 

# Part 1: recap; drawing a bootstrap sample

Let's start by using bootstrapping to estimate the **mean** rent of one-bedroom apartments in Manhattan. 

## Exercise 1

What is the point estimate of the typical rent? 


## Exercise 2

Recap: last time we did a manual bootstrap by sampling with replacement the rent values from a box.

- How many draws do we need for our bootstrap sample? 

```{r boot-sample}
class_bootstrap <- c(2150, 1795, 3800, 3800, 3200, 3950, 3800, 3267, 2300, 2300, 2300, 3267, 2350, 1570, 2350, 2175, 1775, 4195, 2350, 4195)
```

## Exercise 3

- About what value do you expect the bootstrap statistic to take? 
- Calculate the statistic from the bootstrap sample. 

```{r calc-boot-stat}
# add code

```

# Part 2: Bootstrap confidence interval

**We will use the `infer` package, included as part of `tidymodels` to calculate a 95% confidence interval for the mean rent of one-bedroom apartments in Manhattan.**

We start by setting a seed to sure our analysis is reproducible. We'll use 101221 to set our seed to today's date but you can use any value you want on assignments.

```{r set-seed}
set.seed(101221)
```

## Generating the bootstrap distribution

We can use R to take many bootstrap samples and generate a bootstrap distribution

Uncomment the lines and fill in the blanks to create the bootstrap distribution of sample means and save the results in the data frame `boot_dist`.    

Use **500 reps** for the in-class activity. (You will use about 15,000 reps for assignments outsdie of class.)

```{r create-bootstrap-dist}
boot_dist <- manhattan #%>%
  #specify(______) %>%
  #generate(______) %>%
  #calculate(______)
```

- How many rows are in `boot_dist`? 
- What does each row represent? 
- What are the variables in `boot_dist`? What do they mean?

## Visualize the bootstrap distribution 

Visualize the bootstrap distribution using a histogram. Describe the shape, center, and spread of this distribution. 

```{r viz-bootstrap}
# add code
```

## Calculate the confidence interval

Uncomment the lines and fill in the blanks to construct the 95% bootstrap confidence interval for the mean rent of one-bedroom apartments in Manhattan. 

```{r confidence-interval-95}
#___ %>%
#  summarize(lower = quantile(______),
  #          upper = quantile(______))
```

## Interpret the interval

Write the interpretation for the interval calculated above. 

# Part 3: Changing the confidence level 

- Modify the code used to calculate a 95% confidence interval to calculate a **90% confidence interval** for the mean rent of one-bedroom apartments in Manhattan. How does the width of this interval compare to the width of the 95% confidence interval? 

```{r confidence-interval-90}
#calculate a 90% confidence interval
```

- Now let's calculate a 99% confidence interval for the mean rent of one-bedroom apartments in Manhattan. How does the width of this interval compare to the width of the 95% confidence interval? 

```{r confidence-interval-99}
#calculate a 99% confidence interval
```

- What is one advantage to using a 90% confidence interval instead of a 95% confidence interval to estimate a parameter? - What is one advantage to using a 99% confidence interval instead of a 95% confidence interval to estimate a parameter?

# Part 4: Additional practice

Next, use bootstrapping to estimate the median rent for one-bedroom apartments in Manhattan. 

- Generate the bootstrap distribution of the sample medians. Use 500 reps. Save the results in `boot_dist_median`.

```{r boot-dist-median}
## add code
```

- Calculate a 92% confidence interval. 

```{r confidence-interval-92}
## add code
```

- Interpret the 92% confidence interval. 

# Part 5: Even more practice

`law` contains data about LSAT (law school exam) scores and GPA (grade point average)

What's the correlation between LSAT score and GPA?
```{r}
# law %>%
```

Report a 95% bootstrap confidence interval on the sample correlation
```{r}
# add code
```

