---
title: 'AE 18: Sampling distributions & the Central Limit Theorem'
author: "YOUR NAME GOES HERE"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document:
    fig_height: 3
    fig_width: 5
---

# Bulletin

- Due today:
  - quiz 06 on sakai (now)
  - lab 07 by 11:59pm

- Misc
  - Hw 04 due Saturday by 11:59pm
  - Proposal feedback on gradescope, be sure to read the notes
  - If you haven't, check out AE17 solutions and additional example on sakai
  

# Learning goals 

- Recap simulation-based methods and `infer`
- Use Central Limit Theorem to define distribution of sample means
- Calculate probabilities from the normal distribution 

```{r load-package, message=F}
library(tidyverse)
```


# Recap of simulation-based methods and `infer`


# Practice using CLT & Normal distribution

Suppose the bone density for 65-year-old women is normally 
distributed with mean $809 mg/cm^3$ and standard deviation of $140 mg/cm^3$.

Let $x$ be the bone density of 65-year-old women. We can write this distribution of $x$ in mathematical notation as 

$$x \sim N(809, 140)$$
## Visualize the population distribution

```{r visualize-normal}
ggplot(data = data.frame(x = c(809 - 140*3, 809 + 140*3)), aes(x = x)) +
  stat_function(fun = dnorm, args = list(mean = 809, sd = 140),
                color = "black") +
  stat_function(fun = dnorm, args = list(mean = 809, sd = 140/sqrt(10)),
                color = "red",lty = 2) + theme_bw() +
  labs(title = "Black solid line = population dist., Red dotted line = sampling dist.")
```



## Exercise 1

Before typing any code, based on what you know about the normal distribution, what do you expect the median bone density to be?

What bone densities correspond to $Q_1$ (25th percentile), $Q_2$ (50th percentile), and $Q_3$ (the 75th percentile) of this distribution? Use the `qnorm()` function to calculate these values. 

## Exercise 2

The densities of three woods are below:

- Plywood: 540 mg/cubic centimeter
- Pine: 600 mg/cubic centimeter
- Mahogany: 710 mg/cubic centimeter

- What is the probability that a randomly selected 65-year-old woman has bones less dense than Pine?

- Would you be surprised if a randomly selected 65-year-old  woman had bone density less than Mahogany? What if she had bone density less than Plywood? Use the respective probabilities to support your response. 

## Exercise 3

Suppose you want to analyze the mean bone density for a group of 10 randomly selected 65-year-old women. 

- Are the conditions met use the Central Limit Theorem to define the distribution of $\bar{x}$, the mean density of 10 randomly selected 65-year-old women? 
  - Independence? 
  - Sample size/distribution? 
  
- What is the shape, center, and spread of the distribution of $\bar{x}$, the mean bone density for a group of 10 randomly selected 65-year-old women?

- Write the distribution of $\bar{x}$ using mathematical notation. 

## Exercise 4

- What is the probability that the mean bone density for the group of 10 randomly-selected 65-year-old women is less dense than Pine?

- Would you be surprised if a group of 10 randomly-selected 65-year old women had a mean bone density less than Mahogany? What the group had a mean bone density less than Plywood? Use the respective probabilities to support your response. 

## Exercise 5

Explain how your answers differ in Exercises 2 and 4.