---
title: "AE 21: Regression pt. 2 and multiple regression"
author: "YOUR NAME GOES HERE"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r, warning = FALSE, message  = FALSE}
library(tidyverse)
library(tidymodels)
library(scatterplot3d)
```

# Bulletin

- Due today
  - [todo]

- Upcoming
  - [todo]


# Learning goals

- Gain proficiency with multiple linear regression
- Understand what a dummy variable is

To begin, we'll work with a new set of pokemon data.

```{r load-pokemon, message=F}
pokemon <- read_csv("data/pokemon150.csv")
```

## Exercise 1) 

Previously we looked at building a model to predict a pokemon's hit points using height.

Last time we noticed *both* height *and* weight are correlated with a pokemon's hit points.

Other variables may be correlated with `hp`, e.g. a pokemon's legendary status. 

Do legendary pokemon have higher `hp` than non-legendary pokemon? Compare mean `hp` between groups to support your answer.

```{r hp-vs-legendary}
# code here
```


Write down a model to predict a pokemon's hitpoints based on their height, weight, legendary status (use $x$, $y$, $\beta$ notation). Define each variable.

## Exercise 2) 

Use `tidymodel` syntax to build a linear model and estimate each $\beta$.

```{r hp-lm}
# code here
```
Interpret the meaning of your estimates and write a brief description below. Use/discuss the the phrase "all else held constant" with your neighbor.


## Exercise 3)

Do the coefficients match your expectations for the previous exercise? Why or why not? Write code to explore any oddities further.

```{r explain-legendary-coefficient}
# code here
```


## Exercise 4)

Some think that certain pokemon types have higher `hp` than others.

```{r hp-vs-type}
pokemon %>%
  group_by(type_1) %>%
  summarize(mean_hp = mean(hp), n = n())
```
#### a) Construct a linear model in `R` to determine the effect of pokemon type on `hp`. 

```{r hp-type-lm}
# code here
```


#### b) Interpret the meaning of your estimates and write a brief description below. Are any types missing?


#### c) Given the output of your code, write down the linear model (use $x$, $y$, $\beta$ notation). Define each variable.
