---
title: "Data Visualization II"
output: html_document
author: "STA199"
date: "01.19.2022"
editor_options: 
  chunk_output_type: console
---

# Bulletin

- Quiz 01 was due today at the start of class
- [FAQ](https://sta199-sp22-003.netlify.app/faq/faq.html)

# Today

By the end of today you will

- be able to explain continuous, discrete, and categorical variables
- understand how to make visualizations and summarize variables according to their type
- develop a faceted plot

## Notes

There are two types of variables **numeric** and **categorical**.

### Types of variables

Numerical variables can be classified as either **continuous** or **discrete**. 
Continuous numeric variables have an infinite number of values between any 
two values. Discrete numeric variables have a countable number of values.

- height
- number of siblings

Categorical variables can be classified as either **nominal** or **ordinal**. 
Ordinal variables have a natural ordering.

- hair color
- education

### Numeric Variables

To describe the distribution of a numeric we will use the terms below.

- shape
  - skewness: right-skewed, left-skewed, symmetric
  - modality: unimodal, bimodal, multimodal, uniform
- center: mean (`mean`), median (`median`)
- spread: range (`range`), standard deviation (`sd`), interquartile range (`IQR`)
- outliers: observations outside the pattern of the data


## Application exercise
We always begin by loading relevant libraries.

```{r load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
```

Next, we load data.

We will continue our investigation of home prices in Minneapolis, Minnesota.

```{r load-data, message = FALSE}
mn_homes <- read_csv("data/mn_homes.csv")
```

Add a `glimpse` to the code chunk below and identify the following variables as
numeric continuous, numeric discrete, categorical ordinal, or categorical
nominal. 

- area
- beds
- community

```{r glimpse-data}
glimpse(mn_homes$community)
```

The `summary` command is also useful in looking at numerical variables. Use this command to look at the numeric variables from the previous chunk.

```{r summarize}
summary(mn_homes$beds)
```

We can use a **histogram** to summarize a numeric variable.

```{r histogram}
ggplot(data = mn_homes, 
       mapping = aes(x = salesprice)) + 
   geom_histogram(bins = 25)
```

A **density plot** is another option. We just connect the boxes in a histogram
with a smooth curve.

```{r density-plot}
ggplot(data = mn_homes, 
       mapping = aes(x = salesprice)) + 
   geom_density()
```

Side-by-side **boxplots** are helpful to visualize the distribution of a numeric
variable across the levels of a categorical variable.

```{r boxplots}
ggplot(data = mn_homes, 
       mapping = aes(x = community, y = salesprice)) + 
       geom_boxplot() + 
       coord_flip() + 
       labs(main= "Sales Price by Community", x= "Community", y="Sales Price")
```

1. **Question:** What is `coord_flip()` doing in the code chunk above? Try removing it to see.

### Categorical Variables

**Bar plots** allow us to visualize categorical variables.

```{r bar-plot}
ggplot(data = mn_homes) + 
  geom_bar(mapping = aes(x = community)) + coord_flip() + 
  labs(main= "Homes by Community", x= "Community", y="Number of Homes")
```

**Segmented bar plots** can be used to visualize two categorical variables.

```{r segmented-bar-plot}
library(viridis)
ggplot(data = mn_homes, mapping = aes(x = community, fill = fireplace)) + 
  geom_bar() +
  coord_flip() + 
  scale_fill_viridis(discrete = TRUE, option = "D", name = "Fireplace?") +
  labs(main 
       = "Fireplaces by Community", x = "Community", y = "Number of Homes")
```

```{r segmented-bar-plot-fill}
ggplot(data = mn_homes, mapping = aes(x = community, fill = fireplace)) + 
  geom_bar(position = "fill") + coord_flip() + 
  scale_fill_viridis(discrete=TRUE, option = "D", name="Fireplace?") +
  labs(main= "Percentage of Homes with a Fireplace by Community", x=
  "Community", y="Percentage of Homes")
```

2. **Question:** Which of the above two visualizations do you prefer? Why? Is this answer always the same?


3. **Solve** There is something wrong with each of the plots below. Run the code for each plot, read the error, then identify and fix the problem.

```{r fix-errors, eval = FALSE}
ggplot(data = mn_homes) +
  geom_point(mapping = aes(
    x = lotsize,
    y = salesprice,
    shape = 21,
    size = .85
  ))
ggplot(data = mn_homes, mapping = (x = otsize, y = area)) +
  geom_point(, shape = 21, size = .85)
ggplot(data = mn_homes) +
  geom_point(
    mapping = aes(x = lotsize, y = area),
    color = community,
    size = 0.85
  )
ggplot(data = mn_homes) +
  geom_point(mapping = aes(x = 1otsize, y = area))
```

General principles for effective data visualization

- keep it simple
- use color effectively
- tell a story

Why is data visualization important? We will illustrate using the 
`datasaurus_dozen` data from the `datasauRus` package.

```{r read-in-dinos, message = FALSE}
library(datasauRus)
```

```{r glimpse-dinos}
glimpse(datasaurus_dozen)
```

The code below calculates the correlation, mean of y, mean of x, standard deviation of y, and standard deviation of x for each of the 13 datasets.

4. **Question:** What do you notice? 


```{r summarize-dinos}
datasaurus_dozen %>% 
   group_by(dataset) %>%
   summarize(r = cor(x, y), 
             mean_y = mean(y),
             mean_x = mean(x),
             sd_x = sd(x),
             sd_y = sd(y))
```

Let's visualize the relationships

```{r visualize-dinos}
ggplot(data = datasaurus_dozen, 
       mapping = aes(x = x, y = y)) + 
   geom_point(size = .5) + 
   facet_wrap( ~ dataset)
```

5. **Question:** Why is visualization important?

## Practice

(1) Modify the code outline to create a faceted histogram examining the 
distribution of year built within each community.

When you are finished, remove `eval = FALSE` and knit the file to see the changes.

```{r ex-1, eval = FALSE}
ggplot(data = ______, mapping = aes(x = ____)) +
  geom_histogram(binwidth = 10) +
  facet_wrap(~ _____) +
  labs(x = "_____", 
      title = "_______________________", 
      subtitle = "________________")
```

### Additional Resources

- https://ggplot2.tidyverse.org/
- https://raw.githubusercontent.com/rstudio/cheatsheets/master/data-visualization-2.1.pdf
- http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html
- https://medium.com/bbc-visual-and-data-journalism/how-the-bbc-visual-and-data-journalism-team-works-with-graphics-in-r-ed0b35693535
- https://ggplot2-book.org/
= https://ggplot2.tidyverse.org/reference/geom_histogram.html
- https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf
- https://github.com/GraphicsPrinciples/CheatSheet/blob/master/NVSCheatSheet.pdf