---
title: "AE 06: Data Wrangling 2"
subtitle: "due Thursday, May 19 at 9:29am"
author: "Bora Jin"
editor_options:
  chunk_output_type: console
output: 
  html_document: 
    toc: true
    toc_float: true
---

## Getting Started 

- Go to the [course GitHub organization page](https://github.com/sta199-summer22) and find the repository entitled "ae06-GitHubUsername".
- Click the green "code" button and copy the `SSH` URL.
- Go to [RStudio](https://vm-manage.oit.duke.edu/containers), select `File` $\rightarrow$ `New Project` $\rightarrow$ `Version Control` $\rightarrow$ `Git` and paste the URL.
- Open the `.Rmd` file and replace "Your Name" with your name. 

## Demonstrations 

```{r load-package, message=FALSE, warning=FALSE}
library(tidyverse)
```

```{r test-joins}
# lab scores
Lab <- tibble(id = c(2, 3, 4, 6),
              lab1 = c(100, 100, 100, 100), 
              lab2 = c(80, 20, 60, 100))
# ae scores
AE <- tibble(id = c(1, 2, 3, 5),
             ae1 = c(100, 80, 50, 100), 
             ae2 = c(90, 100, 40, 95),  
             ae3 = c(70, 80, 60, 100), 
             ae4 = c(90, 70, 30, 90))
            
Lab
AE
```

We will demonstrate each of the joins on these small, toy datasets. 

Print all the scores on the labs and AEs for all six students. `NA` if the assignment is not turned in. 

```{r full-join}
full_join(Lab, AE)
```

Who has completed all the labs and AEs and what are their scores on the completed assignments?

```{r inner-join}
inner_join(Lab, AE)
```

What are the scores of labs and AEs for those who have completed lab assignments?  

```{r left-join}
left_join(Lab, AE)
```

This time, I want the scores for those who have completed AEs. 

```{r right-join}
right_join(Lab, AE)
```

What are the lab scores for those who have completed both AEs and labs?

```{r semi-join}
semi_join(Lab, AE)
```

What are the lab scores for those who have completed labs but not AEs?

```{r anti-join}
anti_join(Lab, AE)
```

How do the join functions above know to join `Lab` and `AE` by `value`? Examine the names to find out.

```{r names}
names(Lab)
names(AE)
```

We will again work with RDU data stored in `rduall_2021.RDS`. This can be downloaded from the `anyflights` package. 

```{r load-rdu-list, message=FALSE, warning=FALSE}
rdu <- readRDS("data/rduall_2021.RDS")
airports <- rdu$airports
flights <- rdu$flights
```

The `R` object `rdu` is a list where each element is a data frame. We will focus on the following data sets: 

- `flights`: Data on all flights that departed from RDU in 2021
- `airports`: Data on airports at which those flights arrived
- `airlines`: Data to translate between two letter carrier codes and names
- `planes`: Construction information about each plane

```{r load-rdu-data}
flights <- rdu$flights
airports <- rdu$airports
airlines <- rdu$airlines
planes <- rdu$planes
```

**Q - ** How are these datasets related? Examine their variable names. 

```{r colnames}
names(flights)
names(airports)
names(airlines)
names(planes)
```

**Q - ** Last time, it was not straightforward to understand Airlines by their carrier code. Let's append full airline names in `flights`. 

```{r flights-airlines}
flights %>% 
  left_join(airlines, by = "carrier") %>%
  rename(carrier_name = name) %>% # new = old
  select(year:day, contains("carrier")) 
```

**Q - ** Suppose you wanted to make a map of the route of every flight from RDU. What variables would you need from which datasets?

Let's add latitude and longitude information of the destination airports in `flights`. Note these `flights` and `airports` have no variables in common so we will have to specify the variable to join by using `by = `. Check out the documentation for more information.

```{r flights-join}
flights %>% 
  left_join(airports %>% 
              select(faa, lat, lon), 
            by = c("dest" = "faa"))
```

## Practice

(1) Create a new dataset `dest_delays` with the median arrival delay in 2021 for each destination. Note this question does not require you to use joins.

```{r make-dest-delays, include = FALSE}
dest_delays <- flights %>% 
  group_by(dest) %>% 
  summarize(median_arr_delay = median(arr_delay, na.rm = TRUE)) 
```

    - **Q - ** How many different destinations could you visit directly from RDU in 2021? 

(2) Join `dest_delays` and `airports`, only including observations that have both delay and airport information. Note `dest_delays` and `airports` have no variables in common so you will need to specify the variables to join using `by` as in the example above.

```{r new-dataset, include = FALSE}
dest_delays %>% 
  inner_join(airports, by = c("dest" = "faa"))
```

    - **Q - ** Are all of the observations in `dest_delays` included in the new dataset you created by joining `dest_delays` and `airports`? Use an appropriate join function to investigate this issue and determine what is going on here.

```{r mysterious-join, include = FALSE}
dest_delays %>% 
  anti_join(airports, by = c("dest" = "faa"))
```

<!-- SJU did not have airports information.  -->

(3) Is there a relationship between the age of a plane and its delays? The plane tail number is given in the `tailnum` variable in the `flights` dataset. The year the plane was manufactured is given in the `year` variable in the `planes` dataset. 

    - Step #1: Start by finding the average arrival delay for each plane and store the resulting dataset in `plane_delays`.

```{r step-1, include = FALSE}
plane_delays <- flights %>% 
  group_by(tailnum) %>% 
  summarize(median_arr_delay = median(arr_delay, na.rm = TRUE)) 
```

    - Step #2: Join `plane_delays` to the `planes` data using an appropriate join and then use `mutate` to create an age variable. 

```{r step-2, include = FALSE}
plane_delays <- plane_delays %>% 
  left_join(planes, by = "tailnum") %>% 
  mutate(plane_age = 2021 - year)
```

    - Step #3: Finally, create an effective visualization of the data.

```{r step-3, include = FALSE}
plane_delays %>% 
  ggplot(aes(plane_age, median_arr_delay)) + 
  geom_point() +
  geom_smooth() +
  labs(x = "Plane age as of 2021", y = "Median arrival delay")
```

## Submitting Application Exercises

- Once you have completed the activity, push your final changes to your GitHub repo. 
- Make sure you committed at least *three* times. 
- Check that your repo is updated on GitHub, and that’s all you need to do to submit application exercises for participation.

### Additional Resources

- https://rstudio.com/wp-content/uploads/2015/02/data-wrangling-cheatsheet.pdf