---
title: "AE 17: Central Limit Theorem"
subtitle: "due Tuesday, June 7 at 9:29am"
author: "Bora Jin"
editor_options:
  chunk_output_type: console
output: 
  html_document: 
    toc: true
    toc_float: true
---

## Getting Started 

- Clone the repository entitled "ae17-GitHubUsername" at [course GitHub organization page](https://github.com/sta199-summer22) on your [RStudio](https://vm-manage.oit.duke.edu/containers).

- Open the `.Rmd` file and replace "Your Name" with your name. 

## Bone Density for Women Over 65

```{r load-package, message=F}
library(tidyverse)
theme_set(theme_bw())
```

Suppose the bone density for 65-year-old women is normally 
distributed with mean $809 ~mg/cm^3$ and standard deviation of $140 ~mg/cm^3$.

Let $X$ be the bone density of 65-year-old women. We can write this distribution of $X$ in mathematical notation as 

$$X \sim N(809, 140)$$

### Part 1: Population distribution

The population distribution is given below: 

```{r visualize-normal}
ggplot(data = data.frame(x = c(809 - 140*3, 809 + 140*3)), 
       aes(x)) + 
  stat_function(fun = dnorm, args = list(mean = 809, sd = 140)) + 
  labs(title = "Population distribution", x = "", y = "Density") 
```

**Q - ** Before typing any code, based on what you know about the normal distribution, what do you expect the median bone density to be?

**Q - ** What bone densities correspond to Q1 (25th percentile), Q2 (50th percentile), and Q3 (the 75th percentile) of this distribution? Use the `qnorm()` function to calculate these values. 

```{r}

```

The densities of three woods are below:

- Plywood: $540 ~mg/cm^3$
- Pine: $600 ~mg/cm^3$
- Mahogany: $710 ~mg/cm^3$

**Q - ** What is the probability that a randomly selected 65-year-old woman has bones less dense than Pine?

```{r}

```

**Q - ** Would you be surprised if a randomly selected 65-year-old woman had bone density less than Mahogany? What if she had bone density less than Plywood? Use the respective probabilities to support your response. 

```{r}

```

### Part 2: Sample mean and its distribution 

Suppose you want to analyze the mean bone density for a group of 10 randomly selected 65-year-old women. 

**Q - ** Before moving forward, let's check the conditions required to apply the Central Limit Theorem to define the distribution of $\bar{X}$.

  - Independence? 
  - Sample size / distribution? 
  
**Q - ** What is the shape, center, and spread of the distribution of $\bar{X}$, the mean bone density for a group of 10 randomly selected 65-year-old women?

**Q - ** Write the distribution of $\bar{X}$ using mathematical notation. 

The distribution of sample means is given below in comparison with the population distribution: 

```{r visualize-normal2}
ggplot(data = data.frame(x = c(809 - 140*3, 809 + 140*3)), 
       aes(x)) + 
  stat_function(fun = dnorm, args = list(mean = 809, sd = 140), 
            aes(color = "population", linetype = "population")) +
  stat_function(fun = dnorm, args = list(mean = 809, sd = 140/sqrt(10)), 
            aes(color = "sample mean", linetype = "sample mean")) +
  labs(x = "", y = "Density", linetype = "", color = "") +
  scale_color_manual(values = c("population" = "black", 
                                "sample mean" = "red")) +
  scale_linetype_manual(values = c("population" = "solid", 
                                   "sample mean" = "dashed")) +
  geom_vline(xintercept = 809, color = "gray", size = 1) +
  theme(legend.position = c(0.8, 0.8))
```

**Q - ** What is the probability that the mean bone density for a group of 10 randomly-selected 65-year-old women is less dense than Pine?

```{r}

```

**Q - ** Would you be surprised if a group of 10 randomly-selected 65-year old women had a mean bone density less than Mahogany? What the group had a mean bone density less than Plywood? Use the respective probabilities to support your response. 

**Q - ** Explain how and why your answers differ in Part 1 and Part 2. 

### Practice 

Let $X$ follow a Gamma distribution with the shape parameter $\alpha = 1$ and the rate parameter $\beta = 0.1$. Then its distribution looks like this: 

```{r visualize-gamma}
ggplot(data = data.frame(x = c(0,30)), aes(x)) +
  stat_function(fun = dgamma, args = list(shape = 1, rate = .1), 
                color = "black") +
  labs(x = "", y = "Density") 
```

This is far from any normal distributions. Theoretically, its mean is $\alpha/\beta = 10$ and standard deviation is $\sqrt{\alpha/\beta^2} = 10$.

Suppose you take 40 random observations of $X$, take the mean, and save it as $\bar{X}$. 

**Q - ** Can we apply the CLT? Does this scenario satisfy the following conditions? 

- Independence? 
- Sample size / distribution? 

**Q - ** If you said yes from the previous question, write the distribution of $\bar{X}$ by CLT using mathematical notation. 

**Q - ** Visualize the distribution of the sample mean by completing the code chunk below:

```{r visualize-gamma2, eval = FALSE}
ggplot(data = data.frame(x = c(0,30)), aes(x)) +
  stat_function(fun = dgamma, args = list(shape = 1, rate = 0.1), 
                color = "black", linetype = "solid") +
  stat_function(fun = dnorm, args = list(mean = __, sd = __), 
                color = "red", linetype = "dashed") +
  labs(x = "", y = "Density") 
```

## Submitting Application Exercises

- Once you have completed the activity, push your final changes to your GitHub repo. 
- Make sure you committed at least *three* times. 
- Check that your repo is updated on GitHub, and that’s all you need to do to submit application exercises for participation.

