---
title: "AE 19: Introduction to Linear Models"
subtitle: "due Thursday, June 9 at 9:29am"
author: "Bora Jin"
editor_options:
  chunk_output_type: console
output: 
  html_document: 
    toc: true
    toc_float: true
---

## Getting Started 

- Clone the repository entitled "ae19-GitHubUsername" at [course GitHub organization page](https://github.com/sta199-summer22) on your [RStudio](https://vm-manage.oit.duke.edu/containers).

- Open the `.Rmd` file and replace "Your Name" with your name. 

## Used Cars Price 

```{r load-package, message = F}
library(tidyverse)
theme_set(theme_bw())
library(tidymodels)
```

The dataset `car_prices.csv` contains attributes of cars offered for sale on 
cars.com in 2017^[The data is from the `Stat2Data` R package.]. The codebook is 
available below:

- `type`: Model (Accord, Maxima, Mazda6)
- `age`: Age of the used car (in years)
- `price`: Price (in thousands of dollars)
- `mileage`: Previous miles driven (in thousands of miles)

```{r load-data, message = FALSE}
car_prices <- read_csv("data/car_prices.csv")
glimpse(car_prices)
```

### Part 1: Price and Mileage

Consider a regression model with the response `price` and a single predictor `mileage`. 

**Q - ** Write out the equation of a model using parameters and variable names. 

**Q - ** Create a scatterplot of price and mileage.

```{r price-mileage}

```

**Q - ** Use appropriate functions to find the fitted model and display the results in `tidy` format. 

```{r model1fitting, eval = FALSE}
## option 1
linear_reg(engine = "lm") %>% 
  fit(____ ~ _____, data = ______) %>% 
  tidy()

## option 2
lm_pm <- lm(____ ~ _____, data = ______) %>% 
  tidy()
lm_pm
```

- `linear_reg(engine = "lm")`: specify which regression model to use ("lm" = linear model) 
- `fit`: fit, i.e., estimate parameters for a given model. `y ~ x`. use variable names in `data` argument. 
- `tidy`: construct a tidy data frame summarizing model results 
- `lm`: `R` base function to fit a linear model. same formula syntax as `fit`

**Q - ** Write out the equation of the fitted model, and interpret the slope and 
intercept in the context of data. 

- Intercept: 
- Slope: 

**Q - ** What is the predicted selling price of a car with 50,000 miles?

```{r pred}

```

**Q - ** Include a visualization of the linear model on the scatterplot we created above. 
Try two options provided below and compare the two visualizations focusing on any difference. 

```{r price-mileage2, eval = FALSE}
## option 1: intercept and slope

## option 2: predicted values
car_prices <- car_prices %>% 
  mutate(pred = ____)

car_prices %>% 
  ggplot() + 
  geom_point(aes(x = mileage, y = price)) +
  geom_line(aes(x = _____, y = _____), 
              size = 1.5, color = "red") + 
  labs(x = "Mileage (in thousands of miles)", 
       y = "Price (in thousands of dollars)")
```


**Q - ** Suppose my friend has a Honda Accord with 225,000 miles. 
Suppose another friend has a BMW car with 80,000 miles. Is it appropriate 
to use this model to make a prediction for the selling prices? Why or why not?

### Part 2: Price and Type

Consider a regression model with the response `price` and the categorical predictor `type` (Accord, Maxima, Mazda6).

**Q - ** Create side-by-side boxplots of price for each type.

```{r price-type}
```

**Q - ** Use appropriate functions to find the fitted model and display the results in `tidy` format. 
Write out the equation of the fitted model.

```{r model2fitting}
## option 1

## option 2

```


**Q - ** How many terms are in the model for `type`? Is this equal to the number 
of car types in the data set? If not, briefly explain why the number of terms for 
`type` in the model differs from the number car types in the data set.

**Q - ** Interpret the intercept and slope(s) in the context of the problem.

- Intercept: 
- Slope: 

### Part 3: Towards More Complex Models 

**Q - ** Create a scatterplot of price and age. Comment on what you observe. 

```{r price-age}

```

**Q - ** Add fitted linear lines for each `type` of cars on top of the scatterplot of price and age.
Comment on what you observe. 

```{r summary, eval = FALSE}
car_prices %>% 
  ggplot(aes(x = ___, y = ___, color = ___)) + 
  geom_point() +
  geom_XXXX(method = ___, se = FALSE) + 
  labs(x = "Car age (in years)", 
       y = "Price (in thousands of dollars)", 
       color = "Model") +
  scale_color_viridis_d()
```

**Q - ** What are possible limitations of two regression models in Part 1 and 2?

## Submitting Application Exercises

- Once you have completed the activity, push your final changes to your GitHub repo. 
- Make sure you committed at least *three* times. 
- Check that your repo is updated on GitHub, and that’s all you need to do to submit application exercises for participation.

