---
title: "Regression 3: Interactions, model selection and transformations"
date: "March 25 2022"
output:
  html_document:
    df_print: paged
---

```{r, warning = FALSE, message  = FALSE}
library(tidyverse)
library(tidymodels)
```

# Bulletin

- Due Monday
  - Lab 08
  - Draft project report
    - make sure your GitHub repo contains `report.rmd` and `report.pdf`

# Today

By the end of today, you will...

- model interactions between predictors
- understand what's linear about linear regression
- report and explain adjusted $R^2$

We'll continue examining the Palmer penguin dataset.

```{r load-penguins, warning = F, message = F}
data(penguins)
```

Use `?penguins` or [click here](https://github.com/allisonhorst/palmerpenguins) for more info about the dataset.

# Interactions

```{r interaction-plot-1, warning=F, message = F}
penguins %>%
  ggplot(aes(x = bill_length_mm, y = body_mass_g, color = island)) +
  geom_point() + 
  theme_bw() + 
  geom_smooth(method = 'lm', se = F) + 
  labs(x = "Bill length (mm)", y = "Body mass (g)", title = "Body mass vs bill length by island", color = "Island")
```

- What's going on in the figure above?


## The main effects model


### Introducing categorical predictors

```{r main-effect}
main_fit = linear_reg() %>%
set_engine("lm") %>%
fit(body_mass_g ~ bill_length_mm + island, data = penguins)

main_fit %>%
  tidy()
```

The associated linear model:

$$
y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \beta_3 x_3
$$

- $y$: body mass (g)
- $x_1$: bill length (mm)
- $x_2$: whether or not the penguin is from the Dream island, binary (0 or 1)
- $x_3$: whether or not the penguin is from the Torgersen island, binary (0 or 1)

What about Biscoe island?

## The interaction effects model

```{r interaction-effect}
interaction_fit = linear_reg() %>%
set_engine("lm") %>%
fit(body_mass_g ~ bill_length_mm * island, data = penguins)

interaction_fit %>%
  tidy()
```

## Exercise 1

- What changed in the code?

- What does the full model look like?

- Interpret the `bill_length_mm:islandDream` slope.

## Exercise 2

- Before examining the numeric value, which model do you think has the larger $R^2$? Why?

- Compare **adjusted** $R^2$ between models. Which model do you prefer?


```{r compare-adj-r2}
# code here
```


# Adjusted $R^2$

More predictors, means higher $R^2$. For this reason, use adjusted $R^2$ when comparing models with variable number of predictors.

Adjusted $R^2$ penalizes the number of predictors in the model. Therefore, adjusted $R^2$ decreases unless the new variable helps explain the response.

Adjusted $R^2$:

$$
1 - (1 - R^2) \frac{n-1}{n - k - 1}
$$

where $n$ is the number of observations (in the data) and $k$ is the number of predictors (in the model).


# Linearity in linear regression

- Is the model from exercise 1 a linear model?

What's linear about linear regression? The coefficients. We can transform the data in any way we like.

## Example

```{r load-example-data, warning = F, message = F}
example = read_csv("data/example.csv")
```


```{r ex-linear, warning = F, message = F}
example %>%
  ggplot(aes(x = x, y = y)) + 
  geom_point() +
  theme_bw() + 
  geom_smooth(method = 'lm', se = FALSE, color = 'steelblue') +
  labs(x = "X", y = "Y", title = "Naive linear model is a bad fit")
```

## Exercise 3

- Write the equation of the line above (by finding the fitted model). What is the $R^2$ associated with the line above?

```{r bad-fit}
# code here
```


## Exercise 4

- What function better describes the relationship between $x$ and $y$ above?

- Transform the predictor $x$ and plot y versus the transformed predictor below.

```{r transformed-plot}
# code here
```

- Fit a linear model for the transformed $x$. Find $R^2$ of the new model and compare.

# Practice


## Exercise 5

Is the relationship between Body mass (g) and Bill depth (mm) positive or negative? Create a convincing argument from the data.


## Exercise 6
- Create a linear model of body mass using bill depth and one other predictor of your choosing.

- Do you prefer this model to the interaction effects model from exercise 2? Why?



