---
title: "Intro to R, RStudio and R Markdown"
output: html_document
author: "Dr. Alexander Fisher"
date: "08.26.2021"
editor_options: 
  chunk_output_type: console
---

### Logistics

<ul>

-   Questions?
-   Navigating the course website: slides go with videos, application exercises go with lecture.
-   Emails: include in title 'STA 199'
-   Professor Fisher's office hours Friday August 28 10:00am-12:00pm on zoom
    -   <https://duke.zoom.us/j/96917570315>
    -   requires Duke email to join\
    -   full office hours posted next week
-   First homework will be assigned next week
-   Due next Tuesday:
    -   quiz 01 on Sakai (by class time)
    -   lab 01 (by midnight)
    -   prep (watch videos) before class

</ul>

### R language

Print a number to the screen by typing the following into the console

```{r}
3
```

Use R like a calculator

    3 + 5 + 10
    3 * 3
    60 / 5
    11^2
    log(2)

**Store** the output of your calculation by assigning it a name. The named quantity is termed a "variable". A variable is a container that stores data values.

\<- is R's version of the equals sign and denotes the direction of assignment.

    x <- 3 + 5
    a <- 2
    199 -> sta

What will typing `sta` output?

---

Parentheses, as in `log()` above, denotes a function

    2.71 ^ 2
    exp(2)

    squareIt <- function(x) {
      return (x^2)
    }

Can you tell what `squareIt(5)` will return?

Some functions, such as `log()` and `exp()` are pre-defined in base R.

Additional functions may be loaded via packages.

A package is a collection of **data** and **functions**.

Load a package:

```{r}
library(tidyverse)
```

Conflicts may occur when you're loading a package that contains functions that are already defined.

### concept checklist: R

Ask yourself...

-   What's a variable?
-   What's a function?
-   What does it mean to 'print to screen'?
-   What's a package?
-   What's a data frame? How do you access a column?

### RStudio

Let's switch over and explore... [Duke R Containers](https://vm-manage.oit.duke.edu/containers)

### concept checklist: RStudio

-   What's the console?
-   What does the environment tell you?
-   How do you look up documentation?
-   What does it mean to "knit to ... pdf, html?"

**Pro tip: Code --\> Reformat Code**

### R Markdown

This file contains R markdown. Let's look at the [R Markdown cheatsheet](https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf) for quick guide to literate programming.

- Demo: visual editor in the corner (coming soon to containers).

```{r}
as.data.frame(iris)
```

```{r}
as_tibble(iris)
```

### Application Exercise
  Run the following code in the console to download the data and Rmd file for today’s application exercise. (Note: You will learn how to clone a GitHub repo in tomorrow’s lab.) Your RStudio may refresh, and you will see the R Markdown files and data set.
  
  ```
  usethis::use_course("sta199-f21/ae-02")
  ```

### Submission
There is nothing to submit for today’s AE. Starting Monday, you will submit push your AE to your GitHub repo.
