---
title: 'Homework #03: The Joy of Probability'
subtitle: 'due February 25th 11:59 PM '
output:
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    # css: "sta199-hw.css"
link-citations: yes
editor_options:
  chunk_output_type: inline
  markdown: 
    wrap: 85
---

## Goals

- Gain proficiency calculating marginal, joint, and conditional probabilities
- Use visualizations and probabilities to analyze multivariate associations between categorical variables

# Setup

- Go to the [Github Organization page](https://github.com/sta199-s22) and open your `hw03-username` repo
  
- Clone the repository, open a new project in RStudio.

# Data name

We will work with the `tidyverse` package as usual. Our data comes from the `fivethirtyeight` package. You may also want to use `viridis.`

```{r load-package, message = FALSE, warning = FALSE}
library(tidyverse)
library(fivethirtyeight)
library(viridis)
```

Bob Ross was a painter who was most famous for his PBS television show The Joy of Painting. In each episode, Ross created a new oil painting and provided instructions and commentary as he painted it. Ambitious viewers could paint along but viewers also simply enjoyed watching and listening to Ross's soothing voice as he painted an outdoor scene in 30 minutes.

In 2014, Walt Hickey wrote an article for FiveThirtyEight using statistics to analyze the paintings created on the show.The article focused on features that were often seen in Ross’s paintings, such as trees, clouds, cabins, among others. [Click here to see the article](https://fivethirtyeight.com/features/a-statistical-analysis-of-the-work-of-bob-ross/)

In this assignment, you will analyze the data that was used for the article. The data is in the `bob_ross` data set in the fivethirtyeight R package. Each observation represents an episode of the TV show. One painting was created in an episode. To access the full codebook of variables, explore the documentation using `?bob_ross`.

We'll focus on the following variables in this assignment:

- `tree`: Whether or not the painting contains at least 1 tree
- `guest`: Whether or not the episode featured a guest painter
- `steve_ross`: Whether or not Steve Ross was the featured guest
- `cirrus`: Whether the painting contains cirrus clouds
- `cumulus`: Whether the painting contains cumulus clouds
- `mountain`: Whether the painting contains a mountain
- `river`: Whether the painting contains a river
- `cabin`: Whether the painting contains a cabin
- `lake`: Whether the painting contains a lake

## Exercise 1 

- How many episodes are there of the show: **Joy of Painting**?

"There’s nothing wrong with having a tree as a friend."

- In how many episodes was a tree painted?

- What is the probability a randomly selected episode featured a tree?

## Exercise 2

The **Joy of Painting** occasionally featured a guest painter other than Bob Ross. One guest painter was Bob's son Steve Ross.

- What's the probability the show featured Steve Ross given there was a guest painter?

- Did Steve Ross like to paint mountains more or less than the other guest painters? Create a stacked bar plot of guest painters. Have whether or not Steve was the guest painter on the x-axis and fill the bars according to whether or not a mountain exists in the painting. Note: rename observations so that they are more informative than 0 and 1.

## Exercise 3

The next few questions will focus only on paintings created by Bob Ross. Make a new data frame called `ross_paintings` that only includes episodes (and thus paintings) made by Bob Ross. Save this data frame and use it for exercises 4 - 6.

## Exercise 4

"Let's build us a happy, little cloud that floats around the sky."

Are the following two events disjoint? Why or why not?

- A: Bob Ross paints a `cirrus` cloud
- B: Bob Ross paints a `cumulus` cloud

## Exercise 5

In the FiveThirtyEight article, Walt Hickey calculates various probabilities to describe the combination of features typically found in Bob Ross paintings. He states the following about the presence of cabins and lakes in Ross’s paintings: "About 18 percent of his paintings feature a cabin. Given that Ross painted a cabin, there’s a 35 percent chance that it’s on a lake…"

- How many of Bob Ross's paintings feature a cabin? Call this number "M"

- How many of those that feature a cabin also feature a lake? Call this number "X" for the next question.

- Imagine the following: every time Bob Ross paints a cabin, he flips a fair coin to decide whether or not to paint a lake. Given a collection of M Bob Ross paintings with cabins, what is the probability Bob Ross painted X or fewer lakes?

Hint: use the code below as a template

```{r simulation-code, eval = F}
set.seed(2182022) # don't change the seed
num_lakes = rbinom(100000, M, prob = ?)
cabin_lakes = data.frame(num_lakes)
```

## Exercise 6

Suppose you randomly select a Bob Ross painting and see that it features a mountain. Use Bayes Theorem to calculate the probability this painting also features a river. Show your work by using a code chunk as a calculator.

Hint: p(mountain | river) = 0.39

- Follow up question: Does Bob Ross paint mountains independent of whether or not he paints rivers? Why? In other words, is event A independent of B? Here we define events:

- A: Bob Ross paints a mountain
- B: Bob Ross paints a river

## Exercise 7

Your turn! Use this data to explore a question of your choice about paintings created in the TV show The Joy of Painting. Your question should explore the relationship between 3 variables in the data set; at least one of the variables must be one that hasn't been used in exercises 1 - 6. You may use the entire data set or focus the analysis on paintings made by Bob Ross.

- Clearly state the question you’re exploring.
- Create 1 - 2 visualizations that can be used to explore the question. Customize the colors of your visualization using some of the colors commonly used in Bob Ross paintings.[Click here](https://sites.radford.edu/~cdc/seansevillair.html) for a list of common Bob Ross hex codes.

Hint: [Click here for functions to manually create color palettes in ggplot2.](https://ggplot2.tidyverse.org/reference/scale_manual.html)
    
- Calculate the relevant probabilities needed to answer your question.
- Write a short narrative, about 2 - 4 sentences, answering the research question based on the visualization and probabilities. Note: The narrative should not exceed 4 sentences.



# Submission
Knit to PDF to create a PDF document. Stage and commit all remaining changes, 
and push your work to GitHub. Make sure all files are updated on your GitHub 
repo.
Only upload your PDF document to Gradescope. Before you submit the 
uploaded document, mark where each answer is to the exercises. If any answer 
spans multiple pages, then mark all corresponding pages. Associate the "Overall" section with the first page.

## Reminder:
- All plots should follow the best visualization practices: include an informed title, label axes, and carefully consider aesthetic choices.
- All code should follow the tidyverse style guidelines, including not exceeding the 80 character limit.

## Rubric
- Ex 1: 5 pts.
- Ex 2: 8 pts.
- Ex 3: 3 pts.
- Ex 4: 5 pts.
- Ex 5: 8 pts.
- Ex 6: 6 pts.
- Ex 7: 9 pts.
- Workflow and formatting - 6 pts
 - This includes having three meaningful commits, updating name, using
 tidyverse style and naming all code chunks.

