---
title: "Data Visualization"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
library(magick)
```


## Today's Goal 

- Comfortably navigate RStudio fundamentals
- Be able to define common terms including: **data frame**, **function**, **variable**, **package**
- Interact with Git from RStudio (committing and pushing)
- Use a few functions, including: `glimpse()`, `count()`, `group_by()` and `summarize()`
- Create a reproducible report via today's AE using R Markdown

---

## Quiz

- dataset, observation, variable 
- **single entry per observation variable** vs. list (multiple entries per observation)
- `glimpse()`, `nrow()`, `ncol()`, `dim()`
- what is EDA? 
- before plotting, what package to load? `library(tidyverse)`
- ggplot syntax: ggplot, data, aes, geom_xxxx, some more asethetics options (color, shape, size, alpha)
- why is visualization important? reveal a new story that summary statistics cannot tell. 
- mapping vs. setting 
- faceting to explore conditional relationships 
- facet_wrap and facet_grid 
- usually row 1 and column 2: row~column
- guides 

---

class: middle, center

# Questions?

---

## Let's Practice Together! 

Go to [AE 02: Tour of RStudio + R Markdown + GitHub](https://sta199-summer22.netlify.app/appex/ae02_BJ.html)

---

## For Tomorrow 

- Watch videos for [Prepare: May 13](https://sta199-summer22.netlify.app/prepare/week01_may13_BJ.html)

- Submit your `AE02`

