---
title: "Data Visualization 1"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Material 

`r emo::ji("movie_camera")` Watch [Data and Visualization](https://www.youtube.com/watch?v=FddF4b_GuTI) 

- [Slides](https://rstudio-education.github.io/datascience-box/course-materials/slides/u2-d01-data-viz/u2-d01-data-viz.html#1)

`r emo::ji("movie_camera")` Watch [Visualizing Data with ggplot2](https://www.youtube.com/watch?v=s2NF2J36ljE)  

- [Slides](https://rstudio-education.github.io/datascience-box/course-materials/slides/u2-d02-ggplot2/u2-d02-ggplot2.html#1)

---

## Today's Goal 

- Build up an effective visualization systematically layer by layer with `ggplot2`

"The simple graph has brought more information to the data analyst's mind than any other device" - John Tukey

---

## Quiz

**Q - What does each row of a dataset represent?**  

--

Observation 

--

**Q - What does each column of a dataset represent?** 

--

Variable 

--

**Q - All variables in a dataset must be in the same type (character, integer, etc.). (T/F)** 

--

F

--


**Q - In this class, we mainly focus on variables having a single entry per observation. (T/F)** 

--

T

---

## Quiz

**Q - Which function spits out the dimension of a dataset?** 

--

`dim()` (row, column)

--


**Q - How to check the number of rows of a dataset?** 

--

`nrow()`

--

**Q - How to check the number of cols of a dataset?** 

--

`ncol()`

---

## Quiz

**Q - What is the information that you cannot get using `glimpse()`?** 

a. number of rows

b. number of cols

c. list of variables 

d. variable types 

e. summaries of variables 

---

## Quiz

**Q - What is the information that you cannot get using `glimpse()`?** 

a. number of rows

b. number of cols

c. list of variables 

d. variable types 

**e. summaries of variables**

---

## Quiz

**Q - What is EDA?** 

--

- Approach to analysing datasets to summarize and describe its main characteristics
- Often visual 

--

**Q - Why is visualization important?** 

--

It can reveal a new story that summary statistics could miss! 

--

**Q - Before plotting, what package do we load?** 

--

`library(tidyverse)` (`ggplot2` $\in$ `tidyverse`)

---

## Quiz 

**Q - When plotting, what are the three fundamental things to define and the relevant code?** 

--

- Data in `ggplot(data = )`
- Aesthetics in `ggplot(mapping = aes())`
- Geometries in `geom_xxxx()`

layers added by `+`

--

```{r eval=F}
ggplot(data = [dataset], 
       mapping = aes(x = [x-variable], y = [y-variable])) +
   geom_xxx() +
   other options
```

---

## Quiz

**Q - What are *other options* include?** 

--

- Labels and legends in `labs()`
- More aesthetics options (color, shape, size, alpha)
- Faceting
- Many more! 

--

**Q - For aesthetics options, can you explain mapping vs. setting?** 

--

- Mapping: map plotting characters to a specific variable in data through `aes()`
- Setting: determine them without regard to data values through `geom_xxxx()`

---

## Quiz

**Q - Why faceting?** 

--

- To explore conditional relationships 
- Multiple smaller plots in one plot 

--

**Q - Two functions for faceting?** 

--
- `facet_grid([row-variable] ~ [column-variable])`
- `facet_wrap(~ [variable])`

---

class: middle, center

# Questions?

---

## Let's Practice Together! 

Go to [AE 03: Data Visualization 1](https://sta199-summer22.netlify.app/appex/ae03_BJ.html)

---

## For Monday 

- Watch videos for [Prepare: May 16](https://sta199-summer22.netlify.app/prepare/week02_may16_BJ.html)

- Submit your `AE03`

