---
title: "Statistical Inference"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Material 

`r emo::ji("movie_camera")` Watch [Scientific Studies and Confounding](https://youtu.be/WnMzTBrZDcc)

- [Slides](https://rstudio-education.github.io/datascience-box/course-materials/slides/u2-d15-studies-confounding/u2-d15-studies-confounding.html#1)

---

## Today's Goal 

- Understand statistical process terminology
- Understand the types of conclusions that can be made given the underlying statistical process
- Learn to simulate from a binomial distribution

---

## Quiz 

**Q - What is an observational study? Any examples?** 

--

- Collect data in a way that does not interfere with how the data arise 
- Establish **associations** (!= causation)
- Data often cheaper and easier to collect 
- e.g., Smoking behavior and health outcomes 

---

## Quiz 

**Q - What is an experimental study? Any examples?** 

--

- Randomly assign subjects to treatments vs. control 
- Establish **causal connections**
- Often more expensive 
- Sometimes impossible or unethical to design an experiment 
- e.g., Covid-19 vaccine tests in developmental phase 

---

## Quiz 

In the prep video, the following terms are introduced: <br>
response variables and explanatory variables

**Q - What is a response variable?** 

--

- "Dependent variable" 
- Denoted by $y$
- The focus of a research question

--

**Q - What is an explanatory variable?**

--

- "Independent variable" 
- Denoted by $x$
- Variables that might affect or explain changes in the response variable 

---

## Quiz 

**Q - Identify a response variable and an explanatory variable regarding the following examples.**

Do larger homes in good locations lead to higher home selling prices? 

--

- $y$: home selling price
- $x$: size of a house, neighborhood

--

Which procedure prolongs life more for breast cancer patients, chemo or anti-estrogen treatment? 

--

- $y$: survival time 
- $x$: type of treatment
- In real life, you would include several more explanatory variables such as age, general health conditions, weight, etc. 
---

## Quiz 

**Q - What is a confounding variable?**

--

- Extraneous variable that affects both the explanatory and the response variable, and that makes it seem like there is a relationship between them.
- If you fail to account for confounding variables, you may not correctly observe the actual relationship between explanatory variables and a response variable.

---

## Quiz 

**Q - Identify the confounding variable in each of the following statements.**

As the amount of ice cream sales increases, the number of shark attacks also increases.

--

- Confounding variable: summer
- Both increase in the summer, when people go to the beach.

--

The higher the number of firefighters at a fire is, the greater the amount of damage caused by that fire.

--

- Confounding variable: size of a fire 
- More firefighters are sent to larger and severer fires that cause greater damage. 

---

## Quiz 

In the prep video, the following terms are introduced: <br>
random sampling and non-random sampling 

**Q - What is random sampling?** 

--

- A sampling method that allows for randomization of sample selection. 
- Conclusions from a random sample are **generalizable**.
- Simple random sampling: every element in a whole population has an equal chance of being chosen for the sample
<!-- - Stratified random sampling: it divides the population into subgroups or strata such that each subgroup is homogeneous. Then simple random sampling is applied within each stratum.  -->
<!--   - Useful when you need to make sure you get a sufficient sample from each demographic group and then you weight to the approximate percentage in the population. -->

---

## Quiz 

**Q - What is non-random sampling?** 

- In a non-random sample, individuals are selected based on non-random criteria, and not every individual has a chance of being included.
- Conclusions from a non-random sample are **not generalizable**. 
- e.g., I make my close friends complete a survey and analyze the survey data for a final project.
- e.g.. An online survey available to anyone who wishes to provide responses 

---

class: middle, center

# Questions?

---

## Let's Practice Together! 

Go to [AE 12: Foundations of Inference](https://sta199-summer22.netlify.app/appex/ae12_BJ.html)

---

## Bulletin

- Exam 01 

  - to be released today at 1pm
  - due Monday, May 30 at 12:59pm 
  
- May 30 

  - Memorial day holiday. No class! 
  - Have fun after submitting Exam 01 



