---
title: "Confidence Intervals through Bootstrapping"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Material 

`r emo::ji("movie_camera")` Watch [Bootstrapping](https://youtu.be/bdqpI3iVOso)

- [Slides](https://rstudio-education.github.io/datascience-box/course-materials/slides/u4-d11-bootstrap/u4-d11-bootstrap.html#1) 

Optional: `r emo::ji("book")` Read [IMS: Chapter 12 - Confidence Intervals With Bootstrapping](https://openintro-ims.netlify.app/foundations-bootstrapping.html)

---

## Today's Goal 

- Understand how to draw a bootstrap sample and calculate a bootstrap statistic
- Use `infer` (part of `tidymodels`) to obtain a bootstrap distribution
- Calculate a confidence interval from the bootstrap distribution
- Interpret a confidence interval in context of the data

---

## Ideas Behind Bootstrapping   

--

- It is extremely hard to hit the exact population parameter with a single value (a.k.a. point estimate or sample statistic) from a sample.

--

- We would have a better chance with **a range of plausible values** for the population parameter. 

--

- One statistic from one sample... then we need multiple samples to obtain multiple statistics. 

--

- Uh oh, getting multiple samples is generally expensive and time-consuming. If it was 
so easy, we would simply collect data from every unit in the population. 

--

- We just have to make the best of it! With the sample at hand, let's pretend 
that we have multiple samples!! $\Rightarrow$ Bootstrapping is a resampling technique. 

---

## Quiz 

**Q - Correct the following sentences.**

`r emo::ji("emergency")` A bootstrap sample is taken without replacement from the original sample. 

--

`r emo::ji("check")` A bootstrap sample is taken with replacement from the original sample. 

--

`r emo::ji("emergency")` A researcher chooses a bootstrap sample size.

--

`r emo::ji("check")` A bootstrap sample should be of the same size as the original sample.

---

## Quiz 

**Q - Correct the following sentences.**

`r emo::ji("emergency")` A bootstrap distribution is a distribution of resampled data. 

--

`r emo::ji("check")` A bootstrap distribution is a distribution of bootstrap statistics. 

--

`r emo::ji("emergency")` Calculate the bounds of the XX% confidence interval as 
the shortest interval with XX% of the bootstrap distribution. 

--

`r emo::ji("check")` Calculate the bounds of the XX% confidence interval as the middle XX% of the bootstrap distribution. 

---

## Quiz 

**Q - Correct the following sentences.**

`r emo::ji("emergency")` Bootstrapping can be used for mean and median, but not for 
standard deviation.  

--

`r emo::ji("check")` Bootstrapping can be used for any sample statistic. 

--

`r emo::ji("emergency")` The length of a confidence interval has nothing to do with its confidence level.

--

`r emo::ji("check")` The lower the confidence level, the narrower the interval. 

---

## Quiz 

**Q - Correct the following sentences.**

`r emo::ji("emergency")` Bootstrapping can improve statistical inference based on 
a bad sample. 

--

`r emo::ji("check")` Bootstrapping is meaningless if the original sample is not representative because your bootstrap distribution will be centered around the original sample statistic.  

- Bootstrapping is best suited for modeling studies where the data have been generated through **random sampling** from a population.

---

## Quiz 

The 95% confidence interval for the mean price per guest per night
among Airbnb rentals (with at least ten reviews) in Asheville was ($64, $90).

**Q - What is the correct interpretation for this interval?**

a. There is a 95% probability the mean price per night for an Airbnb in Asheville is between $64 and $90.

b. There is a 95% probability the price per night for an Airbnb in Asheville is between $64 and $90.

c. We are 95% confident the mean price per night for Airbnbs in Asheville in our sample is between $64 and $90.

d. We are 95% confident the mean price per night for all Airbnbs in Asheville is between $64 and $90.

---

## Quiz 

The 95% confidence interval for the mean price per guest per night
among Airbnb rentals (with at least ten reviews) in Asheville was ($64, $90).

**Q - What is the correct interpretation for this interval?**

a. There is a 95% probability the mean price per night for an Airbnb in Asheville is between $64 and $90.

b. There is a 95% probability the price per night for an Airbnb in Asheville is between $64 and $90.

c. We are 95% confident the mean price per night for Airbnbs in Asheville in our sample is between $64 and $90.

**d. We are 95% confident the mean price per night for all Airbnbs in Asheville is between $64 and $90.**

---

## Quiz 

a. "There is a 95% probability"

- The true unknown parameter is either in ($64, $90) or not. It can't have a "95% probability" of being in any specific interval. 

b. "the price per night for an Airbnb in Asheville"

- We did not infer on individual price; we did infer on the population mean price per night. 

c. "in our sample"

- I don't need to build a confidence interval for the sample statistic. It's known! 

---

## Quiz 

**Q - What does it mean that "we are 95% confident"?**

--

- Suppose we took 100 independent samples from a population. 
- For each sample, we computed the sample statistic and constructed a 95% confidence interval using bootstrapping. 
- Then we would expect 95 out of these 100 intervals to contain the unknown parameter. 

---

class: middle, center

# Questions?

---

## Let's Practice Together! 

Go to [AE 14: Confidence Intervals through Bootstrapping](https://sta199-summer22.netlify.app/appex/ae14_BJ.html)

---

## Bulletin

- Watch videos for [Prepare: June 2](https://sta199-summer22.netlify.app/prepare/week04_jun02_BJ.html)

- Project proposal due Friday, June 3 at 11:59pm 

- HW03 released 

- Submit `ae14`

