---
title: "Hypothesis Testing 1"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Material 

`r emo::ji("movie_camera")` Watch [Hypothesis Testing](https://warpwire.duke.edu/w/YQ4GAA/)

- [Slides](https://sta199-f21-001.netlify.app/slides/16-hypothesis-test.html#1)

Optional: `r emo::ji("book")` Read [IMS: Chapter 11 - Hypothesis Testing With Randomization](https://openintro-ims.netlify.app/foundations-randomization.html)

---

## Today's Goal 

- Understand important concepts in hypothesis testing framework including the null hypothesis, alternative hypothesis, and p-value, etc.
- Conduct simulation-based hypothesis testing with the `tidymodels` package 

---

## Statistical Inference

Statistical inference is the process of using sample data to make 
conclusions about the underlying population the sample came from.

- **Estimation**: using the sample to estimate a plausible range of values 
for the unknown parameter

- **Testing**: evaluating whether our observed sample provides evidence 
for or against some claim about the population

--

Today we will focus on **Testing**.

We will conduct hypothesis testing using simulation-based methods (**bootstrapping**, again). 

---

## Hypothesis testing framework 

1. Defining the hypotheses

2. Collecting and summarizing data

3. Assessing the observed evidence

4. Making a conclusion

---

## Quiz 

1. Defining the hypotheses

**Q - Choose the correct description in the following sentences:** 

- Two hypotheses are about the ( population / sample ). 

- The null and the alternative hypotheses are defined for ( statistics / parameters ).

- The null hypothesis states ( "there is nothing unusual going on" / "there is something interesting going on" ).

- The alternative hypothesis states ( the status quo / the research question ).

- The alternative hypothesis is denoted by ( $H_1$ / $H_A$ ). 

---

## Quiz 

1. Defining the hypotheses

**Q - Choose the correct description in the following sentences:** 

- Two hypotheses are about the ( **population** / sample ). 

- The null and the alternative hypotheses are defined for ( statistics / **parameters** ).

- The null hypothesis states ( **"there is nothing unusual going on"** / "there is something interesting going on" ).

- The alternative hypothesis states ( the status quo / **the research question** ).

- The alternative hypothesis is denoted by **( $H_1$ / $H_A$ )**. 

---

## Quiz 

1. Defining the hypotheses

**Q - Which of the following is the correct set of hypotheses?**

(a) $H_0: p = 0.10$; $H_A: p \ne 0.10$ 

(b) $H_0: p = 0.10$; $H_A: p > 0.10$ 

(c) $H_0: \hat{p} = 0.10$; $H_A: \hat{p} \ne 0.10$ 

(d) $H_0: \hat{p} = 0.10$; $H_A: \hat{p} < 0.10$ 

- $\hat{\theta}$ used to denote the associated statistic to the parameter $\theta$.

---

## Quiz 

1. Defining the hypotheses

**Q - Which of the following is the correct set of hypotheses?**

**(a) $H_0: p = 0.10$; $H_A: p \ne 0.10$** 

**(b) $H_0: p = 0.10$; $H_A: p > 0.10$** 

(c) $H_0: \hat{p} = 0.10$; $H_A: \hat{p} \ne 0.10$ 

(d) $H_0: \hat{p} = 0.10$; $H_A: \hat{p} < 0.10$ 

- $\hat{\theta}$ used to denote the associated statistic to the parameter $\theta$.

---

## Quiz 

1. Defining the hypotheses

**Q - Identify the null and alternative hypothesis in the following research questions.**

A principal at a certain school claims that the students in the school are above average intelligence. The mean population IQ is 100.

--

- $H_0$: The mean IQ for students attending the school is equal to the mean population IQ.
- $H_1$: The mean IQ for students attending the school is above the mean population IQ.

With $\mu$ being the mean IQ for students attending the school,

- $H_0$: $\mu = 100$
- $H_1$: $\mu > 100$

---

## Quiz 

1. Defining the hypotheses

**Q - Identify the null and alternative hypothesis in the following research questions.**

A researcher wants to test if vitamin C has the ability to prevent the flu in children. The flu infection rate in the US children population is 20%. 

--

- $H_0$: The true infection rate of the flu among children with sufficient vitamin C is equal to the infection rate among all US children. 
- $H_1$: The true infection rate of the flu among children with sufficient vitamin C is lower than the infection rate among all US children. 

With $p$ being the true infection rate of the flu among children with sufficient vitamin C,

- $H_0$: $p = 0.2$
- $H_1$: $p < 0.2$

---

## Quiz 

3. Assessing the observed evidence

**Q - What is p-value?** 

--

- Conditional probability 
- Given $H_0$ is true, what is the probability of observing $\hat{p}$ (our statistic) or something more extreme against the null hypothesis?
- We compute this probability by simulating a null distribution for $\hat{p}$. 

--

**Q - What is the null distribution?** 

--

- Distribution of the observed statistics given the null hypothesis is true ("under the null hypothesis")

--

**Q - We have only one sample. How can we possibly get a distribution?** 

--
Bootstrapping!

---

## Quiz 

4. Making a conclusion

**Q - What are the conclusions we can make from a hypothesis test?**

--

- Reject $H_0$ in favor of $H_1$
- Fail to reject $H_0$ 

--

**Q - We make a conclusion by comparing the p-value to a predetermined numeric cutoff. What is it called?**

--

- Significance level 
- Denoted by $\alpha$
- Depends on the context, but usually set at $\alpha = 0.05$

---

## Quiz 

4. Making a conclusion

**Q - What does it mean that $\alpha = 0.05$?** 

--

- We would expect to incorrectly reject $H_0$ when $H_0$ is true for 5% of the time. 
- P(reject $H_0$ | $H_0$ is true) = 0.05

--

**Q - State a conclusion to make when the p-value $< \alpha$.** 

--

- The results are statistically significant. 
- There is sufficient evidence at $\alpha = 0.05$ to reject the null hypothesis in favor of $H_1$. 

---

## Quiz 

4. Making a conclusion

**Q - State a conclusion to make when the p-value $\geq \alpha$.** 

--

- The results are not statistically significant. 
- We fail to reject the null hypothesis.
- There is insufficient evidence at $\alpha = 0.05$ to reject the null hypothesis. 

--

**Q - What are the two types of errors we can make?** 

--

- P(Type I error) = $\alpha$ = P(reject H_0 | H_0 is true)
- P(Type II error) = $\beta$ = P(fail to reject H_0 | H_0 is false)

--

**Q - How do we assess the capability of a test for detecting "something interesting"?** 

--

- The power of a test: $1-\beta$ = P(reject H_0 | H_0 is false)

---

class: middle, center

# Questions?

---

## Bulletin

- Watch videos for [Prepare: June 3](https://sta199-summer22.netlify.app/prepare/week04_jun03_BJ.html)

- Project proposal due Friday, June 3 at 11:59pm 

- HW03 released 


