---
title: "Central Limit Theorem 1"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

## Material 

`r emo::ji("movie_camera")` Watch [Central Limit Theorem](https://warpwire.duke.edu/w/Vw4GAA/)

- [Slides](https://sta199-fa21-003.netlify.app/slides/clt-intro.html#1)

Optional: `r emo::ji("book")` Read [IMS: Chapter 13 - Inference With Mathematical Models](https://openintro-ims.netlify.app/foundations-mathematical.html?q=central%20limit%20th#CLTsection)

---

## Today's Goal 

- Use Central Limit Theorem to define distribution of sample means
- Calculate probabilities from the normal distribution

---

## Quantifying Variability

We can quantify the variability of sample statistics using different approaches:

- **Simulation**: via bootstrapping or "resampling" techniques 

or

- **Theory**: via the Central Limit Theorem 

--

Today we will focus on **Theory**.

---

## Quiz 

**Q - What is a sampling distribution of the sample mean?**

---

## Quiz 

**Q - Apply the central limit theorem (CLT) on sample means.**

Let a random variable $X$ have a mean $\mu$ and standard deviation $\sigma$. 
Then the sampling distribution of the sample mean $\bar{X} = \frac{X_1 + X_2 + \cdots + X_n}{n}$ 

---

## Quiz 

**Q - Describe density of a normal distribution.** 

---

## Quiz 

**Q - The CLT holds only if** $X \sim N(\mu, \sigma)$. **(T/F)**

[Click!](https://onlinestatbook.com/stat_sim/sampling_dist/index.html)

---

## Quiz 

**Q - What are the two conditions for CLT to hold?** 

---

## Quiz 

**Q - What are the two conditions for CLT to hold?** 

---

## Recap

- If certain assumptions are satisfied, regardless of the shape of the 
population distribution, the sampling distribution of the mean follows an 
approximately normal distribution.

--

- The center of the sampling distribution is at the center of the population 
distribution.

--

- The sampling distribution is less variable than the population distribution 
by a factor of $1/\sqrt{n}$. 

--

- As $n$ increases, the standard error (the spread of the sampling distribution) decreases. 

---

## Quiz 

**Q - What is an appropriate code to calculate** $P(Z < 1.2)$ **where** $Z \sim N(0,1)$**?**

```{r echo = F, out.width = "50%"}
cutoff <- 1.2
ggplot(NULL, aes(c(-3,3))) +
  geom_area(stat = "function", fun = dnorm, 
            fill = "#00998a", xlim = c(-3, cutoff), color = "black") +
  geom_area(stat = "function", fun = dnorm, 
            fill = "#FFFFFF", xlim = c(cutoff, 3), color = "black") +
  labs(x = "Z", y = "Density") +
  scale_x_continuous(breaks = -3:3)
```

---

## Quiz 

**Q - What is an appropriate code to calculate** $P(-2 < X < 7)$ **where** $X \sim N(1,3)$**?**

```{r echo = F, out.width = "50%"}
cutoff1 <- -2
cutoff2 <- 7
ggplot(NULL, aes(c(-8,10))) +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 1, sd = 3), 
            fill = "#FFFFFF", xlim = c(-8, cutoff1), color = "black") +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 1, sd = 3), 
            fill = "#00998a", xlim = c(cutoff1, cutoff2), color = "black") +
  geom_area(stat = "function", fun = dnorm, args = list(mean = 1, sd = 3), 
            fill = "#FFFFFF", xlim = c(cutoff2, 10), color = "black") +
  labs(x = "X", y = "Density") +
  scale_x_continuous(breaks = seq(-8, 10, by = 2))
```

---

## Quiz 

**Q - What is an appropriate code to find** q **s.t.** $P(X > q) = 0.05$ **where** $X \sim N(-1,2)$**?**

```{r echo = F, out.width = "50%"}
cutoff <- qnorm(0.05, mean = -1, sd = 2, lower.tail = FALSE)
ggplot(NULL, aes(c(-7,6))) +
  geom_area(stat = "function", fun = dnorm, 
            args = list(mean = -1, sd = 2),
            fill = "#FFFFFF", xlim = c(-7, cutoff), color = "black") +
  geom_area(stat = "function", fun = dnorm, 
            args = list(mean = -1, sd = 2),
            fill = "#00998a", xlim = c(cutoff, 6), color = "black") +
  labs(x = "X", y = "Density") +
  scale_x_continuous(breaks = seq(-7, 6, by = 2)) +
  geom_point(data = data.frame(x = cutoff, y = 0), aes(x,y)) + 
  geom_text(aes(x = 1.8, y = 0.01, label = "q = ?"))
```

---

class: middle, center

# Questions?

---

## Let's Practice Together! 

Go to [AE 17: Central Limit Theorem](https://sta199-summer22.netlify.app/appex/ae17_BJ.html)

---

## Bulletin

- Watch videos for [Prepare: June 7](https://sta199-summer22.netlify.app/prepare/week05_jun07_BJ.html)

- Project proposal feedback released 

- HW03 due Wednesday, June 8 at 11:59pm 

