---
title: "Central Limit Theorem 2"
author: "Bora Jin"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
theme_set(theme_bw())
```

## Material 

`r emo::ji("movie_camera")` Watch [Inference Using Central Limit Theorem](https://warpwire.duke.edu/w/WQ4GAA/)

- [Slides](https://sta199-fa21-003.netlify.app/slides/clt-inference.html#1)

Optional: `r emo::ji("book")` Read 

- [IMS: Section 16.2 - Mathematical Model for a Proportion](https://openintro-ims.netlify.app/inference-one-prop.html#one-prop-norm)
- [IMS: Section 17.3 - Mathematical Model for Difference in Proportions](https://openintro-ims.netlify.app/inference-two-props.html#math-2prop)
- [IMS: Section 19.2 - Mathematical Model for a Mean](https://openintro-ims.netlify.app/inference-one-mean.html#one-mean-math)
- [IMS: Section 20.3 - Mathematical Model for Testing Difference in Means](https://openintro-ims.netlify.app/inference-two-means.html#math2samp)

---

## Today's Goal 

- Use Central Limit Theorem (CLT) to conduct inference on a population mean
- Conduct CLT-based inference step-by-step and using the `infer` package
- Understand t-distribution vs. standard normal, N(0,1) distribution

---

## Quiz 

**Q - State the central limit theorem.**

For a population with a well-defined mean $\mu$ and standard deviation $\sigma$,
these three properties hold for the distribution of sample average $\bar{X}$,
assuming certain conditions hold:

--

`r emo::ji("white_check_mark")` The distribution of the sample statistic is 

--
**approximately** normal 

--

`r emo::ji("white_check_mark")` The distribution is centered at 

--
the population parameter (often interest of inference)

--

`r emo::ji("white_check_mark")` The variability of the distribution is **inversely** proportional to the square root of 

--
the sample size

---

## Quiz 

**Q - Why do we care about the distribution of sample mean?** 

--

We can estimate / test for a population mean. 

We can construct a confidence interval or conduct a hypothesis test for the population mean using the CLT-based distribution in place of a simulation-based distribution of sample mean. 

---

## Quiz 

**Q - What is the distribution of sample mean by CLT?** 

When the population mean $\mu$ and the population standard deviation $\sigma$ are known, 

--

$$\bar{X} \sim N(\mu, \sigma/\sqrt{n}) \Leftrightarrow Z = \frac{\bar{X}-\mu}{\sigma/\sqrt{n}} \sim N(0,1)$$

approximately, for a large enough $n$. 

--

- N(0,1) is standard normal distribution.
- Often, a random variable following the standard normal distribution is denoted by $Z$.

---

## Quiz 

**Q - What if** $\sigma$ **is unknown?** 

--

- We approximate $\sigma$ with the sample standard deviation. 

--

$$Z = \frac{\bar{X}-\mu}{\sigma/\sqrt{n}} \rightarrow T = \frac{\bar{X}-\mu}{S/\sqrt{n}}$$
where $S^2 = \sum_{i=1}^n(X_i - \bar{X})^2/(n-1)$
--

- $\sigma$ replaced by $S$! (The realized value of $S$ from a sample is $s$.)
--

- This change renders the random variable $T$ follow another distribution than the standard normal distribution, i.e., $$T \sim t_{n-1}$$
where $t_{n-1}$ is a t-distribution with $n-1$ degrees of freedom. 

---

## Quiz 

**Q - List properties of the t-distribution.** 

- Its shape is 

--
unimodal, symmetric, centered at 0 similarly to N(0,1).

--

- Its tails are 

--
thicker than N(0,1).

--

- It is fully defined by 

--
the degrees of freedom.

---

## Quiz 

**Q - Black solid line is N(0,1). What is the t-distribution with df = 1, 3, 10, and 30?**

```{r echo=FALSE, out.width = "60%"}
x <- seq(-5, 5, length = 1000)
d <- rbind(
  data.frame(dist="t(df=1)",  x=x, d=dt(x,1)),
  data.frame(dist="t(df=3)",  x=x, d=dt(x,3)),
  data.frame(dist="t(df=10)", x=x, d=dt(x,10)),
  data.frame(dist="t(df=30)", x=x, d=dt(x,30))
)
ggplot(d) + 
  geom_line(data = data.frame(x=x, d=dnorm(x)), aes(x = x, y = d)) +
  geom_line(aes(x = x, y = d, color = dist, linetype = dist)) +
  labs(x = "", y = "Density") + 
  theme(legend.position = "none")
```

---

## Quiz 

**Q - Black solid line is N(0,1). What is the t-distribution with df = 1, 3, 10, and 30?**

```{r echo=FALSE, out.width = "55%"}
x <- seq(-5, 5, length = 1000)
d <- rbind(
  data.frame(dist="t(df=1)",  x=x, d=dt(x,1)),
  data.frame(dist="t(df=3)",  x=x, d=dt(x,3)),
  data.frame(dist="t(df=10)", x=x, d=dt(x,10)),
  data.frame(dist="t(df=30)", x=x, d=dt(x,30))
)
ggplot(d) + 
  geom_line(data = data.frame(x=x, d=dnorm(x)), aes(x = x, y = d)) +
  geom_line(aes(x = x, y = d, color = dist, linetype = dist)) +
  labs(x = "", y = "Density", 
       color = "Distribution", linetype = "Distribution") +
  theme(legend.position = c(0.8, 0.75), 
        legend.background = element_rect(fill = NA))
```
- thicker tails 
- As the degrees of freedom increases, the t-distribution becomes more like N(0,1)

---

## Quiz 

**Q - What is an appropriate code to calculate** $P(T < 1.2)$ **where** $T \sim t_5$**?**

```{r echo = F, out.width = "50%"}
cutoff <- 1.2
ggplot(NULL, aes(c(-6,6))) +
  geom_area(stat = "function", fun = dt, 
            args = list(df = 5), 
            fill = "#00998a", xlim = c(-6, cutoff), color = "black") +
  geom_area(stat = "function", fun = dt, 
            args = list(df = 5), 
            fill = "#FFFFFF", xlim = c(cutoff, 6), color = "black") +
  labs(x = "", y = "Density") +
  scale_x_continuous(breaks = seq(-6, 6, by = 2))
```
--

```{r}
pt(1.2, df = 5)
```

---

## Quiz 

**Q - What is an appropriate code to calculate** $P(-2 < T < 3)$ **where** $T \sim t_{10}$**?**

```{r echo = F, out.width = "50%"}
cutoff1 <- -2
cutoff2 <- 3
ggplot(NULL, aes(c(-6,6))) +
  geom_area(stat = "function", fun = dt, args = list(df = 10), 
            fill = "#FFFFFF", xlim = c(-6, cutoff1), color = "black") +
  geom_area(stat = "function", fun = dt, args = list(df = 10), 
            fill = "#00998a", xlim = c(cutoff1, cutoff2), color = "black") +
  geom_area(stat = "function", fun = dt, args = list(df = 10), 
            fill = "#FFFFFF", xlim = c(cutoff2, 6), color = "black") +
  labs(x = "", y = "Density") +
  scale_x_continuous(breaks = seq(-6, 6, by = 2))
```
--

```{r}
pt(3, df = 10) - pt(-2, df = 10)
```

---

## Quiz 

**Q - What is an appropriate code to find** q **s.t.** $P(T > q) = 0.25$ **where** $X \sim t_7$**?**

```{r echo = F, out.width = "50%"}
cutoff <- qt(0.25, df = 7, lower.tail = FALSE)
ggplot(NULL, aes(c(-6,6))) +
  geom_area(stat = "function", fun = dt, 
            args = list(df = 7), 
            fill = "#FFFFFF", xlim = c(-6, cutoff), color = "black") +
  geom_area(stat = "function", fun = dt, 
            args = list(df = 7), 
            fill = "#00998a", xlim = c(cutoff, 6), color = "black") +
  labs(x = "", y = "Density") +
  scale_x_continuous(breaks = seq(-6, 6, by = 2)) +
  geom_point(data = data.frame(x = cutoff, y = 0), aes(x,y)) + 
  geom_text(aes(x = 0.8, y = 0.02, label = "q = ?"))
```
--

```{r}
qt(0.25, df = 7, lower.tail = FALSE)
```

---

## Quiz: HT 

Let's conduct a hypothesis test for $H_0: \mu = 5$ vs. $H_1: \mu \neq 5$. 
We don't know the population standard deviation. We have a random sample of size 100. 
The CLT conditions are checked. 

**Q - What is the test statistic and its null distribution by CLT?**

--

- The test statistic is calculated by $t = \frac{\bar{x} - 5}{s/10}$.

--

- Under the null, 
$$T = \frac{\bar{X} - 5}{S/10} \sim t_{99}$$
- Capital letters for random variables and lowercase letters for observed values 

---

## Quiz: HT 

**Q - What does it mean that the test statistic is 3.5?**

--

The observed sample mean $\bar{x}$ is 3.5 standard errors above the hypothesized population mean, 5. 

---

## Quiz: CI

**Q - What is the formula to obtain a** $1-\alpha$ **confidence interval for** $\mu$ **?**

--

$$\bar{x} \pm t^*_{n-1}\times \frac{s}{\sqrt{n}}$$

where $t^*_{n-1}$ is a critical value that satisfies $P(T > t^*_{n-1}) = \alpha/2$ for $T \sim t_{n-1}$. 

--

**Q - What is the R function to calculate** $t^*_{n-1}$?

--

```{r, eval = F}
qt(alpha/2, df = n-1, lower.tail = FALSE)
```

---

## Quiz 

**Q - What is the function in the** `infer` **package to use for CLT-based inference when** $\sigma$ **is unknown?**

--

`t_test()`

---

class: middle, center

# Questions?

---

## Let's Practice Together! 

Go to [AE 18: Central Limit Theorem 2](https://sta199-summer22.netlify.app/appex/ae18_BJ.html)

---

## Bulletin

- Tomorrow is Ask-for-Help day. Bring your questions. 

- Lab06 due tonight at 11:59pm 

- HW03 due Wednesday, June 8 at 11:59pm 

- Tomorrow (June 8) is the last day to withdraw with W

- Submit `ae18`
