---
title: "Welcome to STA 199 Labs!"
date: "January 11, 2022"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

## Introductions

- Meet your TAs!
- Introduce yourself!

---

## What to expect in lab 

- Introduction to the lab assignment (~ 5 - 10 minutes, longer today)

- Work on the lab assignment (individual today, but generally done in teams)

- Lab instructions posted in the [schedule](https://sta199-sp22-003.netlify.app/schedule/) on the course website.

- Start each lab by finding your assignment repo in the [GitHub course organization](https://github.com/sta199-s22)
  - This is where you will find the R Markdown document and data to get started
  
---

## Tips

- You do not have to finish the lab in class, they will always be due the following **Monday at 11:59p**. One work strategy is to get through portions that you think will be most challenging (which initially might be the coding component) during lab when a TA can help you on the spot and leave the narrative writing until later.

- Do not pressure each other to finish early; use the time wisely to really learn the material and produce a quality report.

---

## GitHub setup

- If you did not previously, please follow [instructions to create a GitHub account and submit your username via google form](https://sta199-sp22-003.netlify.app/slides/01-intro.html#29)

- For this first lab, click [here](https://classroom.github.com/a/Febrwd1b) to create your `lab01` repository.

- Visit the [course organization GitHub page](https://github.com/sta199-s22) to find your new repository.

---

class: middle, center

# Demos

---

## Demos 

**Follow along as your TA demonstrates the following:** 

- Configure Git using SSH 
- Clone RStudio repo and start new project 
- Knit document and produce PDF 
- Update name in YAML 
  - Knit, commit, push changes to GitHub 
  - See updates in your GitHub repo
  
---

class: middle 

## You're now ready to complete the rest of lab!

<br>

## Please raise your hand if you need help as you work on the lab.
  

