---
title: "Lab #07: CLT-based Inference"
subtitle: "due Friday, June 10 at 11:59pm"
output: 
  tufte::tufte_html:
    css: "./sta199-labs.css"
    tufte_variant: "envisioned"
    highlight: pygments
    toc: false
    toc_depth: 1
link-citations: yes
editor_options: 
  chunk_output_type: inline
---

```{r include = FALSE}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)
```

# Goals

- Use inference based on the Central Limit Theorem to draw conclusions about a population mean.
- Compare inference based on Central Limit Theorem to simulation-based inference.

# Getting started

- Go to [course GitHub organization page](https://github.com/sta199-summer22) and clone the repository entitled "lab07-GitHubUsername" in RStudio. 
- Find `lab07.Rmd` to open the template R Markdown file. 

For each exercise:

- Show all relevant code and output used to obtain your response. If you use inline code, make sure we can still see the code used to derive that answer.

- Write all narrative in complete sentences, and use clear axis labels and titles on visualizations.

- Use a small number of `reps` (about 500) as you write and test out your code. 
Once you have finalized all of your code, increase the number of `reps` to 15,000 
to produce your final results.
    
- For each simulation exercise, use the **seed** specified in the exercise instructions.

# Packages 

We will use the **tidyverse**, **tidymodels**, **knitr** packages in this lab.

```{r}
library(tidyverse)
library(tidymodels)
library(knitr)
```

# Burrito

What makes a good burrito?

Today's dataset has been adapted from Scott Cole's [Burritos of San Diego 
project](https://srcole.github.io/100burritos/). The goal 
of the project was to identify the best and worst burritos in San Diego, 
characterize variance in burrito quality, and generate predictive models for 
what makes a burrito great. 

As part of this project, 71 participants reviewed burritos from 79 different 
taco shops. Reviewers captured objective measures of the burrito (such as 
whether it contains certain ingredients) and reviewed it on a number of metrics 
(such as quality of the tortilla, the temperature, quality of meat, etc.). For 
the purposes of this lab, you may consider each of these observations to be an
independent and representative sample of **all burritos.**

The subjective ratings in the dataset are as follows. Each variable is ranked
on a 0 to 5 point scale, with 0 being the worst and 5 being the best. 

- `tortilla`: quality of the tortilla
- `temp`: temperature of the burrito
- `meat`: quality of the meat
- `fillings`: quality of non-meat fillings
- `salsa`: quality of the salsa
- `mfr`: meat-to-filling ratio
- `uniformity`: whether each bite contains a uniform slew of ingredients (e.g.,
a bite entirely composed of tortilla and sour cream would probably be terrible)
- `synergy`: how well it all comes together

In addition, the reviewers noted the presence of the following burrito
components. Each of the following variables is a binary variable taking on 
values `present` or `none`:

- `guac`: guacamole
- `cheese`: cheese
- `fries`: fries (it's a thing, [look it up.](https://en.wikipedia.org/wiki/Burrito#San_Diego))
- `sourcream`: sour cream
- `rice`: rice
- `beans`: beans

The data are available in `burritos.csv`

The goal of this analysis is to make inference about the mean synergy rating of burritos based on the Central Limit Theorem (CLT).

1. We'll start by examining the distribution of `synergy`, a rating indicating how well all the ingredients in the burrito come together. 

  - Visualize the distribution of `synergy` using a histogram with binwidth of 0.5. 
  
  - Does the distribution look "normal"? Comment on the shape of the distribution.
  
  - Calculate the following summary statistics: the mean synergy, standard deviation of synergy, and sample size. Save the summary statistics as `summary_stats`. Then display `summary_stat` with `kable()`. 

2. The goal of this analysis is to use CLT-based inference to understand the true mean synergy rating of all burritos. What is your "best guess" for the mean synergy rating of burritos?


**Is the synergy in burritos generally good?** To answer this question, we will conduct a hypothesis test to evaluate if the mean synergy score of all burritos is greater than 3.

3. Before conducting inference, we need to check the conditions to make sure the CLT can be applied in this analysis. For each condition, indicate whether it is satisfied and provide a brief explanation supporting your response. 

  - Independence?
  - Sample size / distribution? 
  
4. State the null and alternative hypotheses to evaluate the question posed in the previous exercise. Write the hypotheses in words and in statistical notation. Clearly define any parameter you introduce. 

5. Let $\bar{X}$ be a random variable for the mean synergy score in a sample of 330 randomly selected burritos. Given the CLT and the hypotheses from the previous exercise,

  - What is the mean of the sampling distribution of $\bar{X}$ under the null? 
  - What is the standard error of the sampling distribution of $\bar{X}$? Assume the true $\sigma$ is 1. 
  - Write the null distribution of $\bar{X}$ concisely in math notation, i.e. $\bar{X} \sim N( , )$. You may use the math symbols given below: 
  
  ```
  $\bar{X} \sim N( , )$
  ```
  
In practice, we never know the true value of $\sigma$, so we estimate it with the observed standard deviation $s$ from our data. Consequently, the null distribution would slightly change. 

6. Recall the formula for the test statistic: 

$$t = \frac{\bar{x}- \mu_{0}}{\hat{se}}$$
where $\mu_0$ is the null value of $\mu$ (the specified value in the null hypothesis for $\mu$), and $\hat{se}$ is the estimated standard error of $\bar{X}$. $\hat{se}$ is calculated by replacing $\sigma$ in the standard error you found in Ex 5 with $s$. 

  - Use R as a "calculator" to calculate the test statistic, $t$. 
  - Explain what this value means in the context of this analysis. 
  - What is the distribution of the test statistic under the null? Be specific.

7. Now let's calculate the p-value and draw a conclusion. 

  - Use the `pt()` function to calculate the p-value. 
  - Explain what this value means in the context of this analysis. 
  - State your conclusion in the context of the data using a significance level of $\alpha = 0.01$.

8. We also want to calculate a 90% confidence interval for the mean synergy rating of all burritos. The confidence interval for a population mean is 

$$\bar{x} \pm t^*_{n-1} \times \hat{se}$$

We already know $\bar{x}$ and $\hat{se}$, so let's focus on the calculating $t^*_{n-1}$. We will use the `qt()` function to calculate the critical value $t^*_{n-1}$. 

```{marginfigure}
Here is an example: If we want to calculate a 95% confidence interval for the mean, we will use `qt(0.975, n-1)`, where 0.975 is the cumulative probability at the upper bound of the 95% confidence interval (recall we used this value to find the upper bound when calculating bootstrap confidence intervals), and (n-1) are the degrees of freedom.
```

  - Calculate the critical value, $t^*_{n-1}$, of the 90% confidence interval for the mean synergy rating of all burritos. 
  - Use R as a "calculator" to calculate the 90% confidence interval. 
  - Interpret the interval in the context of the data. 

9. In the previous exercises, we conducted a hypothesis test and calculated a confidence interval step-by-step. We can also use the `infer` package for the calculations in CLT-based inference using the `t_test()` function. 

  - Fill in the code below to produce the calculations for the hypothesis test you conducted in Ex 4 - 7.

```{marginfigure}
 The results should be the same as the calculations you did in the previous exercises.
```

```{r eval= F}
burrito %>%
  t_test(response = _____, 
         alternative = "______", 
         mu = ______, 
         conf_int = FALSE)
```

  - Next, fill in the code below to produce the 90% confidence interval for the mean synergy rating. 

```{marginfigure}
The results should be the same as the calculations in Ex 8.
```

```{r eval= F}
burrito %>%
  t_test(response = _____, 
         conf_int = _____, 
         conf_level = _____) %>%
  select(lower_ci, upper_ci)
```

# Submission

- Once you are finished with the lab, you will submit the PDF document produced from your final knit, commit, and push to [Gradescope](https://www.gradescope.com/courses/385955).
- Associate the "Workflow & formatting" graded section with the first page of your PDF, and mark the page associated with each exercise. If any answer spans multiple pages, then mark all pages.
- Follow the instructions in previous labs to submit your PDF to Gradescope.

# Grading (50 pts)

<br>

| Component             | Points |
|:----------------------|:-------|
| Ex 1                  |   6    |
| Ex 2                  |   1    |
| Ex 3                  |   3    |
| Ex 4                  |   5    |
| Ex 5                  |   3    |
| Ex 6                  |   6    |
| Ex 7                  |   6    |
| Ex 8                  |   8    |
| Ex 9                  |   5    |
| Workflow & formatting |   7    |

**Grading notes**:

-   The "Workflow & formatting" grade is to assess the reproducible workflow. This includes updating the name on the assignment to your name, having at least 3 informative commit messages, labeling the code chunks, and having readable code that does not exceed 80 characters (i.e., we can read all your code in the knitted PDF.)
